### 2025-02-18 paper.tex
first commit

%%
%% This is file `esapub.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% esapub.dtx  (with options: `manual')
%% ============================================
%% This is the manual describing the usage of
%%      esapub.cls
%% ============================================
%% Copyright 1999 Patrick W Daly
%% Max-Planck-Institut f\"ur Aeronomie
%% Max-Planck-Str. 2
%% D-37191 Katlenburg-Lindau
%% Germany
%% E-mail: daly@linmpi.mpg.de
%%
%% -------------------------------------------------
\documentclass[a4paper,twocolumn]{spaceDebrisC} % European paper
\pagestyle{empty}

% \bibliographystyle{alpha}
\bibliographystyle{apalike}

\usepackage{times}
\usepackage[numbers]{natbib}
\usepackage{graphicx}

% introduce this option for the ESA publications style

\title{Optimal Light Curve Attitude Inversion with Measurement Noise: Two Case Studies}
\author{Liam Robinson}
\author{Carolin Frueh}
\affil{Purdue University, West Lafayette, United States, Email: \texttt{\{robin502, cfrueh\}$@$purdue.edu}}

\begin{document}

\keywords{Light curve inversion; Atitude estimation; Photometry; Inverse problems}

\maketitle

\begin{abstract}

Understanding the orientation and angular velocity state of active or passive human-made space objects is critical for many operations like long-term orbit propagation, aiding a mission, determining mission status, and active debris removal. Beyond low-Earth orbit, optical telescopes are used predominantly to track the space objects. Although this data is abundant, it is generally impossible to resolve even large satellites and rocket bodies in optical ground-based imagery due aperture size and atmospheric distortion. As a result, shape and orientation information is not available in any single image. But a measurement of the object's total brightness can still be obtained. Even if the object's shape and reflective properties are known, any given brightness measurement generally corresponds to infinitely many possible orientations. To constrain the solution space, the brightness can be tracked over time -- producing a sequence of measurements called a light curve. If the identity of the object is known, its attitude profile can be estimated from the light curve through a process known as light curve attitude inversion. Due to the natural and instrumental noise in the light curve, as well as symmetries in the observation geometry, there may be multiple attitude profiles which are equally likely to produce any given light curve.

In this work, we apply a light curve attitude inversion algorithm to real observations taken by the Purdue Optical Ground Station. Given an estimate of the shape and reflective properties of the object, our procedure searches the space of initial orientations, angular velocities, and inertia tensors to find initial conditions that produce light curves with low error compared to the observed values. Specifically, we estimate a full distribution for each candidate light curve and compute the likelihood of the observations being a sample from that distribution. To solve the optimization problem, we create a number of sampled candidate solutions, deterministically arranged throughout the solution space, and perform nonlinear optimization on each using the BFGS algorithm. This way, our optimization procedure reliably converges to all maximum likelihood estimates of the attitude state. Low-error minima are then clustered to identify families of nearby solutions with similar likelihood. The final output of our method is an array of the best solutions, ranked by their likelihood. Formulating the problem this way inherently accounts for the multiple classes of ambiguities resulting from the object and observation geometry, the physical constraints of torque-free rigid body motion, as well as the noise on the light curve.

To test the efficacy of this method, we perform attitude inversion on real observations gathered by the Purdue Optical Ground Station of an inactive GLONASS satellite and a rocket body. Solutions for the GLONASS case are analyzed to highlight the effect of observation geometry ambiguities, while the rocket body case highlights ambiguities introduced by the geometry of the object. Our noise model is validated against further real observations, confirming that all relevant noise sources have been accounted for and modeled accurately. We show that any significant lack of knowledge in the object shape or optical material properties causes the solution to rapidly degrade.

\end{abstract}

\section{General Specifications}

test

\section*{Acknowledgments}

people \cite{robinson2023}

\bibliography{cache/refs.bib}

% \begin{thebibliography}{}

% \bibitem{author73}
% Author T., (1973). Astrophysical Quantities, Athlone Press

% \bibitem{nobody97}
% Nobody B., Somebody G., Who D., et~al., (1997). {\it The book}, Publisher, ed. 2

% \bibitem{smith96}
% Smith A., Jones B., (1996). The new discovery, {\it Other Journal}, {\bf 223}(1), 1029--1101

% \end{thebibliography}
\end{document}

### 2025-02-18 copy_to_dropbox.py

@@ -1,0 +1,8 @@
+import os
+import shutil
+import datetime
+
+dtime = str(datetime.datetime.now().date())
+dest_folder = "/Users/liamrobinson/Library/CloudStorage/Dropbox/progress_PhD/papers/for-review/"
+dest_path = os.path.join(dest_folder, f"digital-twin-{dtime}.pdf")
+shutil.copy("final/paper-final.pdf", dest_path)

### 2025-02-18 figure_normalize.py

@@ -1,0 +1,96 @@
+import re
+import os
+import glob
+import shutil
+from PIL import Image  # Requires Pillow package
+
+output_path = 'final/paper.tex'
+final_refs_name = 'refs.bib'
+
+shutil.rmtree('final')
+os.mkdir('final')
+
+os.system('cp *.sty final')
+os.system('cp *.eps final')
+os.system('cp *.bst final')
+os.system('cp *.cls final')
+
+with open('paper.tex', 'r') as file:
+    content = file.read()
+    content = '\n'.join([x for x in content.splitlines() if not x.startswith('%')])
+
+# also remove the graphicspath declaration
+content = '\n'.join([x for x in content.splitlines() if not x.startswith(r'\graphicspath')])
+
+# also adjust the bibliography declaration
+content_new = []
+for line in content.splitlines():
+    if not line.startswith(r'\bibliography{'):
+        content_new.append(line)
+    else:
+        content_new.append(r'\bibliography{' + final_refs_name + '}')
+content = '\n'.join(content_new)
+
+# Define the regex pattern to find \includegraphics commands
+pattern = r'(\\includegraphics\[.*?\]{)(.*?)(})'
+
+# Find all matches and replace them sequentially
+matches = re.findall(pattern, content)
+for i, match in enumerate(matches, start=1):
+    print(match)
+    # img = Image.open(f'cache/{match[1]}')
+    # img.convert('RGB').save(f'final/fig{i}.jpg', format='JPEG', dpi=(600,600))
+    # replacement = rf"{match[0]}fig{i}.jpg{match[2]}"
+
+    os.system(f'cp cache/{match[1]} final/fig{i}.png')
+    replacement = rf"{match[0]}fig{i}.png{match[2]}"
+    content = content.replace("".join(match), replacement, 1)
+
+# Write the updated content to a new file
+with open(output_path, 'w') as file:
+    file.write(content)
+
+### DEALING WITH CITATIONS
+
+def clean_unused_references(tex_file, bib_file, output_bib_file):
+    # Step 1: Extract citation keys from the LaTeX file
+    with open(tex_file, 'r') as file:
+        tex_content = file.read()
+
+    # Define regex patterns for \citep and \cite
+    pattern = r'\\(citep|cite)\{([^}]+)\}'
+
+    # Find all matches
+    matches = re.findall(pattern, content)
+
+    # Process matches
+    used_citations = []
+    for cite_type, cite_args in matches:
+        # Split the comma-separated arguments
[…]

### 2025-02-18 paper_cache.py

@@ -1,0 +1,33 @@
+import os
+import shutil
+import glob
+
+
+def cache_file(start_path: str, end_dir: str = "cache"):
+    if not os.path.exists(end_dir):
+        os.mkdir(end_dir)
+    im_name = os.path.split(start_path)[1]
+    end_path = os.path.join(end_dir, im_name)
+    if not os.path.exists(start_path):
+        raise ValueError(f"{start_path} does not exist!")
+    shutil.copy(start_path, end_path)
+
+
+imgs = {
+    "/Users/liamrobinson/Documents/maintained-research/mirage/docs/build/html/_images": [
+    ],
+    "/Users/liamrobinson/Documents/presentations-papers/msthesis/static_images": [
+    ],
+    "/Users/liamrobinson/Documents/maintained-research/mirage/docs/source/_static": [
+        "refs.bib"
+    ],
+}
+
+for im_dir, im_names in imgs.items():
+    for im_name in im_names:
+        im_path = os.path.join(im_dir, im_name)
+        globs = glob.glob(im_path)
+        if not globs:
+            raise ValueError(f"Nothing found for pattern {im_name}")
+        for f in globs:
+            cache_file(start_path=f)

### 2025-02-18 paper.tex
added some eqs

@@
 
-% introduce this option for the ESA publications style
+\usepackage{bm} % for bold math
+\usepackage{mathtools} % for \prescript
+\newcommand{\vctr}[1]{\bm{#1}}
+\newcommand{\unitv}[1]{\hat{\vctr{#1}}}
+\newcommand{\preup}[1]{\prescript{#1}{}{}}
+\newcommand{\rf}[1]{\mathcal{\MakeUppercase{#1}}}
+\newcommand{\dcm}[1]{\left[\rf{#1}\right]}
+\newcommand{\vrf}[2]{\preup{\rf{#1}}\vctr{#2}}
+\newcommand{\mbar}[0]{\;\middle|\;}
+\newcommand{\prf}[1]{\preup{\rf{#1}}}
+\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
+
+\newcommand{\matcp}[1]{\left[#1 \times\right]}
+\newcommand{\rthree}[0]{$\mathbb{R}^3$\:}
+\newcommand{\sthree}[0]{$\mathbb{S}^3$\:}
+\newcommand{\sothree}[0]{$SO_3$}
+\newcommand{\pogslla}[0]{$32.900^\circ \textrm{ N}, -105.533^\circ \textrm{ W} \textrm{ }$}
+
+\newcommand{\fighuge}[0]{1.0\textwidth}
+\newcommand{\figbig}[0]{0.8\textwidth}
+\newcommand{\figmed}[0]{0.6\textwidth}
+\newcommand{\figsmall}[0]{0.4\textwidth}
+
+\newcommand{\subfigmed}[0]{0.4\textwidth}
+\newcommand{\subfigsm}[0]{0.29\textwidth}
+\newcommand{\subfigbig}[0]{0.43\textwidth}
+
+\DeclareMathOperator{\atantwo}{atan2}
+\DeclareMathOperator{\fl}{floor}
 
@@
 
-\section{General Specifications}
+\section{Introduction}
 
 test
+
+\section{Inversion Method}
+
+\begin{equation} \label{eq:nll_loss}
+  f_\text{NLL}(\hat{\vctr{S}}, \vctr{S}) = \frac{1}{m}\sum_{k=1}^{m}\left[\ln\sigma_k + \frac{1}{2}\left(\frac{S_k - \hat{S}_k}{\sigma_k}\right)^2 \right].
+ \end{equation}
+
+\begin{equation}
+  \vctr{x} = \begin{bmatrix} 
+  p_1 & p_2 & p_3 & \omega_1 & \omega_2 & \omega_3 & \frac{J_y}{J_x} & \frac{J_z}{J_x}
+  \end{bmatrix}^T.
+ \end{equation}
+
+\begin{align}
+  \vctr{\dot{\sigma}} &= \frac{1}{4} \left[ \left(1 - \vctr{p} \cdot \vctr{p}\right) + 2\vctr{p} \times \vctr{\omega} + 2 \left(\vctr{\omega} \cdot \vctr{p} \right)\vctr{p} \right] \label{eq:mrp_kde} \\
+  \vctr{\dot{\omega}} &= J^{-1} \left[ \left(J \vctr{\omega}\right) \times \vctr{\omega} + \vctr{M}\right] \label{eq:rbtf_dynamics}
+\end{align}
+
+\begin{equation}
+  \vctr{p}'' = \frac{\left(1-\norm{\sigma'}^2\right)\vctr{p} - \left(1-\norm{\vctr{p}}^2\right)\vctr{p}' - 2\vctr{p} \times \vctr{p}'}{1 + \norm{\sigma'}^2 \norm{\vctr{p}}^2 + 2\vctr{p}' \cdot \vctr{p}},
+\end{equation}
+
+\begin{equation} \label{eq:angle_error}
+  \theta = 4 \tan^{-1} \left( \norm{\vctr{p}''} \right).
+\end{equation}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figmed]{obs_geom.png}
+  \caption{Facet geometry including the observer direction $\unitv{o}$, Sun direction $\unitv{s}$, normal vector $\unitv{n}_i$, unshadowed area $\bar{a}_i$, shadowed area $\tilde{a}_i$, and counterclockwise vertex locations $\left\{ \vctr{v}_{i,1}, \vctr{v}_{i,2}, \vctr{v}_{i,3} \right\}$.
+}
+  \label{fig:facet_geom}
+\end{figure}
+
+\begin{equation} \label{eq:us_area}
+  \bar{a} = a - \sum_{d=1}^{n} \sum_{K \in \text{comb}(n,d)} (-1)^{d+1} A\left(\bigcap\limits_{k \in K} P_k\right).
+\end{equation}
+
+\begin{equation} \label{eq:brdf_blinn_phong}
+  f_r(\unitv{s}, \unitv{o}) = \frac{C_d}{\pi} + \frac{n+2}{2\pi} \frac{C_s (\unitv{n} \cdot \unitv{h})^n}{4 (\unitv{n} \cdot \unitv{s})(\unitv{n} \cdot \unitv{o})}.
+ \end{equation}
+
+ \begin{equation} \label{eq:lc_norm}
+  f_p(\unitv{s}_k, \unitv{o}_k) = \sum_{i=1}^n{\prf{B}\left[\bar{a}_i(\unitv{s}_k, \unitv{o}_k)
+  f_{r,i}(\unitv{s}_k, \unitv{o}_k)
+   (\unitv{n}_i \cdot \unitv{s}_k)
+   (\unitv{n}_i \cdot \unitv{o}_k)
+   \right]}. 
+ \end{equation}
+
+ \begin{equation} \label{eq:general_bright}
+  \bar{S}_{\text{SO},k} = f_p(\unitv{s}_k, \unitv{o}_k) \frac{A \cdot \Delta t_k \cdot f_\odot(\vctr{r})}{g R_\oplus^2 r_k^2} \int_{0}^{\infty}{P(\lambda)Q(\lambda)T_k(\lambda) I_\odot(\lambda) \left(\frac{\lambda}{hc}\right)}\,d\lambda.
+\end{equation}
+
+\begin{equation} \label{eq:sensor_noise}
+  \sigma_\text{sensor} = \sqrt{n_\text{pix} \left( \Delta t \cdot \lambda_\text{dark} + \sigma_\text{read}^2 \right)}.
+\end{equation}
+
+\begin{equation}
+  \sigma_\text{flat} = f_k \sqrt{\sum_{i=1}^{n_\text{pix}} S_i^2}.
+\end{equation}
+
+\begin{equation}
+  \lambda_{\text{back},k} = n_{\text{pix},k} \left( \lambda_{\text{moon},k} + \lambda_{\text{star},k} + \lambda_{\text{twi},k} + \lambda_{\text{zod},k} + \lambda_{\text{star},k} + \lambda_{\text{poll},k} \right).
+\end{equation}
+
+\begin{equation} \label{eq:scint_noise}
+  \sigma^2_{Y,k} = 10^{-6} D^{-4/3} (\Delta t)_k^{-1} \cos^{-3}\left(\gamma_k\right) e^{\frac{-2h_\text{obs}}{H}}.
+\end{equation}
+
+\begin{equation}
+  \lambda_{\text{shot},k} = \frac{\bar{S}_{\text{SO},k}}{\sqrt{g}}.
+\end{equation}
+
+\begin{equation} \label{eq:sigma_total}
+  \sigma^2_{S,k} = \lambda_{\text{back},k} + \bar{S}_{\text{SO},k} \sigma^2_{Y,k} + \lambda_{\text{shot},k} + \sigma^2_\text{flat},
+\end{equation}
+
+\begin{equation}
+  S_{\text{SO},k} \sim \mathcal{N}\left( \bar{S}_{\text{SO},k}, \sigma^2_{S,k} \right).
+ \end{equation}
+
+ \begin{table}[ht]
+  \centering
+  \caption{Observation parameters 2}
+  \vspace*{6pt}
+  \begin{tabular}{|l|l|}
+  \hline
+  \textbf{Variable} & \textbf{Value} \\ \hline
+  True inertia ratios $\begin{bmatrix}J_y / J_x & J_z / J_x\end{bmatrix}$ & $\begin{bmatrix} 1.0 & 0.5 \end{bmatrix}$ \\ \hline
+  Initial simulation date & October 10, 2024 08:00:00.000 UTC \\ \hline
+  Object identity (COSPAR ID) & 1987-084D \\ \hline 
+ Light curve duration & $2$ minutes \\ \hline
+ Measurement frequency & $0.5$ Hz \\ \hline
+ Integration time & $0.01$ seconds \\ \hline
+  \end{tabular}
+  \label{tb:case1_in}
+\end{table}
+
+\begin{table}[ht]
+  \centering
+  \caption{Observation parameters 2}
+  \vspace*{6pt}
+  \begin{tabular}{|l|l|}
+  \hline
+  \textbf{Variable} & \textbf{Value} \\ \hline
+  True inertia ratios $\begin{bmatrix}J_y / J_x & J_z / J_x\end{bmatrix}$ & $\begin{bmatrix} 1.0 & 0.5 \end{bmatrix}$ \\ \hline
+  Initial simulation date & October 10, 2024 08:00:00.000 UTC \\ \hline
+  Object identity (COSPAR ID) & 1987-084D \\ \hline 
+ Light curve duration & $2$ minutes \\ \hline
+ Measurement frequency & $0.5$ Hz \\ \hline
+ Integration time & $0.01$ seconds \\ \hline
+  \end{tabular}
+  \label{tb:case1_in}
+\end{table}
 

### 2025-02-18 copy_to_dropbox.py

@@ -4,5 +4,5 @@ import datetime
 
 dtime = str(datetime.datetime.now().date())
 dest_folder = "/Users/liamrobinson/Library/CloudStorage/Dropbox/progress_PhD/papers/for-review/"
-dest_path = os.path.join(dest_folder, f"digital-twin-{dtime}.pdf")
+dest_path = os.path.join(dest_folder, f"sdc-attitude-{dtime}.pdf")
 shutil.copy("final/paper-final.pdf", dest_path)

### 2025-02-18 paper_cache.py

@@ -21,6 +21,8 @@ imgs = {
     "/Users/liamrobinson/Documents/maintained-research/mirage/docs/source/_static": [
         "refs.bib"
     ],
+    '/Users/liamrobinson/Documents/maintained-research/mirage/examples/13-attitude-2024/figs':
+    glob.glob('/Users/liamrobinson/Documents/maintained-research/mirage/examples/13-attitude-2024/figs/*.png'),
 }
 
 for im_dir, im_names in imgs.items():

### 2025-02-21 paper.tex
added some method

--- 2025-02-18 paper.tex
@@
 
-% introduce this option for the ESA publications style
+\usepackage{bm} % for bold math
+\usepackage{mathtools} % for \prescript
+\newcommand{\vctr}[1]{\bm{#1}}
+\newcommand{\unitv}[1]{\hat{\vctr{#1}}}
+\newcommand{\preup}[1]{\prescript{#1}{}{}}
+\newcommand{\rf}[1]{\mathcal{\MakeUppercase{#1}}}
+\newcommand{\dcm}[1]{\left[\rf{#1}\right]}
+\newcommand{\vrf}[2]{\preup{\rf{#1}}\vctr{#2}}
+\newcommand{\mbar}[0]{\;\middle|\;}
+\newcommand{\prf}[1]{\preup{\rf{#1}}}
+\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
+
+\newcommand{\matcp}[1]{\left[#1 \times\right]}
+\newcommand{\rthree}[0]{$\mathbb{R}^3$\:}
+\newcommand{\sthree}[0]{$\mathbb{S}^3$\:}
+\newcommand{\sothree}[0]{$SO_3$}
+\newcommand{\pogslla}[0]{$32.900^\circ \textrm{ N}, -105.533^\circ \textrm{ W} \textrm{ }$}
+
+\newcommand{\figbig}[0]{0.4\textwidth}
+\newcommand{\figmed}[0]{0.4\textwidth}
+\newcommand{\figsmall}[0]{0.3\textwidth}
+
+\newcommand{\subfigmed}[0]{0.4\textwidth}
+\newcommand{\subfigsm}[0]{0.29\textwidth}
+\newcommand{\subfigbig}[0]{0.43\textwidth}
+
+\DeclareMathOperator{\atantwo}{atan2}
+\DeclareMathOperator{\fl}{floor}
 
@@
 
+\end{abstract}
+
+\section{Introduction}
+
+Knowledge of spacecraft orientation and angular velocity states is critical for many space situational awareness (SSA) tasks.
+Understanding the evolution of high area-to-mass ratio objects \cite{frueh2014} and decommissioned satellites \cite{rachman2023}, planning and executing active debris removal \cite{bonnal2013}, and recovering from mission anomalies \cite{umansky2023} all require attitude information.
+
+[\textit{TODO: lit}]
+
 In this work, we apply a light curve attitude inversion algorithm to real observations taken by the Purdue Optical Ground Station.
@@
 
-\end{abstract}
+\section{Inversion Method}
 
-\section{General Specifications}
+\subsection{Objective Function}
 
-test
+Our inversion method is based on the global minimization of a maximum-likelihood loss function which computes the negative log-likelihood of the observed light curve $\vctr{S}$ being a sample from the normal distribution defined by the estimated light curve mean $\hat{\vctr{S}}$ and the estimated standard deviation at each $k$th timestep $\sigma_k$:
+
+\begin{equation} \label{eq:nll_loss}
+  f(\hat{\vctr{S}}, \vctr{S}) = \frac{1}{m}\sum_{k=1}^{m}\left[\ln\sigma_k + \frac{1}{2}\left(\frac{S_k - \hat{S}_k}{\sigma_k}\right)^2 \right].
+ \end{equation}
+
+In this work, the light curve and its standard deviation are always assumed to be expressed in a linear unit, e.g., analog-to-digital units (ADU) or irradiance.
+
+\subsection{State Vector Definition and Dynamics}
+
+The state vector $\vctr{x}$ we optimize to fit the observed light curve is the concatenation of the Modified Rodriguez Parameter (MRP) $\vctr{p} = [p_1, p_2, p_3]^T$, the principal body frame angular velocity $\vctr{\omega} = [\omega_1, \omega_2, \omega_3]$ in radians per second, and the ratios of the principal moments of inertia $J_y / J_x$ and $J_z / J_x$ where the full inertia tensor is $J = \mathrm{diag}\left([J_x, J_y, J_z]\right)$:
+
+\begin{equation}
+  \vctr{x} = \begin{bmatrix} 
+  p_1 & p_2 & p_3 & \omega_1 & \omega_2 & \omega_3 & J_y / J_x & J_z / J_x
+  \end{bmatrix}^T.
+\end{equation}
+
+The first six elements of the state vector are time varying with dynamics given by:
+
+\begin{align}
+  \vctr{\dot{p}} &= \frac{1}{4} \left[ \left(1 - \vctr{p} \cdot \vctr{p}\right) + 2\vctr{p} \times \vctr{\omega} + 2 \left(\vctr{\omega} \cdot \vctr{p} \right)\vctr{p} \right], \label{eq:mrp_kde} \\
+  \vctr{\dot{\omega}} &= J^{-1} \left[ \left(J \vctr{\omega}\right) \times \vctr{\omega} + \vctr{M}\right]. \label{eq:rbtf_dynamics}
+\end{align}
+
+\section{Computing the Light Curve}
+
+\subsection{Object Geometry Definition}
+
+We choose to represent the surface geometry of space objects as a polyhedral mesh composed of a set of flat facets.
+Each $i$th face has $n_v \geq 3$ or more vertices $\left\{ v_{i,1}, v_{i,2}, v_{i,n_v} \right\}$ and an outward-pointing normal vector $\unitv{n}_i$.
+The total facet area is computed for an arbitrary planar polygon via:
+
+\begin{equation} \label{eq:poly_area}
+  A(P) = \sum_{j=0}^{k-3} \frac{\| \left( \vctr{v}_{j+1} - \vctr{v}_{j} \right) \times \left( \vctr{v}_{j+2} - \vctr{v}_{j} \right)\|_2}{2}.
+ \end{equation}
+
+ In any particular illumination and observation condition, a fraction of the facet's area $\tilde{a}_i$ may be shadowed due to obstructing geometry before or after  light reaches the surface element.
+The remainder of the facet's area $\bar{a}_i$ is unshadowed and contributes to the total flux reaching the observer from the object, such that the total area of each facet is $a_i = \tilde{a}_i + \bar{a}_i$.
+The geometry for an individual facet is displayed in Figure \ref{fig:facet_geom}.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figmed]{obs_geom.png}
+  \caption{Facet geometry including the observer direction $\unitv{o}$, Sun direction $\unitv{s}$, normal vector $\unitv{n}_i$, unshadowed area $\bar{a}_i$, shadowed area $\tilde{a}_i$, and counterclockwise vertex positions $\left\{ \vctr{v}_{i,1}, \vctr{v}_{i,2}, \vctr{v}_{i,3} \right\}$.}
+  \label{fig:facet_geom}
+\end{figure}
+
+\subsection{Computing the Unshadowed Area}
+
+The unshadowed area $\bar{a}$ may be computed in a number of ways.
+For convex objects, $\bar{a}=a$ as self-shadowing is impossible.
+For highly-nonconvex and detailed models composed of thousands of faces, $\bar{a}$ can be approximated on a pixel grid using shadow mapping \cite{robinson2022}.
+Self-shadowing can be solved semi-analytically for less detailed approximations of nonconvex objects by computing the mutual intersections of the polygons $P_k$ on other faces whose projections overlap with the face in question.
+Up to a maximum intersection depth $d$, the unshadowed area can be computed via:
+
+\begin{equation} \label{eq:us_area}
+  \bar{a} = a - \sum_{d=1}^{n} \sum_{K \in \text{comb}(n,d)} (-1)^{d+1} A\left(\bigcap\limits_{k \in K} P_k\right).
+\end{equation}
+
+As the objects we investigate in this work are well-approximated with relatively simple nonconvex meshes, we use the semi-analytic method to efficiently compute unshadowed areas via the Sutherland-Hodgeman algorithm \cite{sutherland1974} for each polygon intersection.
+
+\subsection{Surface Reflectivity}
+
+The bidirectional reflectance distribution function (BRDF) defines the amount of incident radiation from $\unitv{s}$ reflected per steradian in the observer's direction $\unitv{o}$.
+We choose the Blinn-Phong \cite{blinn1977} BRDF for this work as it is efficient to compute and satisfies the three main requirements for a physically-meaningful BRDF as it is nonnegative, energy-conserving, and reciprocal \cite{duvenhage2013}.
+The Blinn-Phong BRDF is parameterized by the coefficient of diffuse reflectivity $C_d$, the coefficient of specular reflectivity $C_s$, and the specular exponent $n>0$.
+The coefficients of reflectivity implicity satisfy $C_a + C_s + C_d = 1$ for energy conservation, where $C_a$ is the coefficient of absorption \cite{fan2020thesis}.
+
+\begin{equation} \label{eq:brdf_blinn_phong}
+  f_r(\unitv{s}, \unitv{o}) = \frac{C_d}{\pi} + \frac{n+2}{2\pi} \frac{C_s (\unitv{n} \cdot \unitv{h})^n}{4 (\unitv{n} \cdot \unitv{s})(\unitv{n} \cdot \unitv{o})}.
+\end{equation}
+
+Here, the halfway vector $\unitv{h}$ bisects the illumination and observation directions such that $\unitv{h} = \unitv{h} = (\unitv{s} + \unitv{o})/\norm{\unitv{s} + \unitv{o}}$ \cite{duvenhage2013}.
+
+\subsection{The Light Curve}
+
+Given the observer $\unitv{o}_k$ and illumination directions $\unitv{s}_k$ at the $k$th observation epoch in the body frame $\mathcal{B}$, as well as the BRDF $f_{r,i}$ for each $i$th surface of the object, the fraction of incident power reflected in the direction of the observer is \cite{fan2020thesis}:
+
+\begin{equation} \label{eq:lc_norm}
+  \begin{split} 
+  f_p(\unitv{s}_k, \unitv{o}_k) = \sum_{i=1}^n\prf{B}[&\bar{a}_i(\unitv{s}_k, \unitv{o}_k) f_{r,i}(\unitv{s}_k, \unitv{o}_k) \\ \cdot &(\unitv{n}_i \cdot \unitv{s}_k) (\unitv{n}_i \cdot \unitv{o}_k)].
+  \end{split} 
+\end{equation}
+
+The output of Equation \ref{eq:lc_norm} is a light curve, but as many of the noise characteristics of the signal are defined in image sensor's native unit of ADU, we convert the light curve into the same units via:
+
+\begin{equation} \label{eq:general_bright}
+  \begin{split} 
+  \bar{S}_{\text{SO},k} = &f_p(\unitv{s}_k, \unitv{o}_k) \frac{A_\circ \Delta t_k f_\odot(\vctr{r})}{g R_\oplus^2 r_k^2} \\ \cdot &\int_{0}^{\infty}{P(\lambda)Q(\lambda)T_k(\lambda) I_\odot(\lambda) \left(\frac{\lambda}{hc}\right)}\,d\lambda.
+  \end{split}
+\end{equation}
+
+Here before the integral, $A_\circ$ is the unobstructed aperture area in square meters, $\Delta t_k$ is the exposure time in seconds, $f_\odot(\vctr{r})$ is the fraction of solar irradiance reaching the space object at position $\vctr{r}$ -- accounting for the Earth's shadow, $g$ is the sensor gain in ADU per photoelectron, $R_\oplus$ is the distance from the Sun to the space object in AU, $r_k$ is the distance from the observer to the space object in kilometers.
+Within the integral, we account for the telescope's passband filter $P(\lambda)$, the image sensor quantum efficiency $Q(\lambda)$, the atmospheric absorption spectrum $T_k(\lambda)$, the solar irradiance spectrum $I_\odot(\lambda)$, and the inverse energy of a photon with wavelength $\lambda$, $\lambda / hc$, where $h$ is Planck's constant in Joule-seconds, and $c$ is the speed of light in vacuum in meters per second.
+Taken together, the integral computes the fraction of the energy reflected from the object absorbed into the image sensor, while the the outer factor dimensionalizes the result to yield the mean total sensor response in ADU across all pixels due to the object's signal.
+
+the variance a space object's light curve whose mean is determined by Equation \ref{eq:general_bright} is a combination of a number of relevant independent stochastic processes.
+These distributions have variances $\sigma^2_\text{sensor}$ due to the sensor's integration and readout effects, $\sigma^2_\text{flat}$ from sensor flat-fielding effects, scintillation noise $\bar{S}_{\text{SO},k} \sigma^2_{Y,k}$ \cite{osborn2015}, Poisson signal shot noise $\lambda_{\text{shot},k}$, and Poisson background noise $\lambda_{\text{back},k}$.
+The sum of these variances yields total signal variance in ADU:
+
+\begin{equation} \label{eq:sigma_total}
+  \begin{split}
+  \sigma^2_{S,k} = &\lambda_{\text{back},k} + \bar{S}_{\text{SO},k} \sigma^2_{Y,k} + \lambda_{\text{shot},k} \\ + &\sigma^2_\text{flat} + \sigma^2_\text{sensor}.
+  \end{split}
+\end{equation}
+
+The sensor noise is approximated by the independent combination of Poisson dark current $\lambda_\text{dark}$ and readout noise $\sigma_\text{read}^2$ \cite{krag2003}:
+
+\begin{equation} \label{eq:sensor_noise}
+  \sigma_\text{sensor}^2 = n_\text{pix} \left( \Delta t \cdot \lambda_\text{dark} + \sigma_\text{read}^2 \right).
+\end{equation}
+
+The flat field noise is modeled as a zero-mean Gaussian linearly scaling with the signal in each of the $n_\text{pix}$ pixels of the object signal $S_i$ and a constant $f_k$ fit to the sensor from flat frame observations, yielding the signal standard deviation \cite{newberry1996}:
+
+\begin{equation}
+  \sigma_\text{flat}^2 = f_k^2 \sum_{i=1}^{n_\text{pix}} S_i^2.
+\end{equation}
+
+The background standard deviation is modeled by the sum of six independent Poisson random processes contributing to light entering the telescope optics from environmental sources.
+These processes and the sources of their respective models are: scattered moonlight $\lambda_{\text{moon},k}$ \cite{daniels1977}, integrated starlight $\lambda_{\text{star},k}$ \cite{krag2003}, twilight $\lambda_{\text{twi},k}$ \cite{pata2006}, zodiacal light $\lambda_{\text{zod},k}$ \cite{roach1972}, airglow $\lambda_{\text{air},k}$ \cite{krag2003}, and light pollution $\lambda_{\text{poll},k}$ \cite{falchi2016, falchi2016_data}.
+In summation, the total Poisson background variance is:
+
+\begin{equation}
+  \begin{split}
+  \lambda_{\text{back},k} = n_{\text{pix},k} ( &\lambda_{\text{moon},k} + \lambda_{\text{star},k} + \lambda_{\text{twi},k} \\+ &\lambda_{\text{zod},k} + \lambda_{\text{air},k} + \lambda_{\text{poll},k} ).
+  \end{split}
+\end{equation}
+
+The fractional scintillation noise due to atmospheric turbulence is modeled via Young's approximation \cite{osborn2015}:
+
+\begin{equation} \label{eq:scint_noise}
+  \sigma^2_{Y,k} = 10^{-6} D^{-4/3} (\Delta t)_k^{-1} \cos^{-3}\left(\gamma_k\right) e^{\frac{-2h_\text{obs}}{H}}.
+\end{equation}
+
+[\textit{TODO: describe the variables here, if we want to keep this eq... it's in the cited paper so i am not sure}]
+
+The signal shot noise is a Poisson process as a function of the mean signal in ADU $\bar{S}_{\text{SO},k}$ and the image sensor gain $g$ in ADU per photoelectron:
+
+\begin{equation}
+  \lambda_{\text{shot},k} = \frac{\bar{S}_{\text{SO},k}}{\sqrt{g}}.
+\end{equation}
+
+After summation in Equation \ref{eq:sigma_total}, we can approximate the observed light curve as Gaussian distributed via the mean $\bar{S}_{\text{SO},k}$ from Equation \ref{eq:general_bright} and the variance $\sigma^2_{S,k}$ in $[\text{ADU}^2]$:
+
+\begin{equation}
+  S_{\text{SO},k} \sim \mathcal{N}\left( \bar{S}_{\text{SO},k}, \sigma^2_{S,k} \right).
+ \end{equation}
+
+ \begin{table}[ht]
+  \centering
+  \caption{Observation parameters 2}
+  \vspace*{6pt}
+  \begin{tabular}{|l|l|}
+  \hline
+  \textbf{Variable} & \textbf{Value} \\ \hline
+  Observation start (UTC) & Oct 10, 2024 08:00:00 \\ \hline
+  COSPAR ID & 1987-084D \\ \hline 
+ Light curve duration & $2$ minutes \\ \hline
+ Measurement frequency & $0.5$ Hz \\ \hline
+ Integration time & $0.01$ seconds \\ \hline
+  \end{tabular}
+  \label{tb:case1_in}
+\end{table}
+
+\begin{table}[ht]
+  \centering
+  \caption{Observation parameters 2}
+  \vspace*{6pt}
+  \begin{tabular}{|l|l|}
+  \hline
+  \textbf{Variable} & \textbf{Value} \\ \hline
+  Observation start (UTC) & Oct 10, 2024 08:00:00 \\ \hline
+  COSPAR ID & 1987-084D \\ \hline 
+ Light curve duration & $2$ minutes \\ \hline
+ Measurement frequency & $0.5$ Hz \\ \hline
+ Integration time & $0.01$ seconds \\ \hline
+  \end{tabular}
+  \label{tb:case1_in}
+\end{table}
 
@@
 
-people \cite{robinson2023}
+This work was partially supported by a National Defense Science and Engineering Graduate Fellowship.
+
+[\textit{TODO: Acknowledge AIUB if we end up using some of their data for a GLONASS light curve}]
 

### 2025-02-27 paper.tex
added some initial results/obs

@@
 
-\newcommand{\figbig}[0]{0.4\textwidth}
+\newcommand{\figbig}[0]{0.5\textwidth}
 \newcommand{\figmed}[0]{0.4\textwidth}
@@
 Specifically, we estimate a full distribution for each candidate light curve and compute the likelihood of the observations being a sample from that distribution.
-To solve the optimization problem, we create a number of sampled candidate solutions, deterministically arranged throughout the solution space, and perform nonlinear optimization on each using the BFGS algorithm.
+To solve the optimization problem, we create a number of sampled candidate solutions, scattered throughout the solution space, and perform nonlinear optimization on each using the BFGS algorithm.
 This way, our optimization procedure reliably converges to all maximum likelihood estimates of the attitude state.
@@
 
-In this work, the light curve and its standard deviation are always assumed to be expressed in a linear unit, e.g., analog-to-digital units (ADU) or irradiance.
+In this work, the light curve and its standard deviation are always assumed to be expressed in a linear unit, e.g., analog-to-digital units (ADU) or irradiance proportional to $W/m^2$.
 
@@
 
+\section{Real Data Considerations}
+
+\section{Results}
+
+The two objects selected for study in this work are a Delta I upper stage and the ECHOSTAR 2 satellite.
+
+\subsection{Delta I Upper Stage}
+
+The Delta I upper stage (COSPAR ID 1984-115C) is a Star-37E \cite{delta3914_astronautix} solid fuel apogee kick motor, displayed in Figure \ref{fig:star37e}.
+It is $d=0.93$ meters in diameter and is approximately $h=1.7$ meters in length \cite{star37e_astronautix, star37_gunter}.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figsmall]{star37e.JPG}
+  \caption{The Star-37E upper stage \cite{star37_af}.}
+  \label{fig:star37e}
+\end{figure}
+
+\textit{Computing its MOI}
+
+We compute the inertia tensor of the Star-37E by approximating its geometry as a hollow cylinder of the same aspect ratio such that \cite{serway2019}:
+
+\begin{align}
+  J_a &= \frac{1}{2} m \left(r_o^2+r_i^2\right) \\
+  J_t &= \frac{1}{12} m \left(3 \left(r_o^2+r_i^2\right) + h^2\right). \\
+\end{align}
+
+Here, we choose the inner radius to be $r_i=0.45*d$, the outer radius to be $r_o=0.5d$, and $m=1$ kilogram -- dividing each moment of inertia by a constant does not change the dynamics -- yielding $J_a = 15.3 \: [\text{kg} \cdot m^2]$ and $J_t = 27.9 \: [\text{kg} \cdot m^2]$.
+
+[\textit{MODEL PIC 1}]
+
+\subsection{ECHOSTAR II}
+
+ECHOSTAR II (COSPAR ID 1996-055A) was a communications satellite on the AS-7000 bus \cite{as7000_astronautix}, shown in Figure \ref{fig:echostar1}.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figbig]{EchoStar-1_image.jpg}
+  \caption{ECHOSTAR II artist's rendition \cite{as7000_astronautix}.}
+  \label{fig:echostar1}
+\end{figure}
+
+[\textit{MODEL PIC 2}]
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figbig]{sphx_glr_plot_models_001.png}
+  \caption{Simplified ECHOSTAR II box-wing model used for inversion.}
+  \label{fig:echostar1_simple}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figbig]{sphx_glr_plot_models_002.png}
+  \caption{Simplified Star-37E model used for inversion.}
+  \label{fig:star37e_simple}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figbig]{sphx_glr_plot_lcs_001_2_00x.png}
+  \caption{Selected light curve for the Delta (Star-37E) rocket body, observed by the Purdue Optical Ground Station}
+  \label{fig:rb_lc_obs}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figbig]{sphx_glr_plot_lcs_002_2_00x.png}
+  \caption{Selected light curve for ECHOSTAR II, observed by the Purdue Optical Ground Station}
+  \label{fig:sat_lc_obs}
+\end{figure}
+
  \begin{table}[ht]
@@
 
-[\textit{TODO: Acknowledge AIUB if we end up using some of their data for a GLONASS light curve}]
+% [\textit{TODO: Acknowledge AIUB if we end up using some of their data for a GLONASS light curve}]
 
 \bibliography{cache/refs.bib}
-
-% \begin{thebibliography}{}
-
-% \bibitem{author73}
-% Author T., (1973). Astrophysical Quantities, Athlone Press
-
-% \bibitem{nobody97}
-% Nobody B., Somebody G., Who D., et~al., (1997). {\it The book}, Publisher, ed. 2
-
-% \bibitem{smith96}
-% Smith A., Jones B., (1996). The new discovery, {\it Other Journal}, {\bf 223}(1), 1029--1101
-
-% \end{thebibliography}
 \end{document}

### 2025-02-27 paper_cache.py

@@ -15,6 +15,10 @@ def cache_file(start_path: str, end_dir: str = "cache"):
 
 imgs = {
     "/Users/liamrobinson/Documents/maintained-research/mirage/docs/build/html/_images": [
+        'sphx_glr_plot_models_001.png',
+        'sphx_glr_plot_models_002.png',
+        'sphx_glr_plot_lcs_001_2_00x.png',
+        'sphx_glr_plot_lcs_002_2_00x.png',
     ],
     "/Users/liamrobinson/Documents/presentations-papers/msthesis/static_images": [
     ],

### 2025-03-06 paper.tex
added tables for lc obs and stuff

@@
 % \bibliographystyle{alpha}
-\bibliographystyle{apalike}
+\bibliographystyle{plain}
+
+% \usepackage[style=apa, natbib=true]{biblatex} % Use 'apa' style instead of 'apalike'
+\usepackage{url} % Ensure URLs are properly formatted
+% \addbibresource{cache/refs.bib} % Use biblatex's method for loading the bibliography
 
@@
 \newcommand{\sothree}[0]{$SO_3$}
-\newcommand{\pogslla}[0]{$32.900^\circ \textrm{ N}, -105.533^\circ \textrm{ W} \textrm{ }$}
+\newcommand{\pogslla}[0]{$32.900^\circ \textrm{ N}, -105.533^\circ \textrm{ W}$}
 
@@
 
+% \title{Optimal Full State Light Curve Attitude Inversion: Two Real Case Studies}
 \title{Optimal Light Curve Attitude Inversion with Measurement Noise: Two Case Studies}
@@
 
-Understanding the orientation and angular velocity state of active or passive human-made space objects is critical for many operations like long-term orbit propagation, aiding a mission, determining mission status, and active debris removal.
-Beyond low-Earth orbit, optical telescopes are used predominantly to track the space objects.
-Although this data is abundant, it is generally impossible to resolve even large satellites and rocket bodies in optical ground-based imagery due aperture size and atmospheric distortion.
-As a result, shape and orientation information is not available in any single image.
-But a measurement of the object's total brightness can still be obtained.
-Even if the object's shape and reflective properties are known, any given brightness measurement generally corresponds to infinitely many possible orientations.
+Understanding the orientation and angular velocity state of active or passive human-made space objects is critical for many space situational awareness operations like long-term orbit propagation, anomaly resolution, determining mission status, and active debris removal.
+Beyond low-Earth orbit, optical telescopes are predominantly used to track space objects.
+Due to atmospheric distortion and aperture size, it is generally impossible to resolve even large satellites and rocket bodies in optical ground-based imagery.
+As a result, shape and orientation information is unavailable in any single image, but a measurement of the object's total brightness can still be obtained.
+Even if the object's shape and reflective properties are known, any given brightness measurement will generally correspond to infinitely many possible orientations.
 To constrain the solution space, the brightness can be tracked over time -- producing a sequence of measurements called a light curve.
-If the identity of the object is known, its attitude profile can be estimated from the light curve through a process known as light curve attitude inversion.
-Due to the natural and instrumental noise in the light curve, as well as symmetries in the observation geometry, there may be multiple attitude profiles which are equally likely to produce any given light curve.
+If the object's identity is known, its attitude profile can be estimated from the light curve -- up to certain ambiguities -- through a process known as light curve attitude inversion.
+Due to the environmental and instrumental noise in the light curve, as well as symmetries in the observation geometry, there may be multiple attitude profiles that are equally likely to produce any given light curve.
+To identify these ambiguous solutions, global attitude inversion methods must thoroughly search the solution space.
+
+In this work, we apply a BFGS-based global attitude inversion algorithm to real observations taken by the Purdue Optical Ground Station of a Delta I upper stage and the ECHOSTAR II satellite.
+Given a light curve with an estimate of the shape and reflective properties of the object, our procedure searches the space of initial orientations, angular velocities, and inertia tensors to find initial conditions that produce light curves with low error compared to the observed values.
+Our formulation inherently accounts for the multiple types of ambiguities resulting from the object and observation geometry, the physical constraints of torque-free rigid body motion, and noise in the light curve.
 
@@
 
-[\textit{TODO: lit}]
+Methods for estimating the attitude state of uncontrolled space objects from light curves differ significantly depending on the presence of an initial guess.
+If a guess is available, many filter-based approaches have been developed to process new photometry to update the guess orientation and angular velocity.
+This class of methods includes single Kalman filters \cite{burton2021two, gagnon2024, wetterer2009}, and multi-filter multiple-model adaptive estimation (MMAE) schemes \cite{linares2014space, dianetti2020}.
+These approaches can be very effective if the initial state estimate is close to the truth, but have been reported to diverge for orientation errors as small as $5^\circ$ in orientation and $0.03^\circ/s^{-1}$ in angular velocity \cite{gagnon2024} due measurement ambiguities and nonlinearity.
 
-In this work, we apply a light curve attitude inversion algorithm to real observations taken by the Purdue Optical Ground Station.
-Given an estimate of the shape and reflective properties of the object, our procedure searches the space of initial orientations, angular velocities, and inertia tensors to find initial conditions that produce light curves with low error compared to the observed values.
-Specifically, we estimate a full distribution for each candidate light curve and compute the likelihood of the observations being a sample from that distribution.
-To solve the optimization problem, we create a number of sampled candidate solutions, scattered throughout the solution space, and perform nonlinear optimization on each using the BFGS algorithm.
-This way, our optimization procedure reliably converges to all maximum likelihood estimates of the attitude state.
-Low-error minima are then clustered to identify families of nearby solutions with similar likelihood.
+If no initial attitude state is available, as in most cases of passive debris observation, the inversion problem must be solved globally.
+Filtering approaches can be applied here too; particle \cite{linares2014particle, holzinger2014} and Bayesian multi-hypothesis filters \cite{burton2021two, cabrera2023} have shown promise in prior work.
+If a reliable prior distribution for the attitude state is not available -- which it generally is not -- a Bayesian method may be inadvisable.
+In the absence of any prior knowledge about the object's attitude state, simulated annealing \cite{gagnon2024, clark2020}, genetic algorithms \cite{gagnon2024, piergentili2017, clark2020}, and particle swarm optimizers (PSOs) \cite{clark2020, clark2022, burton2024journal, burton2024scitech, gagnon2024} can be applied to search the solution space exhaustively.
+
+Much of the past literature on attitude inversion from real light curve data is centered on determining the constant spin rate and axis of rotation of the observed object.
+Frequency analysis in the form of periodograms, Fourier transforms, or epoch folding, is often used to determine the spin rate based on the light curve's frequency spectrum \cite{silha2015, silha2021, isoletta2024, schildknecht2015, pittet2017, yanagisawa2012, koshkin2018}.
+If optical material properties are known, the width of a single specular glint can provide sufficient information to determine the rotation rate \cite{hinks2016}.
+There has also been interest in tracking the evolution of spin rates over time for different classes of objects in different orbital regimes \cite{koshkin2018, blacketer2022, karpov2016}.
+These spin rate determination methods are designed for single-axis rotations and can fail for tumbling objects.
+
+The rotation axis is often estimated for diffusely reflecting objects that are well-approximated as ellipsoids via the so-called amplitude method \cite{williams1979location} which has been extended for combined spin and precession motion \cite{yanagisawa2012}.
+For light curves with strong specular peaks, the timing of these glints can be used to determine the constant rotation axis given multiple passes of observations \cite{vananti2023, koshkin2018}.
+These approaches are not applicable for generalized rigid body motion where no simplifying assumptions about the location of the spin axis are available.
+
+Works that seek to estimate a full rigid body attitude state from real data -- a time-varying orientation and angular velocity -- grid search \cite{shafer2017} and genetic algorithms \cite{piergentili2020, gallucci2020} have been used.
+These full-state estimation studies are limited by computational complexity, poor knowledge of material reflectivities, and the presence of measurement noise.
+
+Our method is a global full-state attitude solver inspired by recent PSO work \cite{burton2024journal} but is easier to tune, can be fully parallelized, and is robust to significant measurement noise \cite{robinson2025att}.
+Further, our method differs from other global estimation methods in that we do not seek to find one optimal solution and instead search for a collection of high-quality local minima to directly study the many ambiguous solutions for a given light curve.
+To solve the optimization problem, we sample many candidate solutions scattered throughout the solution space and perform gradient-based optimization on each using the BFGS algorithm.
+Our optimization procedure reliably converges to high-quality maximum likelihood estimates of the attitude state.
 The final output of our method is an array of the best solutions, ranked by their likelihood.
-Formulating the problem this way inherently accounts for the multiple classes of ambiguities resulting from the object and observation geometry, the physical constraints of torque-free rigid body motion, as well as the noise on the light curve.
+Clusters of low-error minima can be analyzed to identify families of solutions.
 
-To test the efficacy of this method, we perform attitude inversion on real observations gathered by the Purdue Optical Ground Station of an inactive GLONASS satellite and a rocket body.
+This work proceeds by describing the inversion method, discussing efficient methods for computing the time-varying probability density of the light curve, and presenting inversion results for the two test cases.
+
+% .
 Solutions for the GLONASS case are analyzed to highlight the effect of observation geometry ambiguities, while the rocket body case highlights ambiguities introduced by the geometry of the object.
 Our noise model is validated against further real observations, confirming that all relevant noise sources have been accounted for and modeled accurately.
-We show that any significant lack of knowledge in the object shape or optical material properties causes the solution to rapidly degrade.
+We show that any significant lack of knowledge of the object's shape or optical material properties causes the solution to rapidly degrade.
 
@@
 
-The state vector $\vctr{x}$ we optimize to fit the observed light curve is the concatenation of the Modified Rodriguez Parameter (MRP) $\vctr{p} = [p_1, p_2, p_3]^T$, the principal body frame angular velocity $\vctr{\omega} = [\omega_1, \omega_2, \omega_3]$ in radians per second, and the ratios of the principal moments of inertia $J_y / J_x$ and $J_z / J_x$ where the full inertia tensor is $J = \mathrm{diag}\left([J_x, J_y, J_z]\right)$:
+The state vector $\vctr{x}$ we optimize to fit the observed light curve is the concatenation of the Modified Rodriguez Parameter (MRP) $\vctr{p} = [p_1, p_2, p_3]^T$, the principal body frame angular velocity $\vctr{\omega} = [\omega_1, \omega_2, \omega_3]$ in radians per second, and the ratios of the principal moments of inertia $J_y / J_x$ and $J_z / J_x$ where the inertia tensor in principal axes is $J = \mathrm{diag}\left([J_x, J_y, J_z]\right)$:
 
@@
 
-The first six elements of the state vector are time varying with dynamics given by:
+The first six elements of the state vector are time-varying with dynamics given by:
 
@@
 
-We choose to represent the surface geometry of space objects as a polyhedral mesh composed of a set of flat facets.
+We choose to represent the surface geometry of space objects as a polyhedral mesh composed of a collection of flat facets.
 Each $i$th face has $n_v \geq 3$ or more vertices $\left\{ v_{i,1}, v_{i,2}, v_{i,n_v} \right\}$ and an outward-pointing normal vector $\unitv{n}_i$.
@@
 
-The unshadowed area $\bar{a}$ may be computed in a number of ways.
+The unshadowed area $\bar{a}$ may be computed in several ways.
 For convex objects, $\bar{a}=a$ as self-shadowing is impossible.
-For highly-nonconvex and detailed models composed of thousands of faces, $\bar{a}$ can be approximated on a pixel grid using shadow mapping \cite{robinson2022}.
+For highly nonconvex and detailed models composed of thousands of faces, $\bar{a}$ can be approximated on a pixel grid using shadow mapping \cite{robinson2022}.
 Self-shadowing can be solved semi-analytically for less detailed approximations of nonconvex objects by computing the mutual intersections of the polygons $P_k$ on other faces whose projections overlap with the face in question.
@@
 
-As the objects we investigate in this work are well-approximated with relatively simple nonconvex meshes, we use the semi-analytic method to efficiently compute unshadowed areas via the Sutherland-Hodgeman algorithm \cite{sutherland1974} for each polygon intersection.
+As the objects we investigate in this work are well-approximated with relatively simple non-convex meshes, we use the semi-analytic method to efficiently compute unshadowed areas via the Sutherland-Hodgeman algorithm \cite{sutherland1974} for each polygon intersection.
 
@@
 The bidirectional reflectance distribution function (BRDF) defines the amount of incident radiation from $\unitv{s}$ reflected per steradian in the observer's direction $\unitv{o}$.
-We choose the Blinn-Phong \cite{blinn1977} BRDF for this work as it is efficient to compute and satisfies the three main requirements for a physically-meaningful BRDF as it is nonnegative, energy-conserving, and reciprocal \cite{duvenhage2013}.
-The Blinn-Phong BRDF is parameterized by the coefficient of diffuse reflectivity $C_d$, the coefficient of specular reflectivity $C_s$, and the specular exponent $n>0$.
-The coefficients of reflectivity implicity satisfy $C_a + C_s + C_d = 1$ for energy conservation, where $C_a$ is the coefficient of absorption \cite{fan2020thesis}.
+We choose the Blinn-Phong \cite{blinn1977} BRDF for this work as it is efficient to compute and satisfies the three main requirements for a physically meaningful BRDF as it is nonnegative, energy-conserving, and reciprocal \cite{duvenhage2013}.
+We avoid more recent microfacet models for computational efficiency.
+In practice, the uncertainty in the reflective properties of the observed objects will often be larger than errors introduced by the choice of reflection model.
+The Blinn-Phong BRDF is parameterized by the coefficient of diffuse reflectivity $C_d$, the coefficient of specular reflectivity $C_s$, and the specular exponent $n>0$:
 
@@
 \end{equation}
+
+The coefficients of reflectivity implicity satisfy $C_a + C_s + C_d = 1$ for energy conservation, where $ 0 \leq C_a \leq 1$ is the coefficient of absorption \cite{fan2020thesis}.
 
@@
 
-The output of Equation \ref{eq:lc_norm} is a light curve, but as many of the noise characteristics of the signal are defined in image sensor's native unit of ADU, we convert the light curve into the same units via:
+The output of Equation \ref{eq:lc_norm} is a light curve, but as many of the noise characteristics of the signal are defined in the image sensor's native unit of ADU, we convert the light curve into the same units via:
 
@@
 Within the integral, we account for the telescope's passband filter $P(\lambda)$, the image sensor quantum efficiency $Q(\lambda)$, the atmospheric absorption spectrum $T_k(\lambda)$, the solar irradiance spectrum $I_\odot(\lambda)$, and the inverse energy of a photon with wavelength $\lambda$, $\lambda / hc$, where $h$ is Planck's constant in Joule-seconds, and $c$ is the speed of light in vacuum in meters per second.
-Taken together, the integral computes the fraction of the energy reflected from the object absorbed into the image sensor, while the the outer factor dimensionalizes the result to yield the mean total sensor response in ADU across all pixels due to the object's signal.
+Taken together, the integral computes the fraction of the energy reflected from the object absorbed into the image sensor, while the outer factor dimensionalizes the result to yield the mean total sensor response in ADU across all pixels due to the object's signal.
 
-the variance a space object's light curve whose mean is determined by Equation \ref{eq:general_bright} is a combination of a number of relevant independent stochastic processes.
+The variance of a space object's light curve whose mean is determined by Equation \ref{eq:general_bright} is a combination of many independent stochastic processes.
 These distributions have variances $\sigma^2_\text{sensor}$ due to the sensor's integration and readout effects, $\sigma^2_\text{flat}$ from sensor flat-fielding effects, scintillation noise $\bar{S}_{\text{SO},k} \sigma^2_{Y,k}$ \cite{osborn2015}, Poisson signal shot noise $\lambda_{\text{shot},k}$, and Poisson background noise $\lambda_{\text{back},k}$.
-The sum of these variances yields total signal variance in ADU:
+The sum of these variances yields the total signal variance in ADU:
 
@@
 The background standard deviation is modeled by the sum of six independent Poisson random processes contributing to light entering the telescope optics from environmental sources.
-These processes and the sources of their respective models are: scattered moonlight $\lambda_{\text{moon},k}$ \cite{daniels1977}, integrated starlight $\lambda_{\text{star},k}$ \cite{krag2003}, twilight $\lambda_{\text{twi},k}$ \cite{pata2006}, zodiacal light $\lambda_{\text{zod},k}$ \cite{roach1972}, airglow $\lambda_{\text{air},k}$ \cite{krag2003}, and light pollution $\lambda_{\text{poll},k}$ \cite{falchi2016, falchi2016_data}.
+These processes and the sources of their respective models are: scattered moonlight $\lambda_{\text{moon},k}$ \cite{daniels1977}, integrated starlight $\lambda_{\text{star},k}$ \cite{krag2003}, twilight $\lambda_{\text{twi},k}$ \cite{patat2006}, zodiacal light $\lambda_{\text{zod},k}$ \cite{roach1972}, airglow $\lambda_{\text{air},k}$ \cite{krag2003}, and light pollution $\lambda_{\text{poll},k}$ \cite{falchi2016, falchi2016_data}.
 In summation, the total Poisson background variance is:
@@
 
-[\textit{TODO: describe the variables here, if we want to keep this eq... it's in the cited paper so i am not sure}]
+[\textit{TODO: describe the variables here, if we want to keep this eq... it's in the cited paper so I am not sure}]
 
@@
 
-After summation in Equation \ref{eq:sigma_total}, we can approximate the observed light curve as Gaussian distributed via the mean $\bar{S}_{\text{SO},k}$ from Equation \ref{eq:general_bright} and the variance $\sigma^2_{S,k}$ in $[\text{ADU}^2]$:
+After summation in Equation \ref{eq:sigma_total}, we can approximate the observed light curve as Gaussian distributed via the mean $\bar{S}_{\text{SO},k}$ from Equation \ref{eq:general_bright} and the variance $\sigma^2_{S,k}$ in $\text{ADU}^2$:
 
@@
   \includegraphics[width=\figsmall]{star37e.JPG}
-  \caption{The Star-37E upper stage \cite{star37_af}.}
+  \caption{Star-37E upper stage \cite{star37_af}.}
   \label{fig:star37e}
-\end{figure}
-
-\textit{Computing its MOI}
-
-We compute the inertia tensor of the Star-37E by approximating its geometry as a hollow cylinder of the same aspect ratio such that \cite{serway2019}:
-
-\begin{align}
-  J_a &= \frac{1}{2} m \left(r_o^2+r_i^2\right) \\
-  J_t &= \frac{1}{12} m \left(3 \left(r_o^2+r_i^2\right) + h^2\right). \\
-\end{align}
-
-Here, we choose the inner radius to be $r_i=0.45*d$, the outer radius to be $r_o=0.5d$, and $m=1$ kilogram -- dividing each moment of inertia by a constant does not change the dynamics -- yielding $J_a = 15.3 \: [\text{kg} \cdot m^2]$ and $J_t = 27.9 \: [\text{kg} \cdot m^2]$.
-
-[\textit{MODEL PIC 1}]
-
-\subsection{ECHOSTAR II}
-
-ECHOSTAR II (COSPAR ID 1996-055A) was a communications satellite on the AS-7000 bus \cite{as7000_astronautix}, shown in Figure \ref{fig:echostar1}.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\figbig]{EchoStar-1_image.jpg}
-  \caption{ECHOSTAR II artist's rendition \cite{as7000_astronautix}.}
-  \label{fig:echostar1}
-\end{figure}
-
-[\textit{MODEL PIC 2}]
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\figbig]{sphx_glr_plot_models_001.png}
-  \caption{Simplified ECHOSTAR II box-wing model used for inversion.}
-  \label{fig:echostar1_simple}
 \end{figure}
@@
 
+We compute the inertia tensor of the Star-37E by approximating its geometry as a hollow cylinder of the same aspect ratio such that \cite{serway2019}:
+
+\begin{align}
+ J_a &= \frac{1}{2} m \left(r_o^2+r_i^2\right) \\
+ J_t &= \frac{1}{12} m \left(3 \left(r_o^2+r_i^2\right) + h^2\right) \\
+ J_\text{R/B} &= \text{diag} \left(J_a, J_t, J_t\right).
+\end{align}
+
+Here, we choose the inner radius to be $r_i=0.45d$, the outer radius to be $r_o=0.5d$, and $m=1$ kilogram -- dividing each moment of inertia by a constant does not change the dynamics -- yielding $J_a = 15.3 \: [kg \cdot m^2]$ and $J_t = 27.9 \: [kg \cdot m^2]$.
+Table \ref{tb:case1_in} summarizes the light curve observations that result in the extracted measurements displayed in Figure \ref{fig:rb_lc_obs}.
+
+\begin{table}[ht]
+  \centering
+  \caption{Delta I rocket body observation parameters}
+  \vspace*{6pt}
+  \begin{tabular}{|l|l|}
+  \hline
+  \textbf{Variable} & \textbf{Value} \\ \hline
+  Telescope location & \pogslla \\ \hline
+  Target COSPAR ID & 1984-115C \\ \hline
+  First obs.\ (UTC) & Mar 2, 2025 01:53:30.251 \\ \hline
+  Light curve duration & $7.5$ minutes \\ \hline
+  Observations & $100$ ($92$ extracted) \\ \hline
+  Obs.\ frequency & $0.222$ Hz \\ \hline
+  Integration time & $3$ seconds \\ \hline
+  \end{tabular}
+  \label{tb:case1_in}
+\end{table}
+
 \begin{figure}[ht]
@@
 \end{figure}
+
+\begin{table}[ht]
+  \centering
+  \caption{Delta I rocket body inversion assumptions}
+  \vspace*{6pt}
+  \begin{tabular}{|l|l|}
+  \hline
+  \textbf{Variable} & \textbf{Value} \\ \hline
+  Uniform BRDF & $C_d=0.2$, $C_s=0.8$, $n=30$ \\ \hline
+  $|\omega(0)|_\text{max}$ & $8.31$ $\text{deg} / \text{s}^{-1}$ \\ \hline
+  \end{tabular}
+  \label{tb:case1_ass}
+\end{table}
+
+\subsection{ECHOSTAR II}
+
+ECHOSTAR II (COSPAR ID 1996-055A) was a communications satellite on the AS-7000 bus \cite{as7000_astronautix}, shown in Figure \ref{fig:echostar1}.
+It has an approximate total wingspan of $s = 23.9$ meters, individual solar panel length of $l_p=8.5$ meters, panel width of $w_p=3.1$ meters, and a mass per panel of $m_p = 60$ kilograms \cite{earl2015}.
+The spacecraft bus is approximately cubic with a side length $w_b=2.3$ meters and a mass of $m_b = 1900$ kilograms \cite{earl2015}.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figbig]{EchoStar-1_image.jpg}
+  \caption{ECHOSTAR II artist's rendition \cite{as7000_astronautix}.}
+  \label{fig:echostar1}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figbig]{sphx_glr_plot_models_001.png}
+  \caption{Simplified ECHOSTAR II box-wing model used for inversion.}
+  \label{fig:echostar1_simple}
+\end{figure}
+
+The contribution of the bus to the total inertia tensor of ECHOSTAR II is spherically symmetric and given by: [\textit{TODO: citation needed}]:
+
+\begin{align}
+ J_\text{bus} &= \frac{1}{6} m_b w_b^2 I_3,
+\end{align}
+
+\noindent
+assuming that the center of mass (COM) of the bus coincides with the COM of the entire vehicle.
+By the intermediate axis theorem, the contribution of a panel to the inertia tensor is a function of the displacement of the panel's COM from the vehicle's overall COM $\vctr{r}_p = [ 0, s/2 - l_p/2, 0]^T$ [\textit{TODO: citation needed}]:
+
+\begin{equation}
+  \begin{split}
+ J_\text{pan} = &\frac{m_p}{12} \cdot \text{diag}\left(l_p^2, w_p^2, \left(l_p^2 + w_p^2\right) \right) + \\&m_p \left[ \left( \vctr{r}_p \cdot \vctr{r}_p \right) I_3 - \vctr{r}_p \vctr{r}_p^T \right],
+  \end{split}
+\end{equation}
+
+\noindent
+assuming that the panels have negligible thickness. The overall inertia tensor for ECHOSTAR II is:
+
+\begin{align}
+ J_\text{SAT} &= J_\text{bus} + 2J_\text{pan} \\
+  &= \text{diag} \left( 9512, 1771, 9609 \right) \: [kg \cdot m^2]
+\end{align}
+
+\begin{table}[ht]
+  \centering
+  \caption{ECHOSTAR II observation parameters}
+  \vspace*{6pt}
+  \begin{tabular}{|l|l|}
+  \hline
+  \textbf{Variable} & \textbf{Value} \\ \hline
+  Telescope location & \pogslla \\ \hline
+  Target COSPAR ID & 1996-055A \\ \hline
+  First obs.\ (UTC) & Feb 26, 2025 04:16:12.279 \\ \hline
+  Light curve duration & $19.5$ minutes \\ \hline
+  Observations & $456$ ($395$ extracted) \\ \hline
+  Obs.\ frequency & $0.388$ Hz \\ \hline
+  Integration time & $1$ seconds \\ \hline
+  \end{tabular}
+  \label{tb:case2_in}
+\end{table}
 
@@
 
+We estimate a maximum bound for the angular velocity magnitude by inspecting the light curve [\textit{TODO: finish sentence}]
+
 \begin{table}[ht]
   \centering
-  \caption{Observation parameters 2}
+  \caption{ECHOSTAR II inversion assumptions}
   \vspace*{6pt}
@@
   \textbf{Variable} & \textbf{Value} \\ \hline
-  Observation start (UTC) & Oct 10, 2024 08:00:00 \\ \hline
-  COSPAR ID & 1987-084D \\ \hline 
- Light curve duration & $2$ minutes \\ \hline
- Measurement frequency & $0.5$ Hz \\ \hline
- Integration time & $0.01$ seconds \\ \hline
+  Aluminum BRDF & $C_d=0.22$, $C_s=0.4$, $n=5$ \\ \hline
+  MLI BRDF & $C_d=0.05$, $C_s=0.24$, $n=20$ \\ \hline
+  Solar panel BRDF & $C_d=0.05$, $C_s=0.13$, $n=10$ \\ \hline
+  $|\omega(0)|_\text{max}$ & $2.32$ $\text{deg} / \text{s}^{-1}$ \\ \hline
   \end{tabular}
-  \label{tb:case1_in}
-\end{table}
-
-\begin{table}[ht]
-  \centering
-  \caption{Observation parameters 2}
-  \vspace*{6pt}
-  \begin{tabular}{|l|l|}
-  \hline
-  \textbf{Variable} & \textbf{Value} \\ \hline
-  Observation start (UTC) & Oct 10, 2024 08:00:00 \\ \hline
-  COSPAR ID & 1987-084D \\ \hline 
- Light curve duration & $2$ minutes \\ \hline
- Measurement frequency & $0.5$ Hz \\ \hline
- Integration time & $0.01$ seconds \\ \hline
-  \end{tabular}
-  \label{tb:case1_in}
+  \label{tb:case2_ass}
 \end{table}
@@
 
-% [\textit{TODO: Acknowledge AIUB if we end up using some of their data for a GLONASS light curve}]
-
+\clearpage
 \bibliography{cache/refs.bib}

### 2025-03-06 copy_to_dropbox.py

@@ -5,4 +5,4 @@ import datetime
 dtime = str(datetime.datetime.now().date())
 dest_folder = "/Users/liamrobinson/Library/CloudStorage/Dropbox/progress_PhD/papers/for-review/"
 dest_path = os.path.join(dest_folder, f"sdc-attitude-{dtime}.pdf")
-shutil.copy("final/paper-final.pdf", dest_path)
+shutil.copy("final/paper.pdf", dest_path)

### 2025-03-09 paper.tex
added plots, i guess

@@
 \usepackage{graphicx}
-
+\usepackage{placeins}
 \usepackage{bm} % for bold math
@@
 \newcommand{\sothree}[0]{$SO_3$}
-\newcommand{\pogslla}[0]{$32.900^\circ \textrm{ N}, -105.533^\circ \textrm{ W}$}
 
@@
 
-% \title{Optimal Full State Light Curve Attitude Inversion: Two Real Case Studies}
 \title{Optimal Light Curve Attitude Inversion with Measurement Noise: Two Case Studies}
@@
 
+\FloatBarrier
 \section{Inversion Method}
@@
 
+\FloatBarrier
 \section{Computing the Light Curve}
@@
 
-\subsection{The Light Curve}
+\subsection{Computing the Mean Observed Signal}
 
@@
 
-The variance of a space object's light curve whose mean is determined by Equation \ref{eq:general_bright} is a combination of many independent stochastic processes.
+\subsection{Approximating the Measurement Noise}
+
+The variance of a space object's observed brightness whose mean is determined by Equation \ref{eq:general_bright} is a combination of many independent stochastic processes.
 These distributions have variances $\sigma^2_\text{sensor}$ due to the sensor's integration and readout effects, $\sigma^2_\text{flat}$ from sensor flat-fielding effects, scintillation noise $\bar{S}_{\text{SO},k} \sigma^2_{Y,k}$ \cite{osborn2015}, Poisson signal shot noise $\lambda_{\text{shot},k}$, and Poisson background noise $\lambda_{\text{back},k}$.
@@
 
-\section{Real Data Considerations}
+\FloatBarrier
+\section{Attitude Inversion Results for Synthetic Data}
 
-\section{Results}
+To demonstrate the effectiveness of this method, we first present results for a simulated light curve where the ground truth is known.
+
+\begin{table}[ht]
+  \centering
+  \caption{Simulated observation parameters}
+  \vspace*{6pt}
+  \begin{tabular}{|l|l|}
+  \hline
+  \textbf{Variable} & \textbf{Value} \\ \hline
+  Target COSPAR ID & 1990-054D \\ \hline
+  First obs.\ (UTC) & Mar 8, 2025 02:00:00.000 \\ \hline
+  Light curve duration & $5$ minutes \\ \hline
+  Observations & $100$ \\ \hline
+  Integration time & $0.5$ seconds \\ \hline
+  \end{tabular}
+  \label{tb:case1_in}
+\end{table}
+
+\begin{table}[ht]
+  \centering
+  \caption{Reflection properties of materials used in the synthetic data test case. The authors estimated the white paint parameters qualitatively}
+  \vspace*{6pt}
+  \begin{tabular}{|l|l|l|l|}
+  \hline
+  \textbf{Material} & $C_d$ & $C_s$ & $n$ \\ \hline
+  Aluminum \cite{fankhauser2023} & $0.34$ & $0.40$ & $8.9$ \\ \hline
+  White paint & $0.9$ & $0.1$ & $1$ \\ \hline
+  \end{tabular}
+  \label{tb:real_matprops}
+ \end{table}
+
+\begin{table}[ht]
+  \centering
+  \caption{True object attitude profile for all test cases}
+  \vspace*{6pt}
+  \begin{tabular}{|l|l|}
+  \hline
+  \textbf{Variable} & \textbf{Value} \\ \hline
+ Initial orientation & $\vctr{p}(0) = -\frac{1}{3} \begin{bmatrix} 1 & 1 & 1 \end{bmatrix}^T$ \\ \hline
+ Initial $\vctr{\omega}$ ($\text{rad}/\text{s}$) & $\vctr{\omega}(0) = \begin{bmatrix} 0.03 & 0.06 & 0.03 \end{bmatrix}^T$ \\ \hline
+ Inertia tensor ratios & $J_y / J_x = 1, \: J_z / J_x = 0.25$ \\ \hline
+ External torque & $\vctr{M}=\begin{bmatrix} 0 & 0 & 0 \end{bmatrix}^T$ \\ \hline
+  \end{tabular}
+  \label{tb:synth_truth}
+\end{table}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figsmall]{rocket_bodies2.png}
+  \caption{Measured rocket body model \cite{nasa_models}.}
+  \label{fig:meas_model}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figsmall]{rocket_bodies1.png}
+  \caption{Simplified rocket body model used for inversion.}
+  \label{fig:inv_model}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figsmall]{light_curves_adu_rocket_body.png}
+  \caption{Simulated observed flux for the rocket body.}
+  \label{fig:obs_adu_synth}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figsmall]{light_curves_mag_rocket_body.png}
+  \caption{Simulated observed V-band magnitude for the rocket body.}
+  \label{fig:obs_mag_synth}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figsmall]{light_curves_std_rocket_body.png}
+  \caption{Simulated observed light curve standard deviation computed with Equation \ref{eq:sigma_total}.}
+  \label{fig:obs_std_synth}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figsmall]{sigma_solsRocket Body.png}
+  \caption{Converged solution initial orientation MRP vectors.}
+  \label{fig:omega_sols1}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figsmall]{omega_sols_Rocket Body.png}
+  \caption{Converged solution initial body-frame angular velocities.}
+  \label{fig:omega_sols1}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\figsmall]{spin_proc_Rocket Body.png}
+  \caption{Converged solution body-frame spin and precession rates.}
+  \label{fig:omega_sols1}
+\end{figure}
+
+\FloatBarrier
+\section{Attitude Inversion Results for Real Data}
 
@@
   \textbf{Variable} & \textbf{Value} \\ \hline
-  Telescope location & \pogslla \\ \hline
   Target COSPAR ID & 1984-115C \\ \hline
@@
   \textbf{Variable} & \textbf{Value} \\ \hline
-  Telescope location & \pogslla \\ \hline
   Target COSPAR ID & 1996-055A \\ \hline
@@
   \hline
-  \textbf{Variable} & \textbf{Value} \\ \hline
+  \textbf{Material} & \textbf{Coefficients} \\ \hline
   Aluminum BRDF & $C_d=0.22$, $C_s=0.4$, $n=5$ \\ \hline
@@
 
+\FloatBarrier
 \section*{Acknowledgments}
@@
 
+\FloatBarrier
+\section*{Appendix}
+
+\begin{table}[ht]
+  \centering
+  \caption{Telescope information common to all test cases}
+  \vspace*{6pt}
+  \begin{tabular}{|l|l|}
+  \hline
+  \textbf{Variable} & \textbf{Value} \\ \hline
+  CCD sensor & KAF-16803 \\ \hline
+  Effective aperture area $A_\circ$ & $0.076$ $m^2$ \\ \hline
+  Observer location & $32.900^\circ \textrm{ N}, -105.533^\circ \textrm{ W}$ \\ \hline
+  Observer altitude & $2.24$ km above MSL \\ \hline
+  Source pixels $n_\text{pix}$ & $100$ \\ \hline
+  Read noise $\sigma_\text{read}$ & $9$ $\text{ADU} / \text{pixel}$ \\ \hline
+  Dark current rate $\lambda_\text{dark}$ & $0.01$ $\text{ADU} / \text{pixel} / \text{s}$ \\ \hline
+  Flat field strength $f_k$ & $0.01$ nondim \\ \hline
+  Gain $g$ & $5.1$ $e^- / \text{ADU}$ \\ \hline
+  \end{tabular}
+  \label{tb:tele_info}
+ \end{table}
+
+ \begin{table}[ht]
+  \centering
+  \caption{BFGS solver configuration for all results cases}
+  \vspace*{6pt}
+  \begin{tabular}{|l|l|}
+  \hline
+  \textbf{Variable} & \textbf{Value} \\ \hline
+  Maximum function evaluations & $1000$ \\ \hline
+  Maximum iterations & $100$ \\ \hline
+  Finite difference step size & $1 \cdot 10^{-5}$ \\ \hline
+  Gradient $\infty$-norm tolerance & $1 \cdot 10^{-5}$ \\ \hline
+  \end{tabular}
+  \label{tb:bfgs_info}
+ \end{table}
+
+\FloatBarrier
 \clearpage

### 2025-03-09 figure_normalize.py

@@ -42,7 +42,7 @@ for i, match in enumerate(matches, start=1):
     # img.convert('RGB').save(f'final/fig{i}.jpg', format='JPEG', dpi=(600,600))
     # replacement = rf"{match[0]}fig{i}.jpg{match[2]}"
 
-    os.system(f'cp cache/{match[1]} final/fig{i}.png')
+    os.system(f'cp "cache/{match[1]}" final/fig{i}.png')
     replacement = rf"{match[0]}fig{i}.png{match[2]}"
     content = content.replace("".join(match), replacement, 1)
 

### 2025-03-09 paper_cache.py

@@ -27,6 +27,8 @@ imgs = {
     ],
     '/Users/liamrobinson/Documents/maintained-research/mirage/examples/13-attitude-2024/figs':
     glob.glob('/Users/liamrobinson/Documents/maintained-research/mirage/examples/13-attitude-2024/figs/*.png'),
+    '/Users/liamrobinson/Documents/maintained-research/mirage/examples/15-attitude-sdc-2025/figs':
+    glob.glob('/Users/liamrobinson/Documents/maintained-research/mirage/examples/15-attitude-sdc-2025/figs/*'),
 }
 
 for im_dir, im_names in imgs.items():

### 2025-03-10 paper.tex
almost done kinda

@@
 
-% \bibliographystyle{alpha}
 \bibliographystyle{plain}
 
-% \usepackage[style=apa, natbib=true]{biblatex} % Use 'apa' style instead of 'apalike'
+\usepackage{float}
+
 \usepackage{url} % Ensure URLs are properly formatted
-% \addbibresource{cache/refs.bib} % Use biblatex's method for loading the bibliography
+\usepackage{xcolor}
+
+\pagecolor[rgb]{0.0,0.0,0.0}
+\color[rgb]{1,1,1}
+
+\usepackage[section]{placeins} % ensures that floats are placed in their section
+\makeatletter
+\AtBeginDocument{%
+  \expandafter\renewcommand\expandafter\subsection\expandafter{%
+    \expandafter\FloatBarrier\subsection
+  }%
+}
+\makeatother
+
+% \makeatletter
+% \AtBeginDocument{%
+%   \expandafter\renewcommand\expandafter\subsubsection\expandafter{%
+%     \expandafter\FloatBarrier\subsubsection
+%   }%
+% }
+% \makeatother
 
@@
 \usepackage{graphicx}
-\usepackage{placeins}
 \usepackage{bm} % for bold math
 \usepackage{mathtools} % for \prescript
+\usepackage{amssymb}
+\usepackage{amsfonts}
+\usepackage{adjustbox}
+
 \newcommand{\vctr}[1]{\bm{#1}}
@@
 \DeclareMathOperator{\fl}{floor}
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 
-\FloatBarrier
 \section{Inversion Method}
@@
 \begin{equation} \label{eq:nll_loss}
- f(\hat{\vctr{S}}, \vctr{S}) = \frac{1}{m}\sum_{k=1}^{m}\left[\ln\sigma_k + \frac{1}{2}\left(\frac{S_k - \hat{S}_k}{\sigma_k}\right)^2 \right].
+ f(\vctr{x}) = \frac{1}{m}\sum_{k=1}^{m}\left[\ln\sigma_k + \frac{1}{2}\left(\frac{S_k - \hat{S}_k(\vctr{x})}{\sigma_k}\right)^2 \right].
  \end{equation}
 
+If the reflectivity of the object's faces is not well-known, we rescale the estimated light curve mean $\hat{S}(\vctr{x})$ to have equal norm to the observations to resolve the inherent albedo-area ambiguity:
+
+\begin{equation}
+  \hat{S}(\vctr{x}) = \frac{\norm{S}}{\norm{\tilde{S}(\vctr{x})}} \tilde{S}(\vctr{x}),
+\end{equation}
+
+\noindent
+where $\tilde{S}(\vctr{x})$ is the original, poorly scaled, estimated light curve.
 In this work, the light curve and its standard deviation are always assumed to be expressed in a linear unit, e.g., analog-to-digital units (ADU) or irradiance proportional to $W/m^2$.
 
-\subsection{State Vector Definition and Dynamics}
+\subsection{State Definition and Dynamics}
 
@@
 
-\FloatBarrier
+\subsection{Running the Solver} \ref{sec:run_solver}
+
+Our method first initializes $n$ guesses scattered throughout the solution space -- possibly sampled from a prior distribution based on available knowledge.
+In the most basic case, the initial state guess is sampled for guess $i$ at iteration $k=0$:
+
+\begin{equation}
+  \vctr{x}_{0,i}(0) = \begin{bmatrix}\vctr{p}_{i,0}(0) & \vctr{\omega}_{i,0}(0) & \left(J_y / J_x\right)_{i,0} & \left(J_z / J_x\right)_{i,0}\end{bmatrix}^T.
+\end{equation}
+
+\noindent
+In this work, $\vctr{p}_{i,0}(0)3$ is uniformly randomly sampled from orientation space, $\vctr{\omega}_{i,0}(0)$ is uniformly randomly sampled from the ball in $\mathbb{R}^3$ of radius $|\omega(0)|_\text{max}$ -- the maximum angular rate to search -- while $\left(J_y / J_x\right)_{i,0}$ and $\left(J_z / J_x\right)_{i,0}$ are sampled from Gaussian prior distributions centered at estimated values derived from the known geometry of the object.
+
+After computing each initial state guess, we run the Broyden-Fletcher-Goldfarb-Shanno (BFGS) algorithm \cite{broyden1970, fletcher1970, goldfarb1970, shanno1970} to optimize each guess to a nearby local minimum $\vctr{x}_{k,i}(0)$ in the objective function such that:
+
+\begin{equation} \label{eq:opt_problem}
+  \vctr{x}_{k,i}(0) = \argmin_x f(\vctr{x}),
+\end{equation}
+
+\noindent
+producing a final initial state which is a good local guess for a state which to produce the observations.
+In this way, we globally search the solution space up to the density implied by $n$.
+Using a simpler algorithm like gradient descent can still reach convergence, but will take significantly more objective function evaluations due to the high nonlinearity of the attitude inversion problem.
+
+In order to implement this optimization effectively without constraints, it is often necessary to take the absolute value of the intermediate inertia tensor values to maintain physical validity.
+
 \section{Computing the Light Curve}
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
   \centering
@@
 \end{figure}
+
+In this work, the Sun position is computed using a SPICE ephemeris kernel \cite{spice} while space object positions are propagated with SGP4 TLEs obtained from Space-Track \cite{spacetrack}.
 
@@
 
-[\textit{TODO: describe the variables here, if we want to keep this eq... it's in the cited paper so I am not sure}]
+Here, the scintillation noise at timestep $k$ as a function of the aperture diameter $D$ in meters, the exposure time $(\Delta t)_k$ in seconds, the zenith angle $\gamma_k$, the observing station's altitude $h_\text{obs}$ in meters, and the turbulence scaleheight $H\approx8000$ meters \cite{osborn2015}.
 
@@
 
-\begin{equation}
+\begin{equation} \label{eq:lc_dist}
  S_{\text{SO},k} \sim \mathcal{N}\left( \bar{S}_{\text{SO},k}, \sigma^2_{S,k} \right).
@@
 
-\FloatBarrier
 \section{Attitude Inversion Results for Synthetic Data}
 
-To demonstrate the effectiveness of this method, we first present results for a simulated light curve where the ground truth is known.
+\subsection{Generating the Observed Light Curve}
 
-\begin{table}[ht]
+To demonstrate the effectiveness of our inversion method, we first present results for a simulated light curve where the ground truth is known and material property mismatch is insignificant.
+We produce a noisy, observed light curve for a high-fidelity scaled model of the Saturn V second stage to act as a template rocket body object, shown in Figures \ref{fig:meas_model}.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\figsmall]{rocket_bodies2.png}
+  \caption{Measured rocket body model \cite{nasa_models}.}
+  \label{fig:meas_model}
+\end{figure}
+
+Table \ref{tb:case1_in} describes the simulated observation scenario while Table \ref{tb:synth_matprops} lists the material properties used in this test case.
+
+\begin{table}[H]
   \centering
@@
 
-\begin{table}[ht]
+\begin{table}[H]
   \centering
@@
   \end{tabular}
-  \label{tb:real_matprops}
+  \label{tb:synth_matprops}
 \end{table}
 
-\begin{table}[ht]
+The true attitude profile and inertia tensor ratios of the object are listed in Table \ref{tb:synth_att}.
+
+\begin{table}[H]
   \centering
@@
   \textbf{Variable} & \textbf{Value} \\ \hline
- Initial orientation & $\vctr{p}(0) = -\frac{1}{3} \begin{bmatrix} 1 & 1 & 1 \end{bmatrix}^T$ \\ \hline
- Initial $\vctr{\omega}$ ($\text{rad}/\text{s}$) & $\vctr{\omega}(0) = \begin{bmatrix} 0.03 & 0.06 & 0.03 \end{bmatrix}^T$ \\ \hline
+ Initial $\vctr{p}(0)$ & $-\frac{1}{3} \begin{bmatrix} 1 & 1 & 1 \end{bmatrix}^T$ \\ \hline
+ True $\vctr{\omega}(0)$ ($\text{rad}/\text{s}$) & $\vctr{\omega}(0) = \begin{bmatrix} 0.03 & 0.06 & 0.03 \end{bmatrix}^T$ \\ \hline
  Inertia tensor ratios & $J_y / J_x = 1, \: J_z / J_x = 0.25$ \\ \hline
- External torque & $\vctr{M}=\begin{bmatrix} 0 & 0 & 0 \end{bmatrix}^T$ \\ \hline
+ External torque $\vctr{M}$ & $\begin{bmatrix} 0 & 0 & 0 \end{bmatrix}^T$ \\ \hline
   \end{tabular}
-  \label{tb:synth_truth}
+  \label{tb:synth_att}
 \end{table}
+\FloatBarrier
 
-\begin{figure}[ht]
+Simulating the mean light curve and its approximate probability distribution via Equation \ref{eq:lc_dist} produces the light curve displayed in ADU in Figure \ref{fig:obs_adu_synth} and V-band magnitude in Figure \ref{fig:obs_mag_synth}.
+
+\begin{figure}[H]
   \centering
-  \includegraphics[width=\figsmall]{rocket_bodies2.png}
-  \caption{Measured rocket body model \cite{nasa_models}.}
-  \label{fig:meas_model}
+  \includegraphics[width=\figsmall]{light_curves_adu_rocket_body.png}
+  \caption{Simulated observed flux for the rocket body.}
+  \label{fig:obs_adu_synth}
 \end{figure}
 
-\begin{figure}[ht]
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\figsmall]{light_curves_mag_rocket_body.png}
+  \caption{Simulated observed V-band magnitude for the rocket body.}
+  \label{fig:obs_mag_synth}
+\end{figure}
+
+% \begin{figure}[H]
+%   \centering
+%   \includegraphics[width=\figsmall]{light_curves_std_rocket_body.png}
+%   \caption{Simulated observed light curve standard deviation computed with Equation \ref{eq:sigma_total}.}
+%   \label{fig:obs_std_synth}
+% \end{figure}
+
+\subsection{Synthetic Light Curve Optimization Results}
+
+To emulate the reality that the precise geometry of the observed object is usually uncertain, we use perform attitude inversion with a significantly simplified rocket body model, shown in Figure \ref{fig:inv_model}.
+
+\begin{figure}[H]
   \centering
@@
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\figsmall]{light_curves_adu_rocket_body.png}
-  \caption{Simulated observed flux for the rocket body.}
-  \label{fig:obs_adu_synth}
-\end{figure}
+A total of $10^5$ initial state guesses were created via the sampling scheme described in Section \ref{sec:run_solver} with $|\omega(0)|_\text{max}=5.26$ degrees per second using an angular velocity padding factor of $f_\omega=1.25$.
+BFGS minimization was then performed for each initial guess to locally solve the optimization problem given by Equation \ref{eq:opt_problem}.
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\figsmall]{light_curves_mag_rocket_body.png}
-  \caption{Simulated observed V-band magnitude for the rocket body.}
-  \label{fig:obs_mag_synth}
-\end{figure}
+Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} display the converged MRP and angular velocity vectors of the best solutions TODO: define how we are determining what's best.
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\figsmall]{light_curves_std_rocket_body.png}
-  \caption{Simulated observed light curve standard deviation computed with Equation \ref{eq:sigma_total}.}
-  \label{fig:obs_std_synth}
-\end{figure}
-
-\begin{figure}[ht]
+\begin{figure}[H]
   \centering
@@
   \caption{Converged solution initial orientation MRP vectors.}
-  \label{fig:omega_sols1}
+  \label{fig:sigma_sols1}
 \end{figure}
 
-\begin{figure}[ht]
+\begin{figure}[H]
   \centering
@@
   \caption{Converged solution body-frame spin and precession rates.}
-  \label{fig:omega_sols1}
+  \label{fig:spin_proc1}
 \end{figure}
 
-\FloatBarrier
+The inversion
+
 \section{Attitude Inversion Results for Real Data}
@@
 
-The contribution of the bus to the total inertia tensor of ECHOSTAR II is spherically symmetric and given by: [\textit{TODO: citation needed}]:
+The contribution of the bus to the total inertia tensor of ECHOSTAR II is spherically symmetric and given by \cite{satterly1958}:
 
@@
 assuming that the center of mass (COM) of the bus coincides with the COM of the entire vehicle.
-By the intermediate axis theorem, the contribution of a panel to the inertia tensor is a function of the displacement of the panel's COM from the vehicle's overall COM $\vctr{r}_p = [ 0, s/2 - l_p/2, 0]^T$ [\textit{TODO: citation needed}]:
+By the intermediate axis theorem, the contribution of a panel to the inertia tensor is a function of the displacement of the panel's COM from the vehicle's overall COM $\vctr{r}_p = [ 0, s/2 - l_p/2, 0]^T$:
 
@@
 \noindent
-assuming that the panels have negligible thickness. The overall inertia tensor for ECHOSTAR II is:
+assuming that the panels have negligible thickness. The overall inertia tensor for ECHOSTAR II is therefore:
 
@@
 
-We estimate a maximum bound for the angular velocity magnitude by inspecting the light curve [\textit{TODO: finish sentence}]
+We estimate a maximum bound for the angular velocity magnitude by taking the Fourier transform of the light curve and choosing the highest-power positive frequency, converting to radians per second and multiplying by a factor of safety $f_\omega > 1$.
+TODO: mention this factor in an equation.
 
-\begin{table}[ht]
+\begin{table}[H]
   \centering
@@
 
-\FloatBarrier
 \section*{Acknowledgments}
@@
 
-\FloatBarrier
 \section*{Appendix}
 
-\begin{table}[ht]
+\begin{table}[H]
   \centering
@@
 
- \begin{table}[ht]
+ \begin{table}[H]
   \centering
@@
 
-\FloatBarrier
-\clearpage
 \bibliography{cache/refs.bib}

### 2025-03-10 paper.tex
need to add plots and discussion for real data

@@
 
-\pagecolor[rgb]{0.0,0.0,0.0}
-\color[rgb]{1,1,1}
+% \pagecolor[rgb]{0.0,0.0,0.0}
+% \color[rgb]{1,1,1}
 
@@
 Due to atmospheric distortion and aperture size, it is generally impossible to resolve even large satellites and rocket bodies in optical ground-based imagery.
-As a result, shape and orientation information is unavailable in any single image, but a measurement of the object's total brightness can still be obtained.
+As a result, shape and orientation information is unavailable in individual images, but a measurement of the object's total brightness can still be obtained.
 Even if the object's shape and reflective properties are known, any given brightness measurement will generally correspond to infinitely many possible orientations.
@@
 Given a light curve with an estimate of the shape and reflective properties of the object, our procedure searches the space of initial orientations, angular velocities, and inertia tensors to find initial conditions that produce light curves with low error compared to the observed values.
-Our formulation inherently accounts for the multiple types of ambiguities resulting from the object and observation geometry, the physical constraints of torque-free rigid body motion, and noise in the light curve.
+Our formulation inherently accounts for the multiple ambiguities resulting from the object and observation geometry, the physical constraints of torque-free rigid body motion, and noise in the light curve.
 
@@
 If a reliable prior distribution for the attitude state is not available -- which it generally is not -- a Bayesian method may be inadvisable.
-In the absence of any prior knowledge about the object's attitude state, simulated annealing \cite{gagnon2024, clark2020}, genetic algorithms \cite{gagnon2024, piergentili2017, clark2020}, and particle swarm optimizers (PSOs) \cite{clark2020, clark2022, burton2024journal, burton2024scitech, gagnon2024} can be applied to search the solution space exhaustively.
+In the absence of prior knowledge about the object's attitude state, simulated annealing \cite{gagnon2024, clark2020}, genetic algorithms \cite{gagnon2024, piergentili2017, clark2020}, and particle swarm optimizers (PSOs) \cite{clark2020, clark2022, burton2024journal, burton2024scitech, gagnon2024} can be applied to search the solution space exhaustively.
 
@@
 
-% .
-Solutions for the GLONASS case are analyzed to highlight the effect of observation geometry ambiguities, while the rocket body case highlights ambiguities introduced by the geometry of the object.
-Our noise model is validated against further real observations, confirming that all relevant noise sources have been accounted for and modeled accurately.
-We show that any significant lack of knowledge of the object's shape or optical material properties causes the solution to rapidly degrade.
-
 \section{Inversion Method}
@@
 
-If the reflectivity of the object's faces is not well-known, we rescale the estimated light curve mean $\hat{S}(\vctr{x})$ to have equal norm to the observations to resolve the inherent albedo-area ambiguity:
+If the reflectivity of the object's faces is not well-known, we rescale the estimated light curve mean $\hat{S}(\vctr{x})$ to have equal $\ell_2$ norm to the observations to resolve the inherent albedo-area ambiguity:
 
@@
 
-\subsection{Running the Solver} \ref{sec:run_solver}
+\subsection{Running the Solver} \label{sec:run_solver}
 
-Our method first initializes $n$ guesses scattered throughout the solution space -- possibly sampled from a prior distribution based on available knowledge.
+To initialize the solver, we initialize $n$ guesses scattered throughout the solution space -- possibly sampled from a prior distribution based on available knowledge.
 In the most basic case, the initial state guess is sampled for guess $i$ at iteration $k=0$:
@@
 \noindent
-In this work, $\vctr{p}_{i,0}(0)3$ is uniformly randomly sampled from orientation space, $\vctr{\omega}_{i,0}(0)$ is uniformly randomly sampled from the ball in $\mathbb{R}^3$ of radius $|\omega(0)|_\text{max}$ -- the maximum angular rate to search -- while $\left(J_y / J_x\right)_{i,0}$ and $\left(J_z / J_x\right)_{i,0}$ are sampled from Gaussian prior distributions centered at estimated values derived from the known geometry of the object.
+In this work, $\vctr{p}_{i,0}(0)3$ is uniformly randomly sampled from orientation space, $\vctr{\omega}_{i,0}(0)$ is uniformly randomly sampled from the ball in $\mathbb{R}^3$ of radius $\norm{\vctr{\omega}}_\text{max}$ -- the maximum angular rate to search -- while $\left(J_y / J_x\right)_{i,0}$ and $\left(J_z / J_x\right)_{i,0}$ are sampled from Gaussian prior distributions centered at estimated values derived from the known geometry of the object.
+We determine $\norm{\vctr{\omega}}_\text{max}$ based on the maximum power positive frequency in the observed light curve $f^\star$ in Hertz and an arbitrary safety factor $f_\omega \geq 1$, selected as $f_\omega = 1.25$ for this work:
+
+\begin{equation} \label{eq:ang_vel_max}
+  \norm{\vctr{\omega}}_\text{max} = 2\pi f_\omega f^\star.
+\end{equation}
 
@@
 \begin{equation} \label{eq:opt_problem}
-  \vctr{x}_{k,i}(0) = \argmin_x f(\vctr{x}),
+ \vctr{x}_{k,i}(0) = \argmin_{\vctr{x}} f(\vctr{x}),
 \end{equation}
@@
 \noindent
-producing a final initial state which is a good local guess for a state which to produce the observations.
+producing a final initial state which is a good local guess for a state that produces the observations.
 In this way, we globally search the solution space up to the density implied by $n$.
@@
 In order to implement this optimization effectively without constraints, it is often necessary to take the absolute value of the intermediate inertia tensor values to maintain physical validity.
+
+\subsection{Classifying Candidate Solutions} \label{sec:candidate_sols}
+
+Although all initial conditions will converge to some nearby local minimum, many of these minima will possess a low likelihood determined by $-\mathrm{exp}(f(\vctr{x}))$.
+As we do not generally know how large the geometric and material mismatch between the observed object and the assumed model is, we cannot put a strict bound on a likelihood level that constitutes a high-quality candidate solution.
+In this work, we identify candidate solutions by choosing a likelihood ratio tolerance $0 < \ell < 1$, such that all solutions $\vctr{x}_{i,k}(0)$ are candidates for the truth if:
+
+\begin{equation}
+  e^{-f(\vctr(x)_{i,k}(0))} > \ell e^{-f(\vctr(x)^\star_{k}(0))}.
+\end{equation}
+
+Here, $\vctr(x)^\star_{k}(0) = \argmin_{\vctr{x}_{i,k}(0)} f(\vctr(x)_{i,k}(0)) $ is the highest likelihood solution found after optimization.
 
@@
 For highly nonconvex and detailed models composed of thousands of faces, $\bar{a}$ can be approximated on a pixel grid using shadow mapping \cite{robinson2022}.
-Self-shadowing can be solved semi-analytically for less detailed approximations of nonconvex objects by computing the mutual intersections of the polygons $P_k$ on other faces whose projections overlap with the face in question.
+Self-shadowing can be efficiently solved semi-analytically for lower-fidelity approximations of real space object geometry by computing the mutual intersections of the polygons $P_k$ of other faces whose projections overlap with the face in question.
 Up to a maximum intersection depth $d$, the unshadowed area can be computed via:
@@
 We choose the Blinn-Phong \cite{blinn1977} BRDF for this work as it is efficient to compute and satisfies the three main requirements for a physically meaningful BRDF as it is nonnegative, energy-conserving, and reciprocal \cite{duvenhage2013}.
-We avoid more recent microfacet models for computational efficiency.
-In practice, the uncertainty in the reflective properties of the observed objects will often be larger than errors introduced by the choice of reflection model.
+We avoid microfacet models for computational efficiency, and because in practice, the uncertainty in the reflective properties of the observed objects will often be larger than errors introduced by the choice of reflection model.
 The Blinn-Phong BRDF is parameterized by the coefficient of diffuse reflectivity $C_d$, the coefficient of specular reflectivity $C_s$, and the specular exponent $n>0$:
@@
 To demonstrate the effectiveness of our inversion method, we first present results for a simulated light curve where the ground truth is known and material property mismatch is insignificant.
-We produce a noisy, observed light curve for a high-fidelity scaled model of the Saturn V second stage to act as a template rocket body object, shown in Figures \ref{fig:meas_model}.
+We produce a noisy, observed light curve for a high-fidelity model of the Saturn V second stage scaled down by a factor $10/3$ to act as a template rocket body object, shown in Figures \ref{fig:meas_model}.
 
@@
 
-Table \ref{tb:case1_in} describes the simulated observation scenario while Table \ref{tb:synth_matprops} lists the material properties used in this test case.
+Table \ref{tb:case1_in} describes the simulated observation scenario, placing the rocket body in the orbit of an SL-12 upper stage in geosynchronous orbit, observed by the Purdue Optical Ground Station.
+Relevant geographical and optical properties of this ground station are provided in the Appendix in Table \ref{tb:tele_info}.
+Table \ref{tb:synth_matprops} lists the material properties used in this test case.
 
@@
   \centering
-  \caption{Simulated observation parameters}
+  \caption{Simulated observation characteristics}
   \vspace*{6pt}
@@
 
-The true attitude profile and inertia tensor ratios of the object are listed in Table \ref{tb:synth_att}.
+The ground truth attitude profile and inertia tensor ratios of the simulated object are listed in Table \ref{tb:synth_att}.
 
@@
   \centering
-  \includegraphics[width=\figsmall]{light_curves_adu_rocket_body.png}
+  \includegraphics[width=\figmed]{light_curves_adu_rocket_body_sdc.png}
   \caption{Simulated observed flux for the rocket body.}
@@
   \centering
-  \includegraphics[width=\figsmall]{light_curves_mag_rocket_body.png}
+  \includegraphics[width=\figmed]{light_curves_mag_rocket_body_sdc.png}
   \caption{Simulated observed V-band magnitude for the rocket body.}
@@
 %   \centering
-%   \includegraphics[width=\figsmall]{light_curves_std_rocket_body.png}
+%   \includegraphics[width=\figsmall]{light_curves_std_rocket_body_sdc.png}
 %   \caption{Simulated observed light curve standard deviation computed with Equation \ref{eq:sigma_total}.}
@@
   \centering
-  \includegraphics[width=\figsmall]{rocket_bodies1.png}
+  \includegraphics[width=\figmed]{rocket_bodies1.png}
   \caption{Simplified rocket body model used for inversion.}
@@
 
-A total of $10^5$ initial state guesses were created via the sampling scheme described in Section \ref{sec:run_solver} with $|\omega(0)|_\text{max}=5.26$ degrees per second using an angular velocity padding factor of $f_\omega=1.25$.
+A total of $10^5$ initial state guesses were created via the sampling scheme described in Section \ref{sec:run_solver} with $\norm{\vctr{\omega}}_\text{max}=5.26$ degrees per second.
+The inertia tensor ratios were initialized at the truth, but are allowed to be changed during optimization.
 BFGS minimization was then performed for each initial guess to locally solve the optimization problem given by Equation \ref{eq:opt_problem}.
-
-Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} display the converged MRP and angular velocity vectors of the best solutions TODO: define how we are determining what's best.
+Classifying candidate solutions with $\ell = 1/10$ produces over $10^4$ solutions within this bound.
+Figure \ref{fig:conv_lcs_synth} shows the mean light curves produced by the identified candidate solutions.
+There is generally good agreement between the candidate light curves and the observations, indicating that high-quality local minima have been identified.
 
@@
   \centering
-  \includegraphics[width=\figsmall]{sigma_solsRocket Body.png}
-  \caption{Converged solution initial orientation MRP vectors.}
+  \includegraphics[width=\figmed]{converged_solutions_1_Rocket Body.png}
+  \caption{Candidate solution light curves found through BFGS optimization.}
+  \label{fig:conv_lcs_synth}
+\end{figure}
+
+Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} display the candidate MRP and angular velocity vectors of the best solutions determined via the procedure in Section \ref{sec:candidate_sols}.
+We notice that for both the initial orientation and angular velocity, the ground truth lies squarely in the surfaces of candidate solutions.
+In angular velocity space, most solutions fall onto the surface of two open-ended cylinders -- one with half the radius of the larger cylinder containing the truth.
+Similarly in orientation space, we solutions fall on two distinct manifolds -- likely resulting from a reflection of the object's body frame about the bisector between the Sun and observer vectors \cite{marto2024, burton2024journal}.
+One of these orientation-space surfaces contains the ground truth.
+These additional families of solutions highlight the reality that a wide variety of attitude profiles may produce the same observations.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\figmed]{sigma_solsRocket Body.png}
+  \caption{Candidate solution initial orientation MRP vectors, ground truth highlighted in red.}
   \label{fig:sigma_sols1}
@@
   \centering
-  \includegraphics[width=\figsmall]{omega_sols_Rocket Body.png}
-  \caption{Converged solution initial body-frame angular velocities.}
+  \includegraphics[width=\figmed]{omega_sols_Rocket Body.png}
+  \caption{Candidate solution initial body-frame angular velocities, ground truth highlighted in red.}
   \label{fig:omega_sols1}
@@
 
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\figsmall]{spin_proc_Rocket Body.png}
-  \caption{Converged solution body-frame spin and precession rates.}
-  \label{fig:spin_proc1}
-\end{figure}
+% While the state vectors of the candidate solutions do not uniquely indicate the ground truth, the object's precession rate is estimated
 
-The inversion
+% \begin{figure}[H]
+%   \centering
+%   \includegraphics[width=\figsmall]{spin_proc_Rocket Body.png}
+%   \caption{Converged solution body frame spin and precession rates.}
+%   \label{fig:spin_proc1}
+% \end{figure}
+
+% The inversion
 
@@
 
-The two objects selected for study in this work are a Delta I upper stage and the ECHOSTAR 2 satellite.
+% The two objects selected for study in this work are a Delta I upper stage and the ECHOSTAR 2 satellite.
 
@@
   \centering
-  \caption{Delta I rocket body observation parameters}
+  \caption{Delta I rocket body observation characteristics}
   \vspace*{6pt}
@@
  Uniform BRDF & $C_d=0.2$, $C_s=0.8$, $n=30$ \\ \hline
-  $|\omega(0)|_\text{max}$ & $8.31$ $\text{deg} / \text{s}^{-1}$ \\ \hline
+  $\norm{\vctr{\omega}}_\text{max}$ & $8.31$ $\text{deg} / \text{s}^{-1}$ \\ \hline
   \end{tabular}
@@
   \centering
-  \caption{ECHOSTAR II observation parameters}
+  \caption{ECHOSTAR II observation characteristics}
   \vspace*{6pt}
@@
 
-We estimate a maximum bound for the angular velocity magnitude by taking the Fourier transform of the light curve and choosing the highest-power positive frequency, converting to radians per second and multiplying by a factor of safety $f_\omega > 1$.
-TODO: mention this factor in an equation.
-
 \begin{table}[H]
@@
  Solar panel BRDF & $C_d=0.05$, $C_s=0.13$, $n=10$ \\ \hline
-  $|\omega(0)|_\text{max}$ & $2.32$ $\text{deg} / \text{s}^{-1}$ \\ \hline
+  $\norm{\vctr{\omega}}_\text{max}$ & $2.32$ $\text{deg} / \text{s}^{-1}$ \\ \hline
   \end{tabular}
@@
   \centering
-  \caption{Telescope information common to all test cases}
+  \caption{Telescope information for synthetic and real data observations}
   \vspace*{6pt}

### 2025-03-11 paper.tex
preparing for final writing bit

@@
   \includegraphics[width=\figbig]{sphx_glr_plot_lcs_001_2_00x.png}
-  \caption{Selected light curve for the Delta (Star-37E) rocket body, observed by the Purdue Optical Ground Station}
+  \caption{Selected light curve for the Delta I (Star-37E) rocket body, observed by the Purdue Optical Ground Station}
   \label{fig:rb_lc_obs}
@@
 \end{table}
+
+\begin{figure*}[t]
+  \centering
+  \includegraphics[width=0.8\textwidth]{sphx_glr_results_002_2_00x.png}
+  \caption{Candidate solution inertia light curves compared to the real measurements in ADU for the Delta I (Star-37E) rocket body.}
+  \label{fig:case2_s}
+\end{figure*}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\figmed]{sphx_glr_results_001_2_00x.png}
+  \caption{Candidate solution inertia ratio distribution for the Delta I (Star-37E) rocket body.}
+  \label{fig:case2_i}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\figmed]{sphx_glr_results_006.png}
+  \caption{Candidate solution orientation MRPs for the Delta I (Star-37E) rocket body.}
+  \label{fig:case2_p}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\figmed]{sphx_glr_results_005.png}
+  \caption{Candidate solution angular velocities for the Delta I (Star-37E) rocket body.}
+  \label{fig:case2_w}
+\end{figure}
 
@@
 
+\begin{figure*}[t]
+  \centering
+  \includegraphics[width=0.8\textwidth]{sphx_glr_results_004_2_00x.png}
+  \caption{Candidate solution inertia light curves compared to the real measurements in ADU for ECHOSTAR II.}
+  \label{fig:case2_s}
+\end{figure*}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\figmed]{sphx_glr_results_003_2_00x.png}
+  \caption{Candidate solution inertia ratio distribution for ECHOSTAR II.}
+  \label{fig:case2_i}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\figmed]{sphx_glr_results_008.png}
+  \caption{Candidate solution orientation MRPs for ECHOSTAR II.}
+  \label{fig:case2_p}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\figmed]{sphx_glr_results_007.png}
+  \caption{Candidate solution angular velocities for ECHOSTAR II.}
+  \label{fig:case2_w}
+\end{figure}
+
+\section{Conclusion}
+
 \section*{Acknowledgments}

### 2025-03-11 paper_cache.py

@@ -19,6 +19,7 @@ imgs = {
         'sphx_glr_plot_models_002.png',
         'sphx_glr_plot_lcs_001_2_00x.png',
         'sphx_glr_plot_lcs_002_2_00x.png',
+        *glob.glob('/Users/liamrobinson/Documents/maintained-research/mirage/docs/build/html/_images/sphx_glr_results_*.png'),
     ],
     "/Users/liamrobinson/Documents/presentations-papers/msthesis/static_images": [
     ],

### 2025-03-11 paper.tex
wrote some stuff

@@
 Here, we choose the inner radius to be $r_i=0.45d$, the outer radius to be $r_o=0.5d$, and $m=1$ kilogram -- dividing each moment of inertia by a constant does not change the dynamics -- yielding $J_a = 15.3 \: [kg \cdot m^2]$ and $J_t = 27.9 \: [kg \cdot m^2]$.
-Table \ref{tb:case1_in} summarizes the light curve observations that result in the extracted measurements displayed in Figure \ref{fig:rb_lc_obs}.
+Table \ref{tb:case2_in} summarizes the light curve observations that result in the extracted measurements displayed in Figure \ref{fig:rb_lc_obs}.
 
@@
   \end{tabular}
-  \label{tb:case1_in}
+  \label{tb:case2_in}
 \end{table}
@@
 \end{figure}
+
+This light curve shows clear periodicity, but has a high estimated standard deviation and a relatively low sampling rate due to the object's low brightness and the sensor readout time.
+Inversion was performed using $10^4$ initial guesses with the material and angular velocity assumptions listed in Table \ref{tb:case2_ass}.
+The candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case2_s}, inertia ratios in Figure \ref{fig:case2_i}, orientation MRPs in Figure \ref{fig:case2_p}, and angular velocities in Figure \ref{fig:case2_w}.
 
@@
  $\norm{\vctr{\omega}}_\text{max}$ & $8.31$ $\text{deg} / \text{s}^{-1}$ \\ \hline
+ $\ell$ & $1/100$ \\ \hline
 \end{tabular}
-  \label{tb:case1_ass}
+  \label{tb:case2_ass}
 \end{table}
@@
   \includegraphics[width=0.8\textwidth]{sphx_glr_results_002_2_00x.png}
-  \caption{Candidate solution inertia light curves compared to the real measurements in ADU for the Delta I (Star-37E) rocket body.}
+  \caption{Candidate solution light curves compared to the real measurements in ADU for the Delta I (Star-37E) rocket body.}
   \label{fig:case2_s}
@@
   \includegraphics[width=\figmed]{sphx_glr_results_001_2_00x.png}
-  \caption{Candidate solution inertia ratio distribution for the Delta I (Star-37E) rocket body.}
+  \caption{Candidate solution inertia ratio distribution for the Delta I (Star-37E) rocket body, non-candidate converged minima are plotted in grey.}
   \label{fig:case2_i}
@@
   \includegraphics[width=\figmed]{sphx_glr_results_006.png}
-  \caption{Candidate solution orientation MRPs for the Delta I (Star-37E) rocket body.}
+  \caption{Candidate solution orientation MRPs for the Delta I (Star-37E) rocket body, non-candidate converged minima are plotted in grey.}
   \label{fig:case2_p}
@@
   \includegraphics[width=\figmed]{sphx_glr_results_005.png}
-  \caption{Candidate solution angular velocities for the Delta I (Star-37E) rocket body.}
+  \caption{Candidate solution angular velocities for the Delta I (Star-37E) rocket body, non-candidate converged minima are plotted in grey.}
   \label{fig:case2_w}
 \end{figure}
+
+The candidate solutions identified by the inversion process produce light curves which adequately approximate the measurements, although they do not reach the lowest observed values.
+This indicates that the true material is more dominated by specular reflections than our assumed coefficients.
+The candidates show some rough clustering in both orientation and angular velocity space, although they vary significantly in angular velocity magnitude.
+The inertia tensor solutions display similar variability, although the solutions are spread relatively uniformly from the initial guess, possibly indicating that it was reasonably unbiased.
+Taken together, these results indicate that the object could be spinning in a number of attitude profiles, and that more information is needed about its true mass distribution and reflectivity to further refine the solution set.
 
@@
 
+Table \ref{tb:case2_in} summarizes the light curve observations that result in the extracted measurements displayed in Figure \ref{fig:rb_lc_obs}.
+
 \begin{table}[H]
@@
   \end{tabular}
-  \label{tb:case2_in}
+  \label{tb:case3_in}
 \end{table}
@@
 \end{figure}
+
+This light curve is sampled more densely than the Delta I observations, and clearly has significantly more intense specular peaks.
+Inversion was performed using $10^4$ initial guesses with the material and angular velocity assumptions listed in Table \ref{tb:case3_ass}.
+The candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case3_s}, inertia ratios in Figure \ref{fig:case3_i}, orientation MRPs in Figure \ref{fig:case3_p}, and angular velocities in Figure \ref{fig:case3_w}.
 
@@
   $\norm{\vctr{\omega}}_\text{max}$ & $2.32$ $\text{deg} / \text{s}^{-1}$ \\ \hline
+  $\ell$ & $1/1000$ \\ \hline
   \end{tabular}
-  \label{tb:case2_ass}
+  \label{tb:case3_ass}
 \end{table}
@@
   \caption{Candidate solution inertia light curves compared to the real measurements in ADU for ECHOSTAR II.}
-  \label{fig:case2_s}
+  \label{fig:case3_s}
 \end{figure*}
@@
   \includegraphics[width=\figmed]{sphx_glr_results_003_2_00x.png}
-  \caption{Candidate solution inertia ratio distribution for ECHOSTAR II.}
-  \label{fig:case2_i}
+  \caption{Candidate solution inertia ratio distribution for ECHOSTAR II, non-candidate converged minima are plotted in grey.}
+  \label{fig:case3_i}
 \end{figure}
@@
   \includegraphics[width=\figmed]{sphx_glr_results_008.png}
-  \caption{Candidate solution orientation MRPs for ECHOSTAR II.}
-  \label{fig:case2_p}
+  \caption{Candidate solution orientation MRPs for ECHOSTAR II, non-candidate converged minima are plotted in grey.}
+  \label{fig:case3_p}
 \end{figure}
@@
   \includegraphics[width=\figmed]{sphx_glr_results_007.png}
-  \caption{Candidate solution angular velocities for ECHOSTAR II.}
-  \label{fig:case2_w}
+  \caption{Candidate solution angular velocities for ECHOSTAR II, non-candidate converged minima are plotted in grey.}
+  \label{fig:case3_w}
 \end{figure}
 
+The candidate solutions identified by the inversion process are more sparse than those found in either of the previous rocket body cases, with some vague clustering in both angular velocity and orientation space.
+That said, the candidate light curves fit the data quite well.
+Notably, the formulation of the negative log-likelihood objective function in Equation \ref{eq:nll_loss} produces the behavior seen in the candidate solutions in Figure \ref{fig:case3_s}, where even the smallest recurrent specular peaks are well-estimated by most solutions.
+A similar effect -- giving up some of the statistical robustness of our formulation -- can be achieved in an unweighted optimization over the observations in magnitudes.
+The inertia tensor solutions are dispersed throughout the solution space, with less symmetry than the Delta I solutions.
+As in the Delta I case, more data about the mass and reflectivity of the object would be necessary to continue refining any of these candidates.
+
 \section{Conclusion}
+
+Attitude information is critical for many tasks informing space situational awareness.
+Orientation and angular velocity data aids high-fidelity orbit propagation, mission recovery efforts, and object selection for active debris removal.
+Obtaining attitude information is often difficult, especially for passive debris objects.
+Due to atmospheric distortion and diffraction-limited optics, ground-based telescopes can only determine a total brightness received from a space object.
+If a shape model is known, the light curve -- a sequence of these brightness measurements -- is one tool for attitude estimation.
+We presented an approach for solving this highly nonlinear and non-convex optimization problem with a gradient-based multistart method.
+We presented attitude inversion results for one case with synthetic data and two cases with real observations.
+Unlike other approaches, our method successfully identifies the numerous ambiguous solutions families -- originating from symmetries in the observation geometry and the object shape -- which may produce the same observations.
+These disparate candidate solutions can serve as starting points for more detailed exploration as additional light curves or further object shape information is obtained.
 

### 2025-03-12 paper.tex
fixed most comments for this round

@@
 If no initial attitude state is available, as in most cases of passive debris observation, the inversion problem must be solved globally.
-Filtering approaches can be applied here too; particle \cite{linares2014particle, holzinger2014} and Bayesian multi-hypothesis filters \cite{burton2021two, cabrera2023} have shown promise in prior work.
-If a reliable prior distribution for the attitude state is not available -- which it generally is not -- a Bayesian method may be inadvisable.
+Filtering approaches can be applied here too; particle \cite{linares2014particle, holzinger2014} and Bayesian Multi-Hypothesis filters \cite{burton2021two, cabrera2023} have shown promise in prior work.
+Bayesian methods may be ineffective if a reliable prior distribution for the attitude state is not available.
 In the absence of prior knowledge about the object's attitude state, simulated annealing \cite{gagnon2024, clark2020}, genetic algorithms \cite{gagnon2024, piergentili2017, clark2020}, and particle swarm optimizers (PSOs) \cite{clark2020, clark2022, burton2024journal, burton2024scitech, gagnon2024} can be applied to search the solution space exhaustively.
 
-Much of the past literature on attitude inversion from real light curve data is centered on determining the constant spin rate and axis of rotation of the observed object.
+While much work has performed full state attitude inversion on simulated light curve data \cite{burton2024journal, gagnon2024, robinson2025att}, studies using real data often focus on determining a constant spin rate and axis of rotation for the observed object.
 Frequency analysis in the form of periodograms, Fourier transforms, or epoch folding, is often used to determine the spin rate based on the light curve's frequency spectrum \cite{silha2015, silha2021, isoletta2024, schildknecht2015, pittet2017, yanagisawa2012, koshkin2018}.
@@
 Further, our method differs from other global estimation methods in that we do not seek to find one optimal solution and instead search for a collection of high-quality local minima to directly study the many ambiguous solutions for a given light curve.
+Our approach was designed specifically for use with real data as we take time-varying measurement noise and uncertainties in the body moments of inertia into account explicitly.
+
 To solve the optimization problem, we sample many candidate solutions scattered throughout the solution space and perform gradient-based optimization on each using the BFGS algorithm.
@@
 
-This work proceeds by describing the inversion method, discussing efficient methods for computing the time-varying probability density of the light curve, and presenting inversion results for the two test cases.
+This work proceeds by describing the inversion method, discussing efficient methods for computing the time-varying probability density of the light curve.
+We present inversion results for a simulated rocket body light curve where the ground truth is known, followed by two cases -- a derelict satellite and rocket body -- using real light curve data collected by the Purdue Optical Ground Station.
+A deeper treatment of our method and additional simulated light curve results can be found in a forthcoming publication by the authors \cite{robinson2025att}.
 
@@
 
-\subsection{Objective Function}
-
-Our inversion method is based on the global minimization of a maximum-likelihood loss function which computes the negative log-likelihood of the observed light curve $\vctr{S}$ being a sample from the normal distribution defined by the estimated light curve mean $\hat{\vctr{S}}$ and the estimated standard deviation at each $k$th timestep $\sigma_k$:
-
-\begin{equation} \label{eq:nll_loss}
- f(\vctr{x}) = \frac{1}{m}\sum_{k=1}^{m}\left[\ln\sigma_k + \frac{1}{2}\left(\frac{S_k - \hat{S}_k(\vctr{x})}{\sigma_k}\right)^2 \right].
- \end{equation}
-
-If the reflectivity of the object's faces is not well-known, we rescale the estimated light curve mean $\hat{S}(\vctr{x})$ to have equal $\ell_2$ norm to the observations to resolve the inherent albedo-area ambiguity:
-
-\begin{equation}
- \hat{S}(\vctr{x}) = \frac{\norm{S}}{\norm{\tilde{S}(\vctr{x})}} \tilde{S}(\vctr{x}),
-\end{equation}
-
-\noindent
-where $\tilde{S}(\vctr{x})$ is the original, poorly scaled, estimated light curve.
-In this work, the light curve and its standard deviation are always assumed to be expressed in a linear unit, e.g., analog-to-digital units (ADU) or irradiance proportional to $W/m^2$.
-
 \subsection{State Definition and Dynamics}
 
-The state vector $\vctr{x}$ we optimize to fit the observed light curve is the concatenation of the Modified Rodriguez Parameter (MRP) $\vctr{p} = [p_1, p_2, p_3]^T$, the principal body frame angular velocity $\vctr{\omega} = [\omega_1, \omega_2, \omega_3]$ in radians per second, and the ratios of the principal moments of inertia $J_y / J_x$ and $J_z / J_x$ where the inertia tensor in principal axes is $J = \mathrm{diag}\left([J_x, J_y, J_z]\right)$:
+The state vector $\vctr{x}$ we optimize to fit the observed light curve is the concatenation of the Modified Rodriguez Parameter (MRP) $\vctr{p} = [p_1, p_2, p_3]^T$, the principal body frame angular velocity $\vctr{\omega} = [\omega_1, \omega_2, \omega_3]$ in radians per second, and the ratios of the principal moments of inertia $J_y / J_x$ and $J_z / J_x$ where the constant inertia tensor in principal axes is $J = \mathrm{diag}\left([J_x, J_y, J_z]\right)$:
 
@@
 \end{align}
+
+\subsection{Objective Function}
+
+Our inversion method is based on the global minimization of a maximum-likelihood loss function which computes the negative log-likelihood of the observed light curve $\vctr{S}$ being a sample from the normal distribution defined by the estimated light curve mean $\hat{\vctr{S}}$ and the estimated standard deviation at each $k$th timestep $\sigma_k$:
+
+\begin{equation} \label{eq:nll_loss}
+ f(\vctr{x}) = \frac{1}{m}\sum_{k=1}^{m}\left[\ln\sigma_k + \frac{1}{2}\left(\frac{\vctr{S}_k - \hat{\vctr{S}}_k(\vctr{x})}{\sigma_k}\right)^2 \right].
+ \end{equation}
+
+If the reflectivity of the object's faces is not well-known, we rescale the estimated light curve mean $\hat{S}(\vctr{x})$ to have equal $\ell_2$ norm to the observations to resolve the inherent albedo-area ambiguity:
+
+\begin{equation}
+ \hat{S}(\vctr{x}) = \frac{\norm{S}}{\norm{\tilde{S}(\vctr{x})}} \tilde{S}(\vctr{x}),
+\end{equation}
+
+\noindent
+where $\tilde{S}(\vctr{x})$ is the original, poorly scaled, estimated light curve.
+In this work, the light curve and its standard deviation are always assumed to be expressed in a linear unit, e.g., analog-to-digital units (ADU) or irradiance proportional to $W/m^2$.
 
@@
 We avoid microfacet models for computational efficiency, and because in practice, the uncertainty in the reflective properties of the observed objects will often be larger than errors introduced by the choice of reflection model.
-The Blinn-Phong BRDF is parameterized by the coefficient of diffuse reflectivity $C_d$, the coefficient of specular reflectivity $C_s$, and the specular exponent $n>0$:
+The Blinn-Phong BRDF is parameterized by the coefficient of diffuse reflectivity $C_d$, the coefficient of specular reflectivity $C_s$, and the specular exponent $n>0$ \cite{duvenhage2013}:
 
@@
 \subsection{Approximating the Measurement Noise}
+
+This section summarizes an in-depth study of relevant noise sources in CCD and CMOS sensor light curves by the authors in a forthcoming publication \cite{robinson2025twin}.
 
@@
 To demonstrate the effectiveness of our inversion method, we first present results for a simulated light curve where the ground truth is known and material property mismatch is insignificant.
-We produce a noisy, observed light curve for a high-fidelity model of the Saturn V second stage scaled down by a factor $10/3$ to act as a template rocket body object, shown in Figures \ref{fig:meas_model}.
+We produce a noisy, observed light curve for a high-fidelity model of the Saturn V second stage scaled down by a factor $10/3$ to act as a template rocket body object, shown on the right of Figure \ref{fig:rb_models}.
 
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\figsmall]{rocket_bodies2.png}
-  \caption{Measured rocket body model \cite{nasa_models}.}
-  \label{fig:meas_model}
-\end{figure}
+% \begin{figure}[H]
+%   \centering
+%   \includegraphics[width=\figsmall]{rocket_bodies2.png}
+%   \caption{Simulated truth rocket body model \cite{nasa_models}.}
+%   \label{fig:meas_model}
+% \end{figure}
 
@@
 
-Simulating the mean light curve and its approximate probability distribution via Equation \ref{eq:lc_dist} produces the light curve displayed in ADU in Figure \ref{fig:obs_adu_synth} and V-band magnitude in Figure \ref{fig:obs_mag_synth}.
+Simulating the mean light curve and its approximate probability distribution via Equation \ref{eq:lc_dist} produces the light curve displayed in V-band magnitude in Figure \ref{fig:obs_mag_synth}.
 
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\figmed]{light_curves_adu_rocket_body_sdc.png}
-  \caption{Simulated observed flux for the rocket body.}
-  \label{fig:obs_adu_synth}
-\end{figure}
+% \begin{figure}[H]
+%   \centering
+%   \includegraphics[width=\figmed]{light_curves_adu_rocket_body_sdc.png}
+%   \caption{Simulated observed flux for the rocket body.}
+%   \label{fig:obs_adu_synth}
+% \end{figure}
 
@@
 
-To emulate the reality that the precise geometry of the observed object is usually uncertain, we use perform attitude inversion with a significantly simplified rocket body model, shown in Figure \ref{fig:inv_model}.
+To emulate the reality that the precise geometry of the observed object is usually uncertain, we use perform attitude inversion with a significantly simplified rocket body model, shown in Figure \ref{fig:rb_models}.
 
@@
   \centering
-  \includegraphics[width=\figmed]{rocket_bodies1.png}
-  \caption{Simplified rocket body model used for inversion.}
-  \label{fig:inv_model}
+  \includegraphics[width=\figmed]{rocket_bodies.png}
+  \caption{Simplified rocket body model used for inversion (left) and simulated ground truth model (right).}
+  \label{fig:rb_models}
 \end{figure}
 
-A total of $10^5$ initial state guesses were created via the sampling scheme described in Section \ref{sec:run_solver} with $\norm{\vctr{\omega}}_\text{max}=5.26$ degrees per second.
-The inertia tensor ratios were initialized at the truth, but are allowed to be changed during optimization.
+A total of $10^5$ initial state guesses are created via the sampling scheme described in Section \ref{sec:run_solver} with $\norm{\vctr{\omega}}_\text{max}=5.26$ degrees per second.
+The inertia tensor ratios are initialized by sampling a symmetric Gaussian distribution with $\sigma=0.05 \: [kg\cdot m^2]$ on each axis to simulate a moderate level of inertia uncertainty.
 BFGS minimization was then performed for each initial guess to locally solve the optimization problem given by Equation \ref{eq:opt_problem}.

### 2025-03-17 paper.tex
merged plots

@@
 
+\usepackage{subcaption}
 \usepackage{times}
@@
 
-To demonstrate the effectiveness of our inversion method, we first present results for a simulated light curve where the ground truth is known and material property mismatch is insignificant.
+To demonstrate the effectiveness of our inversion method, we first present results for a simulated light curve where the ground truth is known.
+In order to make this simulated case realistic, we use the presented measurement noise models and introduce a mismatch between the ground truth object used to generate the measurements and the model used in inversion.
 We produce a noisy, observed light curve for a high-fidelity model of the Saturn V second stage scaled down by a factor $10/3$ to act as a template rocket body object, shown on the right of Figure \ref{fig:rb_models}.
@@
   \centering
-  \includegraphics[width=\figmed]{sigma_solsRocket Body.png}
+  \begin{subfigure}[t]{0.23\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{sigma_solsRocket Body.png}
+    \caption{Uniform prior}
+    \label{fig:sigma_sols1a}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.23\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{sigma_solsRocket Body Prior.png}
+    \caption{Informative prior}
+    \label{fig:sigma_sols1b}
+  \end{subfigure}
+
   \caption{Candidate solution initial orientation MRP vectors, ground truth highlighted in red.}
@@
   \centering
-  \includegraphics[width=\figmed]{omega_sols_Rocket Body.png}
-  \caption{Candidate solution initial body-frame angular velocities, ground truth highlighted in red.}
+  \begin{subfigure}[t]{0.23\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{omega_sols_Rocket Body.png}
+    \caption{Uniform prior}
+    \label{fig:omega_sols1a}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.23\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{omega_sols_Rocket Body Prior.png}
+    \caption{Informative prior}
+    \label{fig:omega_sols1b}
+  \end{subfigure}
+
+  \caption{Candidate solution initial body-frame angular velocity vectors, ground truth highlighted in red. TODO: change the perspective on the angular velocities}
   \label{fig:omega_sols1}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \begin{subfigure}[t]{0.23\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{converged_solutions_2_Rocket Body.png}
+    \caption{Uniform prior}
+    \label{fig:i_sols1a}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.23\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{converged_solutions_2_Rocket Body Prior.png}
+    \caption{Informative prior}
+    \label{fig:i_sols1b}
+  \end{subfigure}
+
+  \caption{Candidate solution inertia ratios, ground truth highlighted in red.}
+  \label{fig:i_sols1}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \begin{subfigure}[t]{0.23\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{converged_solutions_3_Rocket Body.png}
+    \caption{Uniform prior}
+    \label{fig:w_vs_ang_error_sols1a}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.23\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{converged_solutions_3_Rocket Body Prior.png}
+    \caption{Informative prior}
+    \label{fig:w_vs_ang_error_sols1b}
+  \end{subfigure}
+
+  \caption{Candidate solution angular orientation error and angular velocity magnitudes compared to the truth}
+  \label{fig:w_vs_ang_error_sols1}
 \end{figure}
@@
 Inversion was performed using $10^4$ initial guesses with the material and angular velocity assumptions listed in Table \ref{tb:case2_ass}.
-The candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case2_s}, inertia ratios in Figure \ref{fig:case2_i}, orientation MRPs in Figure \ref{fig:case2_p}, and angular velocities in Figure \ref{fig:case2_w}.
+The candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case2_s}, orientation MRPs and angular velocities in Figure \ref{fig:case2_pw}, and inertia ratios in Figure \ref{fig:case2_i}.
 
@@
 
+% \begin{figure}[H]
+%   \centering
+%   \includegraphics[width=\figmed]{sphx_glr_results_006.png}
+%   \caption{}
+%   \label{fig:case2_p}
+% \end{figure}
+
+\begin{figure}[H]
+  \centering
+  \begin{subfigure}[t]{0.23\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{sphx_glr_results_006.png}
+    \caption{MRP}
+    \label{fig:case2_pwa}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.23\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{sphx_glr_results_005.png}
+    \caption{Angular velocity}
+    \label{fig:case2_pwb}
+  \end{subfigure}
+
+  \caption{Candidate solution orientation MRPs and angular velocity vectors for the Delta I (Star-37E) rocket body}
+  \label{fig:case2_pw}
+\end{figure}
+
+% \begin{figure}[H]
+%   \centering
+%   \includegraphics[width=\figmed]{sphx_glr_results_005.png}
+%   \caption{Candidate solution angular velocities for the Delta I (Star-37E) rocket body}
+%   \label{fig:case2_w}
+% \end{figure}
+
 \begin{figure}[H]
@@
   \includegraphics[width=\figmed]{sphx_glr_results_001_2_00x.png}
-  \caption{Candidate solution inertia ratio distribution for the Delta I (Star-37E) rocket body, non-candidate converged minima are plotted in grey.}
+  \caption{Candidate solution inertia ratio distribution for the Delta I (Star-37E) rocket body}
   \label{fig:case2_i}
-\end{figure}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\figmed]{sphx_glr_results_006.png}
-  \caption{Candidate solution orientation MRPs for the Delta I (Star-37E) rocket body, non-candidate converged minima are plotted in grey.}
-  \label{fig:case2_p}
-\end{figure}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\figmed]{sphx_glr_results_005.png}
-  \caption{Candidate solution angular velocities for the Delta I (Star-37E) rocket body, non-candidate converged minima are plotted in grey.}
-  \label{fig:case2_w}
 \end{figure}
@@
 Inversion was performed using $10^4$ initial guesses with the material and angular velocity assumptions listed in Table \ref{tb:case3_ass}.
-The candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case3_s}, inertia ratios in Figure \ref{fig:case3_i}, orientation MRPs in Figure \ref{fig:case3_p}, and angular velocities in Figure \ref{fig:case3_w}.
+The candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case3_s}, orientation MRPs and angular velocities in Figure \ref{fig:case3_pw}, and inertia ratios in Figure \ref{fig:case3_i}.
 
@@
   \centering
-  \includegraphics[width=\figmed]{sphx_glr_results_008.png}
-  \caption{Candidate solution orientation MRPs for ECHOSTAR II, non-candidate converged minima are plotted in grey.}
-  \label{fig:case3_p}
+  \begin{subfigure}[t]{0.23\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{sphx_glr_results_008.png}
+    \caption{MRP}
+    \label{fig:case3_pwa}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.23\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{sphx_glr_results_007.png}
+    \caption{Angular velocity}
+    \label{fig:case3_pwb}
+  \end{subfigure}
+
+  \caption{Candidate solution orientation MRPs and angular velocity vectors for ECHOSTAR II}
+  \label{fig:case3_pw}
 \end{figure}
 
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\figmed]{sphx_glr_results_007.png}
-  \caption{Candidate solution angular velocities for ECHOSTAR II, non-candidate converged minima are plotted in grey.}
-  \label{fig:case3_w}
-\end{figure}
+% \begin{figure}[H]
+%   \centering
+%   \includegraphics[width=\figmed]{sphx_glr_results_008.png}
+%   \caption{Candidate solution orientation MRPs for ECHOSTAR II, non-candidate converged minima are plotted in grey.}
+%   \label{fig:case3_p}
+% \end{figure}
+
+% \begin{figure}[H]
+%   \centering
+%   \includegraphics[width=\figmed]{sphx_glr_results_007.png}
+%   \caption{Candidate solution angular velocities for ECHOSTAR II, non-candidate converged minima are plotted in grey.}
+%   \label{fig:case3_w}
+% \end{figure}
 

### 2025-03-18 paper.tex
corrected 3/18 comments

@@
 If the object's identity is known, its attitude profile can be estimated from the light curve -- up to certain ambiguities -- through a process known as light curve attitude inversion.
-Due to the environmental and instrumental noise in the light curve, as well as symmetries in the observation geometry, there may be multiple attitude profiles that are equally likely to produce any given light curve.
-To identify these ambiguous solutions, global attitude inversion methods must thoroughly search the solution space.
+Due to symmetries in the observation geometry and object shape, there may be multiple attitude profiles that are equally likely to produce any given light curve.
+Environmental and instrumental measurement noise worsens these ambiguities.
+To reliably identify these ambiguous solutions, global attitude inversion methods must thoroughly search the solution space.
 
-In this work, we apply a BFGS-based global attitude inversion algorithm to real observations taken by the Purdue Optical Ground Station of a Delta I upper stage and the ECHOSTAR II satellite.
+In this work, we apply a BFGS-based multi-start global attitude inversion algorithm to synthetic observations of an upper stage and real observations taken by the Purdue Optical Ground Station of a Delta I upper stage and the ECHOSTAR II satellite.
 Given a light curve with an estimate of the shape and reflective properties of the object, our procedure searches the space of initial orientations, angular velocities, and inertia tensors to find initial conditions that produce light curves with low error compared to the observed values.
-Our formulation inherently accounts for the multiple ambiguities resulting from the object and observation geometry, the physical constraints of torque-free rigid body motion, and noise in the light curve.
+Our formulation inherently accounts for the aforementioned ambiguities, the physical constraints of torque-free rigid body motion, and the noise in the light curve.
 
@@
 This class of methods includes single Kalman filters \cite{burton2021two, gagnon2024, wetterer2009}, and multi-filter multiple-model adaptive estimation (MMAE) schemes \cite{linares2014space, dianetti2020}.
-These approaches can be very effective if the initial state estimate is close to the truth, but have been reported to diverge for orientation errors as small as $5^\circ$ in orientation and $0.03^\circ/s^{-1}$ in angular velocity \cite{gagnon2024} due measurement ambiguities and nonlinearity.
+These approaches can be very effective if the initial state estimate is close to the truth, but have been reported to diverge even in cases when the initial guess guess is within $5^\circ$ of the ground truth orientation and $0.03^\circ/s^{-1}$ in angular velocity error \cite{gagnon2024} due to measurement nonlinearity and ambiguities.
 
-If no initial attitude state is available, as in most cases of passive debris observation, the inversion problem must be solved globally.
-Filtering approaches can be applied here too; particle \cite{linares2014particle, holzinger2014} and Bayesian Multi-Hypothesis filters \cite{burton2021two, cabrera2023} have shown promise in prior work.
-Bayesian methods may be ineffective if a reliable prior distribution for the attitude state is not available.
+If no initial attitude guess is available, as in most cases of passive debris observation, the inversion problem must be solved globally.
+Filtering approaches can be applied here too; particle \cite{linares2014particle, holzinger2014} and Bayesian Multi-Hypothesis filters \cite{burton2021two, cabrera2023} have shown promise in prior work, but still rely on high-quality priors as with the Kalman filters.
 In the absence of prior knowledge about the object's attitude state, simulated annealing \cite{gagnon2024, clark2020}, genetic algorithms \cite{gagnon2024, piergentili2017, clark2020}, and particle swarm optimizers (PSOs) \cite{clark2020, clark2022, burton2024journal, burton2024scitech, gagnon2024} can be applied to search the solution space exhaustively.
@@
 While much work has performed full state attitude inversion on simulated light curve data \cite{burton2024journal, gagnon2024, robinson2025att}, studies using real data often focus on determining a constant spin rate and axis of rotation for the observed object.
+
 Frequency analysis in the form of periodograms, Fourier transforms, or epoch folding, is often used to determine the spin rate based on the light curve's frequency spectrum \cite{silha2015, silha2021, isoletta2024, schildknecht2015, pittet2017, yanagisawa2012, koshkin2018}.
@@
 
-Works that seek to estimate a full rigid body attitude state from real data -- a time-varying orientation and angular velocity -- grid search \cite{shafer2017} and genetic algorithms \cite{piergentili2020, gallucci2020} have been used.
-These full-state estimation studies are limited by computational complexity, poor knowledge of material reflectivities, and the presence of measurement noise.
+The simplifying assumptions of constant, single-axis rotation that are commonly made when working with real data enable elegant solutions, but break down when the observed object is in a complex spin profile.
+Methods for full attitude state solutions -- time-varying orientation and angular velocities -- have been developed \cite{clark2022, burton2024journal, gagnon2024, linares2014particle}, but are seldom applied to real data.
+Works that have been successful at estimating a full rigid body attitude state from real data include genetic algorithms \cite{piergentili2020, gallucci2020} and brute-force grid search \cite{shafer2017}.
+Any full-state estimation method must tackle the inherent computational complexity, lack of knowledge of material reflectivities, and the presence of measurement noise.
 
-Our method is a global full-state attitude solver inspired by recent PSO work \cite{burton2024journal} but is easier to tune, can be fully parallelized, and is robust to significant measurement noise \cite{robinson2025att}.
+In this paper, we describe a global full-state attitude solver inspired by recent PSO work \cite{burton2024journal} but is easier to tune, can be fully parallelized, and is robust to significant measurement noise \cite{robinson2025att}.
 Further, our method differs from other global estimation methods in that we do not seek to find one optimal solution and instead search for a collection of high-quality local minima to directly study the many ambiguous solutions for a given light curve.
@@
 We present inversion results for a simulated rocket body light curve where the ground truth is known, followed by two cases -- a derelict satellite and rocket body -- using real light curve data collected by the Purdue Optical Ground Station.
-A deeper treatment of our method and additional simulated light curve results can be found in a forthcoming publication by the authors \cite{robinson2025att}.
+A deeper treatment of our method and additional simulated light curve results can be found in \cite{robinson2025att}.
 
@@
 
-The state vector $\vctr{x}$ we optimize to fit the observed light curve is the concatenation of the Modified Rodriguez Parameter (MRP) $\vctr{p} = [p_1, p_2, p_3]^T$, the principal body frame angular velocity $\vctr{\omega} = [\omega_1, \omega_2, \omega_3]$ in radians per second, and the ratios of the principal moments of inertia $J_y / J_x$ and $J_z / J_x$ where the constant inertia tensor in principal axes is $J = \mathrm{diag}\left([J_x, J_y, J_z]\right)$:
+The state vector $\vctr{x}$ is defined to be the concatenation of the Modified Rodriguez Parameter (MRP) $\vctr{p} = [p_1, p_2, p_3]^T$, the principal body frame angular velocity $\vctr{\omega} = [\omega_1, \omega_2, \omega_3]$ in radians per second, and the ratios of the principal moments of inertia $J_y / J_x$ and $J_z / J_x$ where the constant inertia tensor in principal axes is $J = \mathrm{diag}\left([J_x, J_y, J_z]\right)$:
 
@@
 
-Our inversion method is based on the global minimization of a maximum-likelihood loss function which computes the negative log-likelihood of the observed light curve $\vctr{S}$ being a sample from the normal distribution defined by the estimated light curve mean $\hat{\vctr{S}}$ and the estimated standard deviation at each $k$th timestep $\sigma_k$:
+Our inversion method is based on the global minimization of a maximum-likelihood loss function which computes the negative log-likelihood of the observed light curve $\vctr{S}$ being a sample from the normal distribution defined by the estimated light curve mean $\hat{\vctr{S}}$ and the standard deviation at each $k$th timestep $\sigma_k$:
 
@@
 
-\subsection{Running the Solver} \label{sec:run_solver}
+\subsection{Computing Local Minima} \label{sec:run_solver}
 
-To initialize the solver, we initialize $n$ guesses scattered throughout the solution space -- possibly sampled from a prior distribution based on available knowledge.
+Our multi-start solution method begins by creating $n$ guesses scattered throughout the solution space -- possibly sampled from a prior distribution based on available knowledge.
 In the most basic case, the initial state guess is sampled for guess $i$ at iteration $k=0$:
@@
 \noindent
-In this work, $\vctr{p}_{i,0}(0)3$ is uniformly randomly sampled from orientation space, $\vctr{\omega}_{i,0}(0)$ is uniformly randomly sampled from the ball in $\mathbb{R}^3$ of radius $\norm{\vctr{\omega}}_\text{max}$ -- the maximum angular rate to search -- while $\left(J_y / J_x\right)_{i,0}$ and $\left(J_z / J_x\right)_{i,0}$ are sampled from Gaussian prior distributions centered at estimated values derived from the known geometry of the object.
+In this work, $\vctr{p}_{i,0}(0)$ is uniformly randomly sampled from orientation space, $\vctr{\omega}_{i,0}(0)$ is uniformly randomly sampled from the ball in $\mathbb{R}^3$ of radius $\norm{\vctr{\omega}}_\text{max}$ -- the maximum angular rate at which to initialize guesses -- while $\left(J_y / J_x\right)_{i,0}$ and $\left(J_z / J_x\right)_{i,0}$ are sampled from Gaussian prior distributions centered at estimated values derived from the known geometry of the object.
 We determine $\norm{\vctr{\omega}}_\text{max}$ based on the maximum power positive frequency in the observed light curve $f^\star$ in Hertz and an arbitrary safety factor $f_\omega \geq 1$, selected as $f_\omega = 1.25$ for this work:
@@
 \end{equation}
+
+Even for a well-chosen bound $\norm{\vctr{\omega}}_\text{max}$, there may still be numerous high-quality solutions at significantly lower or higher angular velocities that possess equal $f^\star$ values due to object shape symmetries or harmonics in spin and precession rates, respectively.
+As a result, it is usually unwise to directly constrain the angular velocity to a given magnitude unless additional, unambiguous information is available.
 
@@
 
- In any particular illumination and observation condition, a fraction of the facet's area $\tilde{a}_i$ may be shadowed due to obstructing geometry before or after light reaches the surface element.
+ In any particular illumination and observation condition, a fraction of the facet's area $\tilde{a}_i$ may be Sun- or observer-shadowed due to obstructing geometry before or after light reaches the surface element, respectively.
 The remainder of the facet's area $\bar{a}_i$ is unshadowed and contributes to the total flux reaching the observer from the object, such that the total area of each facet is $a_i = \tilde{a}_i + \bar{a}_i$.
@@
 Self-shadowing can be efficiently solved semi-analytically for lower-fidelity approximations of real space object geometry by computing the mutual intersections of the polygons $P_k$ of other faces whose projections overlap with the face in question.
-Up to a maximum intersection depth $d$, the unshadowed area can be computed via:
+Up to a maximum intersection depth $d$, selected for computation time, the unshadowed area can be computed via:
 
@@
 As the objects we investigate in this work are well-approximated with relatively simple non-convex meshes, we use the semi-analytic method to efficiently compute unshadowed areas via the Sutherland-Hodgeman algorithm \cite{sutherland1974} for each polygon intersection.
+Further explanation of the polygon clipping procedure for computing the total shadowed area and accelerating this computation during inversion with the so-called shadow cache is included here \textbf{TODO: should this be a here?
+wording this felt weird if we're avoiding saying like "the authors'"} \cite{robinson2025att}.
 
@@
 
-This section summarizes an in-depth study of relevant noise sources in CCD and CMOS sensor light curves by the authors in a forthcoming publication \cite{robinson2025twin}.
+This section summarizes an in-depth study of relevant noise sources in CCD and CMOS sensor light curves which can be found in \cite{robinson2025twin}.
 
@@
   \centering
-  \caption{Reflection properties of materials used in the synthetic data test case. The authors estimated the white paint parameters qualitatively}
+  \caption{Reflection properties of materials used in the synthetic data test case. The white paint parameters were estimated qualitatively}
   \vspace*{6pt}
@@
   \includegraphics[width=\figmed]{light_curves_mag_rocket_body_sdc.png}
-  \caption{Simulated observed V-band magnitude for the rocket body.}
+  \caption{Simulated mean and sampled observations for the rocket body in V-band magnitude, with the variance included in blue.}
   \label{fig:obs_mag_synth}
@@
 The inertia tensor ratios are initialized by sampling a symmetric Gaussian distribution with $\sigma=0.05 \: [kg\cdot m^2]$ on each axis to simulate a moderate level of inertia uncertainty.
-BFGS minimization was then performed for each initial guess to locally solve the optimization problem given by Equation \ref{eq:opt_problem}.
+BFGS minimization is then performed for each initial guess to locally solve the optimization problem given by Equation \ref{eq:opt_problem}.
 Classifying candidate solutions with $\ell = 1/10$ produces over $10^4$ solutions within this bound.
 Figure \ref{fig:conv_lcs_synth} shows the mean light curves produced by the identified candidate solutions.
-There is generally good agreement between the candidate light curves and the observations, indicating that high-quality local minima have been identified.
+There is good agreement between the candidate light curves and the observations, indicating that high-quality local minima have been identified.
 
@@
 Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} display the candidate MRP and angular velocity vectors of the best solutions determined via the procedure in Section \ref{sec:candidate_sols}.
-We notice that for both the initial orientation and angular velocity, the ground truth lies squarely in the surfaces of candidate solutions.
-In angular velocity space, most solutions fall onto the surface of two open-ended cylinders -- one with half the radius of the larger cylinder containing the truth.
-Similarly in orientation space, we solutions fall on two distinct manifolds -- likely resulting from a reflection of the object's body frame about the bisector between the Sun and observer vectors \cite{marto2024, burton2024journal}.
+We notice that for both the initial orientation in Figure \ref{fig:sigma_sols1} and angular velocity in Figure \ref{fig:omega_sols1}, the ground truth, plotted in red, lies squarely on a candidate solution surface.
+In angular velocity space, shown in Figure \ref{fig:omega_sols1}, most solutions fall onto the surface of two open-ended cylinders -- one with half the radius of the larger cylinder containing the truth.
+Similarly in orientation space, shown in Figure \ref{fig:sigma_sols1}, the solutions fall on two distinct manifolds -- likely resulting from a reflection of the object's body frame about the bisector between the Sun and observer vectors \cite{marto2024, burton2024journal}.
 One of these orientation-space surfaces contains the ground truth.
-These additional families of solutions highlight the reality that a wide variety of attitude profiles may produce the same observations.
+These additional families of solutions highlight the reality that a wide variety of attitude profiles may produce nearly identical observations.
 
@@
 
-This light curve shows clear periodicity, but has a high estimated standard deviation and a relatively low sampling rate due to the object's low brightness and the sensor readout time.
+This light curve shows clear periodicity, but has a high measurement standard deviation and a relatively low sampling rate due to the object's low brightness and the sensor readout time.
 Inversion was performed using $10^4$ initial guesses with the material and angular velocity assumptions listed in Table \ref{tb:case2_ass}.
@@
 If a shape model is known, the light curve -- a sequence of these brightness measurements -- is one tool for attitude estimation.
-We presented an approach for solving this highly nonlinear and non-convex optimization problem with a gradient-based multistart method.
+We presented an approach for solving this highly nonlinear and non-convex optimization problem with a gradient-based multi-start method.
 We presented attitude inversion results for one case with synthetic data and two cases with real observations.

### 2025-03-20 paper.tex
added discussion of new results plots

@@
 
+\section{Global TODO list}
+
+\begin{itemize}
+  \item Add the constant term back into the objective function (in text and code) so we can start treating it as a real likelihood, useful for the future
+  \item Mention the unknown solar panel rotation angle for ECHOSTAR
+  \item Update perspective on some the case 1a angular velocity plot
+  \item Update $n_\text{guess}$ values everywhere to reflect the scripts
+  \item Update $\omega$ sampling scheme to go from $0.5-2$ as reflected in scripts
+  \item Re-run all cases to ensure that values are up to date with recent changes, update where those statistics are reported in the text
+  \item Try to get rid of big gaps where possible
+\end{itemize}
+
 \section{Introduction}
@@
 Filtering approaches can be applied here too; particle \cite{linares2014particle, holzinger2014} and Bayesian Multi-Hypothesis filters \cite{burton2021two, cabrera2023} have shown promise in prior work, but still rely on high-quality priors as with the Kalman filters.
-In the absence of prior knowledge about the object's attitude state, simulated annealing \cite{gagnon2024, clark2020}, genetic algorithms \cite{gagnon2024, piergentili2017, clark2020}, and particle swarm optimizers (PSOs) \cite{clark2020, clark2022, burton2024journal, burton2024scitech, gagnon2024} can be applied to search the solution space exhaustively.
+In the absence of prior knowledge about the object's attitude state, simulated annealing \cite{gagnon2024, clark2020}, genetic algorithms \cite{gagnon2024, piergentili2017, clark2020}, and particle swarm optimizers (PSOs) \cite{clark2020, clark2022, burton2024journal, burton2024scitech, gagnon2024, gagnon2025} can be applied to search the solution space exhaustively.
 
@@
 Further, our method differs from other global estimation methods in that we do not seek to find one optimal solution and instead search for a collection of high-quality local minima to directly study the many ambiguous solutions for a given light curve.
-Our approach was designed specifically for use with real data as we take time-varying measurement noise and uncertainties in the body moments of inertia into account explicitly.
+Our approach is designed specifically for use with real data as we take time-varying measurement noise and uncertainties in the body moments of inertia into account explicitly.
 
@@
 
-Our multi-start solution method begins by creating $n$ guesses scattered throughout the solution space -- possibly sampled from a prior distribution based on available knowledge.
+Our multi-start solution method begins by creating $n_\text{guess}$ guesses scattered throughout the solution space -- possibly sampled from a prior distribution based on available knowledge.
 In the most basic case, the initial state guess is sampled for guess $i$ at iteration $k=0$:
@@
 \begin{equation}
- \vctr{x}_{0,i}(0) = \begin{bmatrix}\vctr{p}_{i,0}(0) & \vctr{\omega}_{i,0}(0) & \left(J_y / J_x\right)_{i,0} & \left(J_z / J_x\right)_{i,0}\end{bmatrix}^T.
+ \vctr{x}_{0,i}(0) = \begin{bmatrix}\vctr{p}_{i,0}(0) \\ \vctr{\omega}_{i,0}(0) \\ \left(J_y / J_x\right)_{i,0} \\ \left(J_z / J_x\right)_{i,0}\end{bmatrix}.
 \end{equation}
 
-\noindent
-In this work, $\vctr{p}_{i,0}(0)$ is uniformly randomly sampled from orientation space, $\vctr{\omega}_{i,0}(0)$ is uniformly randomly sampled from the ball in $\mathbb{R}^3$ of radius $\norm{\vctr{\omega}}_\text{max}$ -- the maximum angular rate at which to initialize guesses -- while $\left(J_y / J_x\right)_{i,0}$ and $\left(J_z / J_x\right)_{i,0}$ are sampled from Gaussian prior distributions centered at estimated values derived from the known geometry of the object.
+In this work, initial guesses for the state vector are sampled from a prior distribution, usually defined independently for the MRP, angular velocity, and inertia tensor ratios.
+If no information is available, a uniform distribution is sampled from for the first six components of the state.
+In particular, $\vctr{p}_{i,0}(0)$ is uniformly randomly sampled from orientation space, $\vctr{\omega}_{i,0}(0)$ is randomly sampled from a uniform distribution of magnitudes up to $\norm{\vctr{\omega}}_\text{max}$ -- the maximum angular rate at which to initialize guesses.
 We determine $\norm{\vctr{\omega}}_\text{max}$ based on the maximum power positive frequency in the observed light curve $f^\star$ in Hertz and an arbitrary safety factor $f_\omega \geq 1$, selected as $f_\omega = 1.25$ for this work:
@@
 
+Uniformly sampling over angular velocity directions, we can express the default sampling strategy for an angular velocity guess as:
+
+\begin{equation} \label{eq:omega_sampler}
+  \vctr{\omega}_{i,0}(0) = \frac{m}{\sqrt{g_1^2+g_2^2+g_3^2}} \cdot \begin{bmatrix} g_1 & g_2 & g_3 \end{bmatrix}^T,
+\end{equation}
+
+\noindent
+where $g_{1:3} \in \mathcal{N}(0,1)$ and $m=\mathcal{U}(0,\norm{\vctr{\omega}}_\text{max})$.
+
 Even for a well-chosen bound $\norm{\vctr{\omega}}_\text{max}$, there may still be numerous high-quality solutions at significantly lower or higher angular velocities that possess equal $f^\star$ values due to object shape symmetries or harmonics in spin and precession rates, respectively.
 As a result, it is usually unwise to directly constrain the angular velocity to a given magnitude unless additional, unambiguous information is available.
+
+Similarly for the MRP, we first sample a quaternion $\vctr{q}_i(0) = [q_x, q_y, q_z, q_w]^T$ uniformly over orientation space:
+
+\begin{equation}
+  \vctr{q}_i(0) = \frac{1}{\sqrt{g_1^2 + g_2^2 + g_3^2 + g_4^2}} \begin{bmatrix} g_1 & g_2 & g_3 & g_4 \end{bmatrix}^T,
+\end{equation}
+
+\noindent
+and project the quaternion into MRP space:
+
+\begin{equation} \label{eq:mrp_sampler}
+  \vctr{p}_{i,0}(0) = \frac{1}{1 + q_w} \begin{bmatrix} q_x & q_y & q_z \end{bmatrix}^T.
+\end{equation}
+
+Unlike the MRP and angular velocity, the default sampling method for inertia ratios initial guesses $\left(J_y / J_x\right)_{i,0}$ and $\left(J_z / J_x\right)_{i,0}$ draws from independent, identically distributed Gaussian prior distributions with $\sigma_J$ centered at estimated values derived from the known geometry of the object $\left(J_y / J_x\right)_\text{est}$ and $\left(J_z / J_x\right)_\text{est}$, such that:
+
+\begin{equation}
+  \begin{bmatrix}
+    \left(J_y / J_x\right)_{i,0} \\ \left(J_z / J_x\right)_{i,0}
+  \end{bmatrix} \sim \mathcal{N} \left(\begin{bmatrix}
+    \left(J_y / J_x\right)_\text{est} \\
+    \left(J_z / J_x\right)_\text{est} \end{bmatrix}, \sigma_J I_2 \right),
+\end{equation}
+
+\noindent
+where $I_2$ is the $2\times2$ identity matrix.
 
@@
 \noindent
-producing a final initial state which is a good local guess for a state that produces the observations.
+producing a converged initial state $\vctr{x}_{k,i}(0)$ which is a good local guess to have produced the observations.
 In this way, we globally search the solution space up to the density implied by $n$.
@@
 
-\begin{equation}
-  e^{-f(\vctr(x)_{i,k}(0))} > \ell e^{-f(\vctr(x)^\star_{k}(0))}.
+\begin{equation} \label{eq:ell_selection_criteria}
+  e^{-f\left(\vctr{x}_{i,k}(0)\right)} > \ell e^{-f\left(\vctr{x}^\star_{k}(0)\right)}.
 \end{equation}
 
-Here, $\vctr(x)^\star_{k}(0) = \argmin_{\vctr{x}_{i,k}(0)} f(\vctr(x)_{i,k}(0)) $ is the highest likelihood solution found after optimization.
+Here, $\vctr{x}^\star_{k}(0) = \argmin_{\vctr{x}_{i,k}(0)} f\left(\vctr{x}_{i,k}(0)\right) $ is the highest likelihood solution found after optimization.
 
 \section{Computing the Light Curve}
+
+The inversion method in this paper is not dependent on a particular approach for computing light curves and their expected standard deviations.
+Here, we present the models used to produce the results in Sections \ref{sec:synth_results} and \ref{sec:real_results}.
 
@@
 As the objects we investigate in this work are well-approximated with relatively simple non-convex meshes, we use the semi-analytic method to efficiently compute unshadowed areas via the Sutherland-Hodgeman algorithm \cite{sutherland1974} for each polygon intersection.
-Further explanation of the polygon clipping procedure for computing the total shadowed area and accelerating this computation during inversion with the so-called shadow cache is included here \textbf{TODO: should this be a here?
-wording this felt weird if we're avoiding saying like "the authors'"} \cite{robinson2025att}.
+Further explanation of the polygon clipping procedure for computing the total shadowed area and accelerating this computation during inversion with the so-called shadow cache is included here \textbf{TODO: should this be a `here'?
+wording this felt weird if we're avoiding saying like `the authors'"} \cite{robinson2025att}.
 
@@
 
-\section{Attitude Inversion Results for Synthetic Data}
+\section{Attitude Inversion Results for Synthetic Data} \label{sec:synth_results}
+
+To demonstrate the effectiveness of our inversion method, we first present results for a simulated light curve where the ground truth attitude profile is available for comparison.
 
@@
 
-To demonstrate the effectiveness of our inversion method, we first present results for a simulated light curve where the ground truth is known.
 In order to make this simulated case realistic, we use the presented measurement noise models and introduce a mismatch between the ground truth object used to generate the measurements and the model used in inversion.
@@
   \centering
-  \caption{Reflection properties of materials used in the synthetic data test case. The white paint parameters were estimated qualitatively}
+  \caption{Reflection properties of materials used in the synthetic data test case. The white paint parameters are estimated qualitatively}
   \vspace*{6pt}
@@
  Initial $\vctr{p}(0)$ & $-\frac{1}{3} \begin{bmatrix} 1 & 1 & 1 \end{bmatrix}^T$ \\ \hline
- True $\vctr{\omega}(0)$ ($\text{rad}/\text{s}$) & $\vctr{\omega}(0) = \begin{bmatrix} 0.03 & 0.06 & 0.03 \end{bmatrix}^T$ \\ \hline
+ True $\vctr{\omega}(0)$ ($\text{rad}\cdot\text{s}^{-1}$) & $\vctr{\omega}(0) = \begin{bmatrix} 0.03 & 0.06 & 0.03 \end{bmatrix}^T$ \\ \hline
  Inertia tensor ratios & $J_y / J_x = 1, \: J_z / J_x = 0.25$ \\ \hline
@@
   \includegraphics[width=\figmed]{light_curves_mag_rocket_body_sdc.png}
-  \caption{Simulated mean and sampled observations for the rocket body in V-band magnitude, with the variance included in blue.}
+  \caption{Simulated mean and sampled observations for the rocket body in black in V-band magnitude, with the variance included in blue.}
   \label{fig:obs_mag_synth}
@@
 
-\subsection{Synthetic Light Curve Optimization Results}
+\subsection{Case 1a and 1b: Synthetic Light Curve Optimization Results}
 
@@
 
-A total of $10^5$ initial state guesses are created via the sampling scheme described in Section \ref{sec:run_solver} with $\norm{\vctr{\omega}}_\text{max}=5.26$ degrees per second.
-The inertia tensor ratios are initialized by sampling a symmetric Gaussian distribution with $\sigma=0.05 \: [kg\cdot m^2]$ on each axis to simulate a moderate level of inertia uncertainty.
+Two distinct inversion cases are computed.
+Case 1a presents inversion results given no attitude information besides an estimate of the inertia tensor ratios derived from the known geometry, while Case 1b presents results given good initial guesses for orientation and angular velocity.
+Tables \ref{tb:case1a_prior} and \ref{tb:case1b_prior} summarize the initial condition sampling strategies employed for each case.
+
+For both cases, a total of $n_\text{guess}=10^5$ initial state guesses are created via the sampling schemes described their respective tables.
 BFGS minimization is then performed for each initial guess to locally solve the optimization problem given by Equation \ref{eq:opt_problem}.
-Classifying candidate solutions with $\ell = 1/10$ produces over $10^4$ solutions within this bound.
+
+\begin{table}[H]
+  \centering
+  \caption{Initial condition sampling parameters used for inversion in Case 1a}
+  \vspace*{6pt}
+  \begin{tabular}{|l|l|}
+  \hline
+  \textbf{Variable} & \textbf{Value} \\ \hline
+  $\vctr{p}$ sampling & Uniform, via Equation \ref{eq:mrp_sampler} \\ \hline
+  $\vctr{\omega}$ sampling & Uniform, via Equation \ref{eq:omega_sampler} \\ \hline
+  $J$ sampling & Gaussian, via Equation \ref{eq:omega_sampler} \\ \hline
+  $\sigma_J$ & $0.05$ $[\text{nondim}]$ \\ \hline
+  $\norm{\vctr{\omega}}_\text{max}$ & $5.26$ $[\text{deg}\cdot \text{s}^{-1}]$ \\ \hline
+\end{tabular}
+  \label{tb:case1a_prior}
+\end{table}
+
+\begin{table}[H]
+  \centering
+  \caption{Informative prior parameters used for inversion in Case 1b.
+The standard deviations $\sigma_\omega$ and $\sigma_J$ are for spherically-symmetric Gaussians with centered on the ground truth.
+MRPs are sampled by rotating by $\theta\sim \mathcal{N}(0,\sigma_p)$ degrees away from the ground truth orientation around an axis uniformly sampled on the sphere.}
+  \vspace*{6pt}
+  \begin{tabular}{|l|l|}
+  \hline
+  \textbf{Variable} & \textbf{Value} \\ \hline
+  $\vctr{p}$ sampling & Gaussian, $\sigma_\omega = 0.05$ $[\text{deg}\cdot\text{s}^{-1}]$ \\ \hline
+  $\vctr{\omega}$ sampling & Gaussian, $\sigma_p=20$ $[\text{deg}]$ \\ \hline
+  $J$ sampling & Gaussian, $\sigma_J = 0.05$ $[\text{nondim}]$ \\ \hline
+\end{tabular}
+  \label{tb:case1b_prior}
+\end{table}
+
+Solutions are then classified as candidates for the ground truth with $\ell = 1/10$ via Equation \ref{eq:ell_selection_criteria}.
 Figure \ref{fig:conv_lcs_synth} shows the mean light curves produced by the identified candidate solutions.
@@
   \centering
-  \includegraphics[width=\figmed]{converged_solutions_1_Rocket Body.png}
+  \includegraphics[width=\figbig]{converged_solutions_1_Rocket Body.png}
   \caption{Candidate solution light curves found through BFGS optimization.}
@@
 
-Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} display the candidate MRP and angular velocity vectors of the best solutions determined via the procedure in Section \ref{sec:candidate_sols}.
-We notice that for both the initial orientation in Figure \ref{fig:sigma_sols1} and angular velocity in Figure \ref{fig:omega_sols1}, the ground truth, plotted in red, lies squarely on a candidate solution surface.
+Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} display the candidate MRP and angular velocity of the candidate solutions determined via the procedure in Section \ref{sec:candidate_sols}.
+In the left Case 1a figures for orientation and angular velocity, we notice that the ground truth lies squarely on a surface containing numerous candidate solution.
+This indicates that the local minima have correctly converged to the non-singular subspace of states that approximate the observations well for the assumed shape model.
 In angular velocity space, shown in Figure \ref{fig:omega_sols1}, most solutions fall onto the surface of two open-ended cylinders -- one with half the radius of the larger cylinder containing the truth.
-Similarly in orientation space, shown in Figure \ref{fig:sigma_sols1}, the solutions fall on two distinct manifolds -- likely resulting from a reflection of the object's body frame about the bisector between the Sun and observer vectors \cite{marto2024, burton2024journal}.
-One of these orientation-space surfaces contains the ground truth.
+Similarly in orientation space, shown in Figure \ref{fig:sigma_sols1}, the solutions fall on two distinct manifolds, one containing the truth -- likely resulting from a reflection of the object's body frame about the bisector between the Sun and observer vectors \cite{marto2024, burton2024journal}.
 These additional families of solutions highlight the reality that a wide variety of attitude profiles may produce nearly identical observations.
+
+The Case 1b results on the right within Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} show that the majority of the initial guesses sampled close to the truth have converged to a local subset of the solution manifolds identified by Case 1a.
+We should not expect these solutions to have shrunk down exponentially towards the ground truth -- the mismatch in the assumed shape model and the measurement noise inherently limits the minimum size of the feasible solution space.
 
@@
     \includegraphics[width=\textwidth]{sigma_solsRocket Body.png}
-    \caption{Uniform prior}
+    \caption{}
     \label{fig:sigma_sols1a}
@@
     \includegraphics[width=\textwidth]{sigma_solsRocket Body Prior.png}
-    \caption{Informative prior}
+    \caption{}
     \label{fig:sigma_sols1b}
@@
 
-  \caption{Candidate solution initial orientation MRP vectors, ground truth highlighted in red.}
+  \caption{Candidate solution initial orientation MRP vectors for Cases 1a (left) and 1b (right), with the ground truth highlighted in red.}
   \label{fig:sigma_sols1}
@@
     \includegraphics[width=\textwidth]{omega_sols_Rocket Body.png}
-    \caption{Uniform prior}
+    \caption{}
     \label{fig:omega_sols1a}
@@
     \includegraphics[width=\textwidth]{omega_sols_Rocket Body Prior.png}
-    \caption{Informative prior}
+    \caption{}
     \label{fig:omega_sols1b}
   \end{subfigure}
-
-  \caption{Candidate solution initial body-frame angular velocity vectors, ground truth highlighted in red. TODO: change the perspective on the angular velocities}
+  \caption{Candidate solution initial body-frame angular velocity vectors for Cases 1a (left) and 1b (right), with the ground truth highlighted in red.
+TODO: change the perspective on the angular velocities}
   \label{fig:omega_sols1}
 \end{figure}
+
+Figure \ref{fig:i_sols1} displays the inertia tensor solutions for Cases 1a and 1b.
+Notably, the majority of candidate solutions remain within a few standard deviations of the mean of their initial distribution, with the remaining minority exploring directions that maintain the same precession rate.
+When a good initial guess for the other attitude state components is available in Case 1b, the inertia tensor solutions are more well-clustered, although due to the low observability of the rocket body's spin rate in the observed light curve, there are still ambiguous solutions with higher or lower $J_z / J_x$ values.
 
@@
     \includegraphics[width=\textwidth]{converged_solutions_2_Rocket Body.png}
-    \caption{Uniform prior}
+    \caption{}
     \label{fig:i_sols1a}
@@
     \includegraphics[width=\textwidth]{converged_solutions_2_Rocket Body Prior.png}
-    \caption{Informative prior}
+    \caption{}
     \label{fig:i_sols1b}
   \end{subfigure}
-
-  \caption{Candidate solution inertia ratios, ground truth highlighted in red.}
+  \caption{Candidate solution inertia ratios for Cases 1a (left) and 1b (right), with the ground truth highlighted in red.}
   \label{fig:i_sols1}
 \end{figure}
+
+Figure \ref{fig:w_vs_ang_error_sols1} summarizes the distribution of candidate solutions with respect to the ground truth by displaying the angular velocity magnitude error and angular orientation error compared to the ground truth.
+Case 1a shows significant variability in angular velocity magnitude, although the mean error is low at $2\%$.
+Due to the underdetermined orientation solution space, there are far more solutions at high angular offsets from the truth orientation, leading to a large mean angular error of $127^\circ$.
+For Case 1b, the mean angular error is much lower at $25^\circ$, although convergence here is still inherently limited by the noisy measurements and model mismatch.
 
@@
     \includegraphics[width=\textwidth]{converged_solutions_3_Rocket Body.png}
-    \caption{Uniform prior}
+    \caption{}
     \label{fig:w_vs_ang_error_sols1a}
@@
     \includegraphics[width=\textwidth]{converged_solutions_3_Rocket Body Prior.png}
-    \caption{Informative prior}
+    \caption{}
     \label{fig:w_vs_ang_error_sols1b}
@@
 
-  \caption{Candidate solution angular orientation error and angular velocity magnitudes compared to the truth}
+  \caption{Candidate solution angular orientation error and angular velocity magnitudes for Cases 1a (left) and 1b (right) compared to the ground truth}
   \label{fig:w_vs_ang_error_sols1}
@@
 
-% The inversion
-
-\section{Attitude Inversion Results for Real Data}
+\section{Attitude Inversion Results for Real Data} \label{sec:real_results}
 
@@
 
-\subsection{Delta I Upper Stage}
+\subsection{Case 2: Delta I Upper Stage}
 
-The Delta I upper stage (COSPAR ID 1984-115C) is a Star-37E \cite{delta3914_astronautix} solid fuel apogee kick motor, displayed in Figure \ref{fig:star37e}.
+In Case 2, we perform attitude inversion on real data observed for a Delta I upper stage rocket body.
+The chosen Delta I upper stage (COSPAR ID 1984-115C) is a Star-37E \cite{delta3914_astronautix} solid fuel apogee kick motor, displayed in Figure \ref{fig:star37e}.
 It is $d=0.93$ meters in diameter and is approximately $h=1.7$ meters in length \cite{star37e_astronautix, star37_gunter}.
@@
   \centering
-  \includegraphics[width=\figsmall]{star37e.JPG}
+  \includegraphics[width=0.25\textwidth]{star37e.JPG}
   \caption{Star-37E upper stage \cite{star37_af}.}
@@
 This light curve shows clear periodicity, but has a high measurement standard deviation and a relatively low sampling rate due to the object's low brightness and the sensor readout time.
-Inversion was performed using $10^4$ initial guesses with the material and angular velocity assumptions listed in Table \ref{tb:case2_ass}.
+Inversion is performed using $n_\text{guess}=10^4$ initial guesses with the material and angular velocity assumptions listed in Table \ref{tb:case2_ass}.
+As in Case 1a, orientation and angular velocity initial guesses are sampled via Equations \ref{eq:mrp_sampler} and \ref{eq:omega_sampler}, respectively.
+All initial inertia tensor guesses are set to the estimated $J_\text{R/B}$.
 The candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case2_s}, orientation MRPs and angular velocities in Figure \ref{fig:case2_pw}, and inertia ratios in Figure \ref{fig:case2_i}.
@@
  Uniform BRDF & $C_d=0.2$, $C_s=0.8$, $n=30$ \\ \hline
- $\norm{\vctr{\omega}}_\text{max}$ & $8.31$ $\text{deg} / \text{s}^{-1}$ \\ \hline
+ $\norm{\vctr{\omega}}_\text{max}$ & $8.31$ $\text{deg} \cdot \text{s}^{-1}$ \\ \hline
  $\ell$ & $1/100$ \\ \hline
@@
 
-\begin{figure*}[t]
+\begin{figure}[H]
   \centering
-  \includegraphics[width=0.8\textwidth]{sphx_glr_results_002_2_00x.png}
+  \includegraphics[width=\figbig]{sphx_glr_results_002_2_00x.png}
   \caption{Candidate solution light curves compared to the real measurements in ADU for the Delta I (Star-37E) rocket body.}
   \label{fig:case2_s}
-\end{figure*}
+\end{figure}
 
@@
     \includegraphics[width=\textwidth]{sphx_glr_results_006.png}
-    \caption{MRP}
+    \caption{}
     \label{fig:case2_pwa}
@@
     \includegraphics[width=\textwidth]{sphx_glr_results_005.png}
-    \caption{Angular velocity}
+    \caption{}
     \label{fig:case2_pwb}
@@
 
-  \caption{Candidate solution orientation MRPs and angular velocity vectors for the Delta I (Star-37E) rocket body}
+  \caption{Candidate solution orientation MRPs (left) and angular velocity vectors (right) for the Delta I (Star-37E) rocket body}
   \label{fig:case2_pw}
@@
   \centering
-  \includegraphics[width=\figmed]{sphx_glr_results_001_2_00x.png}
+  \includegraphics[width=\figsmall]{sphx_glr_results_001_2_00x.png}
   \caption{Candidate solution inertia ratio distribution for the Delta I (Star-37E) rocket body}
@@
 The candidates show some rough clustering in both orientation and angular velocity space, although they vary significantly in angular velocity magnitude.
-The inertia tensor solutions display similar variability, although the solutions are spread relatively uniformly from the initial guess, possibly indicating that it was reasonably unbiased.
-Taken together, these results indicate that the object could be spinning in a number of attitude profiles, and that more information is needed about its true mass distribution and reflectivity to further refine the solution set.
+The inertia tensor solutions display similar variability, although the solutions are spread relatively uniformly from the initial guess, possibly indicating that the geometry-driven estimate is approximately unbiased.
+Taken together, these results indicate that the object could be spinning in a number of attitude profiles, although the light curve inversion process has significantly narrowed the search space to small subsets of the original unconstrained eight-dimensional volume.
+These constraints may be combined with additional light curves, more information about the object's true mass distribution, or reflectivity to further refine the solution set.
 
-\subsection{ECHOSTAR II}
+\subsection{Case 3: ECHOSTAR II}
 
+In Case 3, we perform attitude inversion on real data observed for a derelict satellite.
 ECHOSTAR II (COSPAR ID 1996-055A) was a communications satellite on the AS-7000 bus \cite{as7000_astronautix}, shown in Figure \ref{fig:echostar1}.
@@
 This light curve is sampled more densely than the Delta I observations, and clearly has significantly more intense specular peaks.
-Inversion was performed using $10^4$ initial guesses with the material and angular velocity assumptions listed in Table \ref{tb:case3_ass}.
+Inversion is performed using $n_\text{guess}=10^4$ initial guesses with the material and angular velocity assumptions listed in Table \ref{tb:case3_ass}.
+Initial guess sampling is accomplished identically to Case 2, with a different maximum angular veloity and initial inertia ratios.
 The candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case3_s}, orientation MRPs and angular velocities in Figure \ref{fig:case3_pw}, and inertia ratios in Figure \ref{fig:case3_i}.
+Since the light curve spans three orders of magnitude with a high signal-to-noise ratio, the shape and material mismatch produces specular peaks that are good fits, but are possibly hundreds of standard deviations away from the computed measurement distribution.
+To combat this, $\ell$ is selected to be much more permissive in selecting candidate solutions compared to previous cases.
 
@@
   \hline
-  \textbf{Material} & \textbf{Coefficients} \\ \hline
+  \textbf{Variable} & \textbf{Value} \\ \hline
  Aluminum BRDF & $C_d=0.22$, $C_s=0.4$, $n=5$ \\ \hline
@@
 
-\begin{figure*}[t]
+\begin{figure}[H]
   \centering
-  \includegraphics[width=0.8\textwidth]{sphx_glr_results_004_2_00x.png}
-  \caption{Candidate solution inertia light curves compared to the real measurements in ADU for ECHOSTAR II.}
+  \includegraphics[width=\figbig]{sphx_glr_results_004_2_00x.png}
+  \caption{Candidate solution light curves compared to the real measurements in ADU for ECHOSTAR II.}
   \label{fig:case3_s}
-\end{figure*}
+\end{figure}
 
@@
   \centering
-  \includegraphics[width=\figmed]{sphx_glr_results_003_2_00x.png}
-  \caption{Candidate solution inertia ratio distribution for ECHOSTAR II, non-candidate converged minima are plotted in grey.}
+  \includegraphics[width=\figsmall]{sphx_glr_results_003_2_00x.png}
+  \caption{Candidate solution inertia ratio distribution for ECHOSTAR II with initial estimate in red.}
   \label{fig:case3_i}
@@
     \includegraphics[width=\textwidth]{sphx_glr_results_008.png}
-    \caption{MRP}
+    \caption{}
     \label{fig:case3_pwa}
@@
     \includegraphics[width=\textwidth]{sphx_glr_results_007.png}
-    \caption{Angular velocity}
+    \caption{}
     \label{fig:case3_pwb}
@@
 
-  \caption{Candidate solution orientation MRPs and angular velocity vectors for ECHOSTAR II}
+  \caption{Candidate solution orientation MRPs (left) and angular velocity vectors (right) for ECHOSTAR II}
   \label{fig:case3_pw}
@@
 %   \includegraphics[width=\figmed]{sphx_glr_results_008.png}
-%   \caption{Candidate solution orientation MRPs for ECHOSTAR II, non-candidate converged minima are plotted in grey.}
+%   \caption{Candidate solution orientation MRPs for ECHOSTAR II}
 %   \label{fig:case3_p}
@@
 %   \includegraphics[width=\figmed]{sphx_glr_results_007.png}
-%   \caption{Candidate solution angular velocities for ECHOSTAR II, non-candidate converged minima are plotted in grey.}
+%   \caption{Candidate solution angular velocities for ECHOSTAR II}
 %   \label{fig:case3_w}
@@
 The inertia tensor solutions are dispersed throughout the solution space, with less symmetry than the Delta I solutions.
-As in the Delta I case, more data about the mass and reflectivity of the object would be necessary to continue refining any of these candidates.
+As in the Delta I case, the attitude inversion solver has successfully constrained the search space for later optimization as more data about the mass and reflectivity of the object becomes available to continue refining any of these candidates.
 

### 2025-03-24 paper.tex
almost done

@@
 
+\usepackage{colortbl}
+\usepackage{booktabs}
 \usepackage{subcaption}
@@
 \usepackage{adjustbox}
+
+\newcommand{\grule}[0]{\arrayrulecolor{darkgray}\cmidrule(){1-2}}
+\newcommand{\brule}[0]{\arrayrulecolor{black} \bottomrule}
 
@@
 
-In this work, we apply a BFGS-based multi-start global attitude inversion algorithm to synthetic observations of an upper stage and real observations taken by the Purdue Optical Ground Station of a Delta I upper stage and the ECHOSTAR II satellite.
+In this work, we apply a BFGS-based multi-start global attitude inversion algorithm to efficiently solve the attitude inversion problem globally.
+Results for synthetic observations are presented for a generic upper stage.
+Results for real observations taken by the Purdue Optical Ground Station of Delta I upper stage and the ECHOSTAR II satellite are also presented.
 Given a light curve with an estimate of the shape and reflective properties of the object, our procedure searches the space of initial orientations, angular velocities, and inertia tensors to find initial conditions that produce light curves with low error compared to the observed values.
@@
 
-\section{Global TODO list}
-
-\begin{itemize}
-  \item Add the constant term back into the objective function (in text and code) so we can start treating it as a real likelihood, useful for the future
-  \item Mention the unknown solar panel rotation angle for ECHOSTAR
-  \item Update perspective on some the case 1a angular velocity plot
-  \item Update $n_\text{guess}$ values everywhere to reflect the scripts
-  \item Update $\omega$ sampling scheme to go from $0.5-2$ as reflected in scripts
-  \item Re-run all cases to ensure that values are up to date with recent changes, update where those statistics are reported in the text
-  \item Try to get rid of big gaps where possible
-\end{itemize}
-
 \section{Introduction}
@@
 Methods for estimating the attitude state of uncontrolled space objects from light curves differ significantly depending on the presence of an initial guess.
-If a guess is available, many filter-based approaches have been developed to process new photometry to update the guess orientation and angular velocity.
+If prior knowledge is available, many filter-based approaches have been developed to process new photometry to update the guess orientation and angular velocity.
 This class of methods includes single Kalman filters \cite{burton2021two, gagnon2024, wetterer2009}, and multi-filter multiple-model adaptive estimation (MMAE) schemes \cite{linares2014space, dianetti2020}.
-These approaches can be very effective if the initial state estimate is close to the truth, but have been reported to diverge even in cases when the initial guess guess is within $5^\circ$ of the ground truth orientation and $0.03^\circ/s^{-1}$ in angular velocity error \cite{gagnon2024} due to measurement nonlinearity and ambiguities.
+These approaches can be very effective if the initial state estimate is close to the truth, but have been reported to diverge even in cases when the initial guess is within $5^\circ$ of the ground truth orientation and $0.03^\circ/s^{-1}$ in angular velocity error \cite{gagnon2024} due to measurement nonlinearity and ambiguities.
 
 If no initial attitude guess is available, as in most cases of passive debris observation, the inversion problem must be solved globally.
-Filtering approaches can be applied here too; particle \cite{linares2014particle, holzinger2014} and Bayesian Multi-Hypothesis filters \cite{burton2021two, cabrera2023} have shown promise in prior work, but still rely on high-quality priors as with the Kalman filters.
-In the absence of prior knowledge about the object's attitude state, simulated annealing \cite{gagnon2024, clark2020}, genetic algorithms \cite{gagnon2024, piergentili2017, clark2020}, and particle swarm optimizers (PSOs) \cite{clark2020, clark2022, burton2024journal, burton2024scitech, gagnon2024, gagnon2025} can be applied to search the solution space exhaustively.
-
-While much work has performed full state attitude inversion on simulated light curve data \cite{burton2024journal, gagnon2024, robinson2025att}, studies using real data often focus on determining a constant spin rate and axis of rotation for the observed object.
+Filtering approaches can be applied here too; particle \cite{linares2014particle, holzinger2014} and Bayesian Multi-Hypothesis filters \cite{burton2021two, cabrera2023} have shown promise in prior work, but still rely on high-quality priors as with the single-hypothesis Kalman filters.
+In the absence of any prior knowledge about the object's attitude state, simulated annealing \cite{gagnon2024, clark2020}, genetic algorithms \cite{gagnon2024, piergentili2017, clark2020}, and particle swarm optimizers (PSOs) \cite{clark2020, clark2022, burton2024journal, burton2024scitech, gagnon2024, gagnon2025} can be applied to search the solution space exhaustively.
+
+While much work has performed full state attitude inversion on simulated light curve data \cite{burton2024journal, gagnon2024, robinson2025att}, studies using real data often focus on estimating an assumed constant spin rate and axis of rotation for the observed object.
+These assumptions simplifies the estimation process, but can lead to misleading conclusions when the observed object has a complex rotation state.
 
@@
 This work proceeds by describing the inversion method, discussing efficient methods for computing the time-varying probability density of the light curve.
-We present inversion results for a simulated rocket body light curve where the ground truth is known, followed by two cases -- a derelict satellite and rocket body -- using real light curve data collected by the Purdue Optical Ground Station.
+We present inversion results for a simulated rocket body light curve where the ground truth is available to demonstrate the efficacy of the method, followed by results for real light curve data collected by the Purdue Optical Ground Station for a derelict ECHOSTAR satellite and Delta I rocket body.
 A deeper treatment of our method and additional simulated light curve results can be found in \cite{robinson2025att}.
@@
 \begin{equation} \label{eq:nll_loss}
- f(\vctr{x}) = \frac{1}{m}\sum_{k=1}^{m}\left[\ln\sigma_k + \frac{1}{2}\left(\frac{\vctr{S}_k - \hat{\vctr{S}}_k(\vctr{x})}{\sigma_k}\right)^2 \right].
- \end{equation}
-
+ f(\vctr{x}) = \frac{1}{m}\sum_{k=1}^{m}\left[\frac{1}{2} \ln2\pi + \ln\sigma_k + \frac{1}{2}\left(\frac{\vctr{S}_k - \hat{\vctr{S}}_k(\vctr{x})}{\sigma_k}\right)^2 \right].
+ \end{equation}
+
+Note that the constant $0.5\ln2\pi$ could be removed from the objective function, but it is retained such that $\text{exp}\left(-f(\vctr{x})\right)$ is a true likelihood.
 If the reflectivity of the object's faces is not well-known, we rescale the estimated light curve mean $\hat{S}(\vctr{x})$ to have equal $\ell_2$ norm to the observations to resolve the inherent albedo-area ambiguity:
@@
 
-Our multi-start solution method begins by creating $n_\text{guess}$ guesses scattered throughout the solution space -- possibly sampled from a prior distribution based on available knowledge.
-In the most basic case, the initial state guess is sampled for guess $i$ at iteration $k=0$:
+Our multi-start solution method begins by creating $n_\text{sample}$ sampled states scattered throughout the solution space -- possibly sampled from a prior distribution based on available knowledge.
+In the most basic case, the initial states are drawn for sample $i$ at iteration $k=0$:
 
@@
 
-In this work, initial guesses for the state vector are sampled from a prior distribution, usually defined independently for the MRP, angular velocity, and inertia tensor ratios.
-If no information is available, a uniform distribution is sampled from for the first six components of the state.
-In particular, $\vctr{p}_{i,0}(0)$ is uniformly randomly sampled from orientation space, $\vctr{\omega}_{i,0}(0)$ is randomly sampled from a uniform distribution of magnitudes up to $\norm{\vctr{\omega}}_\text{max}$ -- the maximum angular rate at which to initialize guesses.
+In this work, initial state vectors are sampled from a prior distribution, usually defined independently for the MRP, angular velocity, and inertia tensor ratios.
+If no information is available, as is usually the case, a uniform distribution is sampled from for the first six components of the state.
+In particular, $\vctr{p}_{i,0}(0)$ is uniformly randomly sampled from orientation space, $\vctr{\omega}_{i,0}(0)$ is randomly sampled from a uniform distribution of magnitudes up to $\norm{\vctr{\omega}}_\text{max}$ -- the maximum angular rate at which to initialize state samples.
 We determine $\norm{\vctr{\omega}}_\text{max}$ based on the maximum power positive frequency in the observed light curve $f^\star$ in Hertz and an arbitrary safety factor $f_\omega \geq 1$, selected as $f_\omega = 1.25$ for this work:
@@
 
-Uniformly sampling over angular velocity directions, we can express the default sampling strategy for an angular velocity guess as:
+Uniformly sampling over angular velocity directions, we can express the default sampling strategy for an angular velocity state as:
 
@@
 \noindent
-where $g_{1:3} \in \mathcal{N}(0,1)$ and $m=\mathcal{U}(0,\norm{\vctr{\omega}}_\text{max})$.
-
-Even for a well-chosen bound $\norm{\vctr{\omega}}_\text{max}$, there may still be numerous high-quality solutions at significantly lower or higher angular velocities that possess equal $f^\star$ values due to object shape symmetries or harmonics in spin and precession rates, respectively.
-As a result, it is usually unwise to directly constrain the angular velocity to a given magnitude unless additional, unambiguous information is available.
-
-Similarly for the MRP, we first sample a quaternion $\vctr{q}_i(0) = [q_x, q_y, q_z, q_w]^T$ uniformly over orientation space:
+where $g_{1:3} \in \mathcal{N}(0,1)$ and $m=\mathcal{U}(c_1 \norm{\vctr{\omega}}_\text{max}, c_2 \norm{\vctr{\omega}}_\text{max})$.
+The constants $c_1<1$ and $c_2>1$ are chosen to correct for the fact that even for a well-chosen bound $\norm{\vctr{\omega}}_\text{max}$, there can still be numerous high-quality solutions at significantly lower or higher angular velocities that possess equal $f^\star$ values.
+This is a result of object shape symmetries or harmonics in spin and precession rates.
+We select $c_1=0.5$, $c_2=2$ for this work.
+
+For the MRP, we first sample a quaternion $\vctr{q}_i(0) = [q_x, q_y, q_z, q_w]^T$ uniformly over orientation space:
 
@@
 \noindent
-and project the quaternion into MRP space:
+where $g_{1:4} \in \mathcal{N}(0,1)$, and project the quaternion into MRP space:
 
@@
 
-Unlike the MRP and angular velocity, the default sampling method for inertia ratios initial guesses $\left(J_y / J_x\right)_{i,0}$ and $\left(J_z / J_x\right)_{i,0}$ draws from independent, identically distributed Gaussian prior distributions with $\sigma_J$ centered at estimated values derived from the known geometry of the object $\left(J_y / J_x\right)_\text{est}$ and $\left(J_z / J_x\right)_\text{est}$, such that:
+Unlike the MRP and angular velocity, the default sampling method for inertia ratios initial samples $\left(J_y / J_x\right)_{i,0}$ and $\left(J_z / J_x\right)_{i,0}$ draws from independent, identically distributed Gaussian prior distributions with $\sigma_J$ centered at estimated values derived from the known geometry of the object $\left(J_y / J_x\right)_\text{est}$ and $\left(J_z / J_x\right)_\text{est}$, such that:
 
@@
 
-After computing each initial state guess, we run the Broyden-Fletcher-Goldfarb-Shanno (BFGS) algorithm \cite{broyden1970, fletcher1970, goldfarb1970, shanno1970} to optimize each guess to a nearby local minimum $\vctr{x}_{k,i}(0)$ in the objective function such that:
+After sampling each initial state, we run the Broyden-Fletcher-Goldfarb-Shanno (BFGS) algorithm \cite{broyden1970, fletcher1970, goldfarb1970, shanno1970} to optimize each sample to a nearby local minimum $\vctr{x}_{k,i}(0)$ in the objective function such that:
 
@@
 
-Although all initial conditions will converge to some nearby local minimum, many of these minima will possess a low likelihood determined by $-\mathrm{exp}(f(\vctr{x}))$.
+Although all initial conditions will converge to some nearby local minimum, many of these minima will possess a low likelihood determined by $-\mathrm{exp}(f(\vctr{x}))$ where $f(\vctr{x})$ is the objective function in Equation \ref{eq:nll_loss}.
 As we do not generally know how large the geometric and material mismatch between the observed object and the assumed model is, we cannot put a strict bound on a likelihood level that constitutes a high-quality candidate solution.
@@
 
- In any particular illumination and observation condition, a fraction of the facet's area $\tilde{a}_i$ may be Sun- or observer-shadowed due to obstructing geometry before or after light reaches the surface element, respectively.
+ In any particular illumination and observation condition, a fraction of the facet's area $\tilde{a}_i$ might be Sun or observer-shadowed due to obstructing geometry before or after light reaches the surface element, respectively.
 The remainder of the facet's area $\bar{a}_i$ is unshadowed and contributes to the total flux reaching the observer from the object, such that the total area of each facet is $a_i = \tilde{a}_i + \bar{a}_i$.
@@
 
-The unshadowed area $\bar{a}$ may be computed in several ways.
+The unshadowed area $\bar{a}$ can be computed in several ways.
 For convex objects, $\bar{a}=a$ as self-shadowing is impossible.
@@
 As the objects we investigate in this work are well-approximated with relatively simple non-convex meshes, we use the semi-analytic method to efficiently compute unshadowed areas via the Sutherland-Hodgeman algorithm \cite{sutherland1974} for each polygon intersection.
-Further explanation of the polygon clipping procedure for computing the total shadowed area and accelerating this computation during inversion with the so-called shadow cache is included here \textbf{TODO: should this be a `here'?
-wording this felt weird if we're avoiding saying like `the authors'"} \cite{robinson2025att}.
+Further explanation of the polygon clipping procedure for computing the total shadowed area, and accelerating this computation during inversion with the so-called shadow cache is available here \cite{robinson2025att}.
 
@@
 
-The output of Equation \ref{eq:lc_norm} is a light curve, but as many of the noise characteristics of the signal are defined in the image sensor's native unit of ADU, we convert the light curve into the same units via:
+As many of the noise characteristics of the signal are defined in the image sensor's native unit of ADU, we use $f_p(\unitv{s}_k, \unitv{o}_k)$ to compute the light curve in ADU via:
 
@@
 
-The sensor noise is approximated by the independent combination of Poisson dark current $\lambda_\text{dark}$ and readout noise $\sigma_\text{read}^2$ \cite{krag2003}:
+The sensor noise is approximated by the independent combination of Poisson dark current $\lambda_\text{dark}$ and zero-mean Gaussian readout noise $\sigma_\text{read}^2$ \cite{krag2003}:
 
@@
 In order to make this simulated case realistic, we use the presented measurement noise models and introduce a mismatch between the ground truth object used to generate the measurements and the model used in inversion.
-We produce a noisy, observed light curve for a high-fidelity model of the Saturn V second stage scaled down by a factor $10/3$ to act as a template rocket body object, shown on the right of Figure \ref{fig:rb_models}.
+The selected object is a high-fidelity model of the Saturn V second stage scaled down by a factor $10/3$ to act as a template rocket body object, shown on the right of Figure \ref{fig:rb_models}
 
@@
 
-Table \ref{tb:case1_in} describes the simulated observation scenario, placing the rocket body in the orbit of an SL-12 upper stage in geosynchronous orbit, observed by the Purdue Optical Ground Station.
+Table \ref{tb:case1_in} describes the simulated observation scenario used to produce the noisy light curve observations, placing the rocket body in the orbit of an SL-12 upper stage in geosynchronous orbit, observed by the Purdue Optical Ground Station.
 Relevant geographical and optical properties of this ground station are provided in the Appendix in Table \ref{tb:tele_info}.
@@
   \centering
+  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
   \caption{Simulated observation characteristics}
   \vspace*{6pt}
-  \begin{tabular}{|l|l|}
-  \hline
-  \textbf{Variable} & \textbf{Value} \\ \hline
- Target COSPAR ID & 1990-054D \\ \hline
- First obs.\ (UTC) & Mar 8, 2025 02:00:00.000 \\ \hline
- Light curve duration & $5$ minutes \\ \hline
- Observations & $100$ \\ \hline
- Integration time & $0.5$ seconds \\ \hline
+  \begin{tabular}{@{} l l @{}}
+    \toprule
+    Variable & Value \\ \midrule
+    Target COSPAR ID & 1990-054D \\ \grule
+    First obs.\ (UTC) & Mar 8, 2025 02:00:00.000 \\ \grule
+    Light curve duration & $5$ minutes \\ \grule
+    Observations & $100$ \\ \grule
+    Integration time & $0.5$ seconds \\ \brule
   \end{tabular}
@@
   \centering
-  \caption{Reflection properties of materials used in the synthetic data test case. The white paint parameters are estimated qualitatively}
+  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
+  \caption{Reflection properties of materials used in the synthetic data test case. The white paint parameters are estimated qualitatively.}
   \vspace*{6pt}
-  \begin{tabular}{|l|l|l|l|}
-  \hline
-  \textbf{Material} & $C_d$ & $C_s$ & $n$ \\ \hline
- Aluminum \cite{fankhauser2023} & $0.34$ & $0.40$ & $8.9$ \\ \hline
- White paint & $0.9$ & $0.1$ & $1$ \\ \hline
+  \begin{tabular}{@{} l l l l @{}}
+    \toprule
+    Material & $C_d$ & $C_s$ & $n$ \\ \midrule
+    Aluminum \cite{fankhauser2023} & $0.34$ & $0.40$ & $8.9$ \\ \arrayrulecolor{darkgray}\cmidrule(){1-4}
+    White paint & $0.9$ & $0.1$ & $1$ \\ \brule
   \end{tabular}
@@
   \centering
+  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
   \caption{True object attitude profile for all test cases}
   \vspace*{6pt}
-  \begin{tabular}{|l|l|}
-  \hline
-  \textbf{Variable} & \textbf{Value} \\ \hline
- Initial $\vctr{p}(0)$ & $-\frac{1}{3} \begin{bmatrix} 1 & 1 & 1 \end{bmatrix}^T$ \\ \hline
- True $\vctr{\omega}(0)$ ($\text{rad}\cdot\text{s}^{-1}$) & $\vctr{\omega}(0) = \begin{bmatrix} 0.03 & 0.06 & 0.03 \end{bmatrix}^T$ \\ \hline
- Inertia tensor ratios & $J_y / J_x = 1, \: J_z / J_x = 0.25$ \\ \hline
- External torque $\vctr{M}$ & $\begin{bmatrix} 0 & 0 & 0 \end{bmatrix}^T$ \\ \hline
+  \begin{tabular}{@{} l l @{}}
+    \toprule
+    Variable & Value \\ \midrule
+    Initial $\vctr{p}(0)$ & $-\frac{1}{3} \begin{bmatrix} 1 & 1 & 1 \end{bmatrix}^T$ \\ \grule
+    True $\vctr{\omega}(0)$ ($\text{rad}\cdot\text{s}^{-1}$) & $\vctr{\omega}(0) = \begin{bmatrix} 0.03 & 0.06 & 0.03 \end{bmatrix}^T$ \\ \grule
+    Inertia tensor ratios & $J_y / J_x = 1, \: J_z / J_x = 0.25$ \\ \grule
+    External torque $\vctr{M}$ & $\begin{bmatrix} 0 & 0 & 0 \end{bmatrix}^T$ \\ \brule
   \end{tabular}
@@
 Two distinct inversion cases are computed.
-Case 1a presents inversion results given no attitude information besides an estimate of the inertia tensor ratios derived from the known geometry, while Case 1b presents results given good initial guesses for orientation and angular velocity.
+Case 1a presents inversion results given no attitude information besides an estimate of the inertia tensor ratios derived from the known geometry, while Case 1b presents results given good prior attitude state distributions.
 Tables \ref{tb:case1a_prior} and \ref{tb:case1b_prior} summarize the initial condition sampling strategies employed for each case.
 
-For both cases, a total of $n_\text{guess}=10^5$ initial state guesses are created via the sampling schemes described their respective tables.
-BFGS minimization is then performed for each initial guess to locally solve the optimization problem given by Equation \ref{eq:opt_problem}.
+For both cases, a total of $n_\text{sample}=10^5$ initial state samplees are created via the sampling schemes described their respective tables.
+BFGS minimization is then performed for each initial state sample to locally solve the optimization problem given by Equation \ref{eq:opt_problem}.
 
@@
   \centering
+  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
   \caption{Initial condition sampling parameters used for inversion in Case 1a}
   \vspace*{6pt}
-  \begin{tabular}{|l|l|}
-  \hline
-  \textbf{Variable} & \textbf{Value} \\ \hline
-  $\vctr{p}$ sampling & Uniform, via Equation \ref{eq:mrp_sampler} \\ \hline
-  $\vctr{\omega}$ sampling & Uniform, via Equation \ref{eq:omega_sampler} \\ \hline
-  $J$ sampling & Gaussian, via Equation \ref{eq:omega_sampler} \\ \hline
-  $\sigma_J$ & $0.05$ $[\text{nondim}]$ \\ \hline
-  $\norm{\vctr{\omega}}_\text{max}$ & $5.26$ $[\text{deg}\cdot \text{s}^{-1}]$ \\ \hline
+  \begin{tabular}{@{} l l @{}}
+    \toprule
+    Variable & Value \\ \midrule
+    $\vctr{p}$ sampling & Uniform, via Equation \ref{eq:mrp_sampler} \\ \grule
+    $\vctr{\omega}$ sampling & Uniform, via Equation \ref{eq:omega_sampler} \\ \grule
+    $J$ sampling & Gaussian, via Equation \ref{eq:omega_sampler} \\ \grule
+    $\sigma_J$ & $0.1$ $[\text{nondim}]$ \\ \grule
+    $\norm{\vctr{\omega}}_\text{max}$ & $5.26$ $[\text{deg}\cdot \text{s}^{-1}]$ \\ \brule
   \end{tabular}
@@
   \centering
+  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
   \caption{Informative prior parameters used for inversion in Case 1b.
-The standard deviations $\sigma_\omega$ and $\sigma_J$ are for spherically-symmetric Gaussians with centered on the ground truth.
+The standard deviations $\sigma_\omega$ and $\sigma_J$ are for spherically-symmetric Gaussians centered on the ground truth.
 MRPs are sampled by rotating by $\theta\sim \mathcal{N}(0,\sigma_p)$ degrees away from the ground truth orientation around an axis uniformly sampled on the sphere.}
   \vspace*{6pt}
-  \begin{tabular}{|l|l|}
-  \hline
-  \textbf{Variable} & \textbf{Value} \\ \hline
-  $\vctr{p}$ sampling & Gaussian, $\sigma_\omega = 0.05$ $[\text{deg}\cdot\text{s}^{-1}]$ \\ \hline
-  $\vctr{\omega}$ sampling & Gaussian, $\sigma_p=20$ $[\text{deg}]$ \\ \hline
-  $J$ sampling & Gaussian, $\sigma_J = 0.05$ $[\text{nondim}]$ \\ \hline
+  \begin{tabular}{@{} l l @{}}
+    \toprule
+    Variable & Value \\ \midrule
+    $\vctr{p}$ sampling & Gaussian, $\sigma_p=20$ $[\text{deg}]$ \\ \grule
+    $\vctr{\omega}$ sampling & Gaussian, $\sigma_\omega = 0.1$ $[\text{deg}\cdot\text{s}^{-1}]$ \\ \grule
+    $J$ sampling & Gaussian, $\sigma_J = 0.05$ $[\text{nondim}]$ \\ \brule
   \end{tabular}
@@
 
-Solutions are then classified as candidates for the ground truth with $\ell = 1/10$ via Equation \ref{eq:ell_selection_criteria}.
+Solutions are then classified as candidates for the ground truth with $\ell = 1/2$ via Equation \ref{eq:ell_selection_criteria}.
 Figure \ref{fig:conv_lcs_synth} shows the mean light curves produced by the identified candidate solutions.
@@
 Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} display the candidate MRP and angular velocity of the candidate solutions determined via the procedure in Section \ref{sec:candidate_sols}.
-In the left Case 1a figures for orientation and angular velocity, we notice that the ground truth lies squarely on a surface containing numerous candidate solution.
+In the left Case 1a figures for orientation and angular velocity, we notice that the ground truth in red lies squarely on a surface containing numerous candidate solutions.
 This indicates that the local minima have correctly converged to the non-singular subspace of states that approximate the observations well for the assumed shape model.
@@
 Similarly in orientation space, shown in Figure \ref{fig:sigma_sols1}, the solutions fall on two distinct manifolds, one containing the truth -- likely resulting from a reflection of the object's body frame about the bisector between the Sun and observer vectors \cite{marto2024, burton2024journal}.
-These additional families of solutions highlight the reality that a wide variety of attitude profiles may produce nearly identical observations.
-
-The Case 1b results on the right within Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} show that the majority of the initial guesses sampled close to the truth have converged to a local subset of the solution manifolds identified by Case 1a.
+These additional families of solutions highlight the reality that a wide variety of attitude profiles can produce nearly identical observations.
+
+The Case 1b results on the right within Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} show that the majority of the initial states sampled close to the truth have converged to a local subset of the solution manifolds identified by Case 1a.
 We should not expect these solutions to have shrunk down exponentially towards the ground truth -- the mismatch in the assumed shape model and the measurement noise inherently limits the minimum size of the feasible solution space.
@@
     \centering
-    \includegraphics[width=\textwidth]{sigma_solsRocket Body Prior.png}
+    \includegraphics[width=\textwidth]{sigma_solsRocket Body Prior_pin.png}
     \caption{}
@@
     \centering
-    \includegraphics[width=\textwidth]{omega_sols_Rocket Body.png}
+    \includegraphics[width=\textwidth]{omega_sols_Rocket Body_pin.png}
     \caption{}
@@
     \centering
-    \includegraphics[width=\textwidth]{omega_sols_Rocket Body Prior.png}
+    \includegraphics[width=\textwidth]{omega_sols_Rocket Body Prior_pin.png}
     \caption{}
@@
 Notably, the majority of candidate solutions remain within a few standard deviations of the mean of their initial distribution, with the remaining minority exploring directions that maintain the same precession rate.
-When a good initial guess for the other attitude state components is available in Case 1b, the inertia tensor solutions are more well-clustered, although due to the low observability of the rocket body's spin rate in the observed light curve, there are still ambiguous solutions with higher or lower $J_z / J_x$ values.
+When a good prior for the other attitude state components is available in Case 1b, the inertia tensor solutions are more well-clustered, although due to the low observability of the rocket body's spin rate in the observed light curve, there are still ambiguous solutions with higher or lower $J_z / J_x$ values.
 
@@
 Figure \ref{fig:w_vs_ang_error_sols1} summarizes the distribution of candidate solutions with respect to the ground truth by displaying the angular velocity magnitude error and angular orientation error compared to the ground truth.
-Case 1a shows significant variability in angular velocity magnitude, although the mean error is low at $2\%$.
-Due to the underdetermined orientation solution space, there are far more solutions at high angular offsets from the truth orientation, leading to a large mean angular error of $127^\circ$.
-For Case 1b, the mean angular error is much lower at $25^\circ$, although convergence here is still inherently limited by the noisy measurements and model mismatch.
+Case 1a shows significant variability in angular velocity magnitude, although this is almost entirely due to the unobservability of the spin rate about the object's axis of symmetry in the light curve.
+The mean precession rate error is low at $0.3\%$.
+Due to the underdetermined orientation solution space, there are far more solutions at high angular offsets from the truth orientation, leading to a large mean angular error of $126^\circ$.
+For Case 1b, the mean orientation error is much lower at $18^\circ$, although convergence here is still inherently limited by the noisy measurements and model mismatch.
 
@@
   \centering
+  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
   \caption{Delta I rocket body observation characteristics}
   \vspace*{6pt}
-  \begin{tabular}{|l|l|}
-  \hline
-  \textbf{Variable} & \textbf{Value} \\ \hline
- Target COSPAR ID & 1984-115C \\ \hline
- First obs.\ (UTC) & Mar 2, 2025 01:53:30.251 \\ \hline
- Light curve duration & $7.5$ minutes \\ \hline
- Observations & $100$ ($92$ extracted) \\ \hline
- Obs.\ frequency & $0.222$ Hz \\ \hline
- Integration time & $3$ seconds \\ \hline
+  \begin{tabular}{@{} l l @{}}
+    \toprule
+    Variable & Value \\ \midrule
+    Target COSPAR ID & 1984-115C \\ \grule
+    First obs.\ (UTC) & Mar 2, 2025 01:53:30.251 \\ \grule
+    Light curve duration & $7.5$ minutes \\ \grule
+    Observations & $100$ ($92$ extracted) \\ \grule
+    Obs.\ frequency & $0.222$ Hz \\ \grule
+    Integration time & $3$ seconds \\ \brule
   \end{tabular}
@@
 This light curve shows clear periodicity, but has a high measurement standard deviation and a relatively low sampling rate due to the object's low brightness and the sensor readout time.
-Inversion is performed using $n_\text{guess}=10^4$ initial guesses with the material and angular velocity assumptions listed in Table \ref{tb:case2_ass}.
-As in Case 1a, orientation and angular velocity initial guesses are sampled via Equations \ref{eq:mrp_sampler} and \ref{eq:omega_sampler}, respectively.
-All initial inertia tensor guesses are set to the estimated $J_\text{R/B}$.
+Inversion is performed using $n_\text{sample}=10^4$ samples with the material and angular velocity assumptions listed in Table \ref{tb:case2_ass}.
+As in Case 1a, orientation and angular velocity states are sampled via Equations \ref{eq:mrp_sampler} and \ref{eq:omega_sampler}, respectively.
+All initial inertia tensor samples are set to the estimated $J_\text{R/B}$.
 The candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case2_s}, orientation MRPs and angular velocities in Figure \ref{fig:case2_pw}, and inertia ratios in Figure \ref{fig:case2_i}.
@@
   \centering
+  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
   \caption{Delta I rocket body inversion assumptions}
   \vspace*{6pt}
-  \begin{tabular}{|l|l|}
-  \hline
-  \textbf{Variable} & \textbf{Value} \\ \hline
- Uniform BRDF & $C_d=0.2$, $C_s=0.8$, $n=30$ \\ \hline
- $\norm{\vctr{\omega}}_\text{max}$ & $8.31$ $\text{deg} \cdot \text{s}^{-1}$ \\ \hline
- $\ell$ & $1/100$ \\ \hline
+  \begin{tabular}{@{} l l @{}}
+    \toprule
+    Variable & Value \\ \midrule
+    Uniform BRDF & $C_d=0.2$, $C_s=0.8$, $n=30$ \\ \grule
+    $\norm{\vctr{\omega}}_\text{max}$ & $6.93$ $\text{deg} \cdot \text{s}^{-1}$ \\ \grule
+    $\ell$ & $1/100$ \\ \brule
   \end{tabular}
@@
     \centering
-    \includegraphics[width=\textwidth]{sphx_glr_results_005.png}
+    \includegraphics[width=\textwidth]{delta_ang_vel_pin.png}
     \caption{}
@@
 The candidates show some rough clustering in both orientation and angular velocity space, although they vary significantly in angular velocity magnitude.
-The inertia tensor solutions display similar variability, although the solutions are spread relatively uniformly from the initial guess, possibly indicating that the geometry-driven estimate is approximately unbiased.
+The inertia tensor solutions display similar variability, although the solutions are spread relatively uniformly from the initial estimate, possibly indicating that the geometry-driven estimate is approximately unbiased.
 Taken together, these results indicate that the object could be spinning in a number of attitude profiles, although the light curve inversion process has significantly narrowed the search space to small subsets of the original unconstrained eight-dimensional volume.
-These constraints may be combined with additional light curves, more information about the object's true mass distribution, or reflectivity to further refine the solution set.
+These constraints can be combined with additional light curves, more information about the object's true mass distribution, or reflectivity to further refine the solution set.
 
@@
   \centering
+  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
   \caption{ECHOSTAR II observation characteristics}
   \vspace*{6pt}
-  \begin{tabular}{|l|l|}
-  \hline
-  \textbf{Variable} & \textbf{Value} \\ \hline
- Target COSPAR ID & 1996-055A \\ \hline
- First obs.\ (UTC) & Feb 26, 2025 04:16:12.279 \\ \hline
- Light curve duration & $19.5$ minutes \\ \hline
- Observations & $456$ ($395$ extracted) \\ \hline
- Obs.\ frequency & $0.388$ Hz \\ \hline
- Integration time & $1$ seconds \\ \hline
+  \begin{tabular}{@{} l l @{}}
+    \toprule
+    Variable & Value \\ \midrule
+    Target COSPAR ID & 1996-055A \\ \grule
+    First obs.\ (UTC) & Feb 26, 2025 04:16:12.279 \\ \grule
+    Light curve duration & $19.5$ minutes \\ \grule
+    Observations & $456$ ($395$ extracted) \\ \grule
+    Obs.\ frequency & $0.388$ Hz \\ \grule
+    Integration time & $1$ seconds \\ \brule
   \end{tabular}
@@
 This light curve is sampled more densely than the Delta I observations, and clearly has significantly more intense specular peaks.
-Inversion is performed using $n_\text{guess}=10^4$ initial guesses with the material and angular velocity assumptions listed in Table \ref{tb:case3_ass}.
-Initial guess sampling is accomplished identically to Case 2, with a different maximum angular veloity and initial inertia ratios.
+Inversion is performed using $n_\text{sample}=10^4$ samples with the material and angular velocity assumptions listed in Table \ref{tb:case3_ass}.
+Initial state sampling is accomplished identically to Case 2, with a different maximum angular veloity and initial inertia ratios.
 The candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case3_s}, orientation MRPs and angular velocities in Figure \ref{fig:case3_pw}, and inertia ratios in Figure \ref{fig:case3_i}.
@@
   \centering
+  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
   \caption{ECHOSTAR II inversion assumptions}
   \vspace*{6pt}
-  \begin{tabular}{|l|l|}
-  \hline
-  \textbf{Variable} & \textbf{Value} \\ \hline
- Aluminum BRDF & $C_d=0.22$, $C_s=0.4$, $n=5$ \\ \hline
- MLI BRDF & $C_d=0.05$, $C_s=0.24$, $n=20$ \\ \hline
- Solar panel BRDF & $C_d=0.05$, $C_s=0.13$, $n=10$ \\ \hline
-  $\norm{\vctr{\omega}}_\text{max}$ & $2.32$ $\text{deg} / \text{s}^{-1}$ \\ \hline
-  $\ell$ & $1/1000$ \\ \hline
+  \begin{tabular}{@{} l l @{}}
+    \toprule
+    Variable & Value \\ \midrule
+    Aluminum BRDF & $C_d=0.22$, $C_s=0.4$, $n=5$ \\ \grule
+    MLI BRDF & $C_d=0.05$, $C_s=0.24$, $n=20$ \\ \grule
+    Solar panel BRDF & $C_d=0.05$, $C_s=0.13$, $n=10$ \\ \grule
+    $\norm{\vctr{\omega}}_\text{max}$ & $2.32$ $\text{deg} / \text{s}^{-1}$ \\ \grule
+    $\ell$ & $1/1000$ \\ \brule
   \end{tabular}
@@
 We presented attitude inversion results for one case with synthetic data and two cases with real observations.
-Unlike other approaches, our method successfully identifies the numerous ambiguous solutions families -- originating from symmetries in the observation geometry and the object shape -- which may produce the same observations.
+Unlike other approaches, our method successfully identifies the numerous ambiguous solutions families -- originating from symmetries in the observation geometry and the object shape -- which can produce the same observations.
 These disparate candidate solutions can serve as starting points for more detailed exploration as additional light curves or further object shape information is obtained.
 
+\section{Global TODO list}
+
+\begin{itemize}
+  % \item Add the constant term back into the objective function (in text and code) so we can start treating it as a real likelihood, useful for the future
+  \item Mention the unknown solar panel rotation angle for ECHOSTAR
+  \item make a table of execution times
+  % \item Update perspective on some the case 1a angular velocity plot
+  % \item Update $n_\text{sample}$ values everywhere to reflect the scripts
+  % \item Update $\omega$ sampling scheme to go from $0.5-2$ as reflected in scripts
+  % \item Re-run all cases to ensure that values are up to date with recent changes, update where those statistics are reported in the text
+  \item Try to get rid of big gaps where possible
+  % \item Reformat tables so they're attractive to look at
+\end{itemize}
+
 \section*{Acknowledgments}
@@
   \centering
+  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
   \caption{Telescope information for synthetic and real data observations}
   \vspace*{6pt}
-  \begin{tabular}{|l|l|}
-  \hline
-  \textbf{Variable} & \textbf{Value} \\ \hline
- CCD sensor & KAF-16803 \\ \hline
- Effective aperture area $A_\circ$ & $0.076$ $m^2$ \\ \hline
- Observer location & $32.900^\circ \textrm{ N}, -105.533^\circ \textrm{ W}$ \\ \hline
- Observer altitude & $2.24$ km above MSL \\ \hline
- Source pixels $n_\text{pix}$ & $100$ \\ \hline
- Read noise $\sigma_\text{read}$ & $9$ $\text{ADU} / \text{pixel}$ \\ \hline
- Dark current rate $\lambda_\text{dark}$ & $0.01$ $\text{ADU} / \text{pixel} / \text{s}$ \\ \hline
- Flat field strength $f_k$ & $0.01$ nondim \\ \hline
- Gain $g$ & $5.1$ $e^- / \text{ADU}$ \\ \hline
+  \begin{tabular}{@{} l l @{}}
+    \toprule
+    Variable & Value \\ \midrule
+    CCD sensor & KAF-16803 \\ \grule
+    Aperture area ($A_\circ$) & $0.076$ m$^2$ \\ \grule
+    Observer location & $32.900^\circ$ N, $-105.533^\circ$ W \\ \grule
+    Observer altitude & $2.24$ km above MSL \\ \grule
+    Read noise ($\sigma_\text{read}$) & $9$ ADU/pixel \\ \grule
+    Dark current rate ($\lambda_\text{dark}$) & $0.01$ ADU/pixel/s \\ \grule
+    Flat field strength ($f_k$) & $0.01$ (dimensionless) \\ \grule
+    Gain ($g$) & $5.1$ e$^-$/ADU \\ \brule
   \end{tabular}
@@
   \centering
+  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
   \caption{BFGS solver configuration for all results cases}
   \vspace*{6pt}
-  \begin{tabular}{|l|l|}
-  \hline
-  \textbf{Variable} & \textbf{Value} \\ \hline
- Maximum function evaluations & $1000$ \\ \hline
- Maximum iterations & $100$ \\ \hline
- Finite difference step size & $1 \cdot 10^{-5}$ \\ \hline
- Gradient $\infty$-norm tolerance & $1 \cdot 10^{-5}$ \\ \hline
+  \begin{tabular}{@{} l l @{}}
+    \toprule
+    Variable & Value \\ \midrule
+    Maximum function evaluations & $1000$ \\ \grule
+    Maximum iterations & $100$ \\ \grule
+    Finite difference step size & $1 \cdot 10^{-5}$ \\ \grule
+    Gradient $\infty$-norm tolerance & $1 \cdot 10^{-5}$ \\ \brule
   \end{tabular}

### 2025-03-25 paper.tex
submitted

@@
 
-% \makeatletter
-% \AtBeginDocument{%
-%   \expandafter\renewcommand\expandafter\subsubsection\expandafter{%
-%     \expandafter\FloatBarrier\subsubsection
-%   }%
-% }
-% \makeatother
-
+\usepackage{hyperref}
 \usepackage{colortbl}
@@
 
-\newcommand{\figbig}[0]{0.5\textwidth}
-\newcommand{\figmed}[0]{0.4\textwidth}
+\newcommand{\figbig}[0]{0.4\textwidth}
+\newcommand{\figmed}[0]{0.35\textwidth}
 \newcommand{\figsmall}[0]{0.3\textwidth}
 
-\newcommand{\subfigmed}[0]{0.4\textwidth}
+\newcommand{\subfigbig}[0]{0.4\textwidth}
 \newcommand{\subfigsm}[0]{0.29\textwidth}
-\newcommand{\subfigbig}[0]{0.43\textwidth}
+\newcommand{\subfigmed}[0]{0.43\textwidth}
 
@@
 This class of methods includes single Kalman filters \cite{burton2021two, gagnon2024, wetterer2009}, and multi-filter multiple-model adaptive estimation (MMAE) schemes \cite{linares2014space, dianetti2020}.
-These approaches can be very effective if the initial state estimate is close to the truth, but have been reported to diverge even in cases when the initial guess is within $5^\circ$ of the ground truth orientation and $0.03^\circ/s^{-1}$ in angular velocity error \cite{gagnon2024} due to measurement nonlinearity and ambiguities.
+These approaches can be very effective if the initial state estimate is close to the truth, but have been reported to diverge even in cases when the initial guess is within $5^\circ$ of the ground truth orientation and $0.03^\circ\cdot s^{-1}$ in angular velocity error \cite{gagnon2024} due to measurement nonlinearity and ambiguities.
 
@@
 
-While much work has performed full state attitude inversion on simulated light curve data \cite{burton2024journal, gagnon2024, robinson2025att}, studies using real data often focus on estimating an assumed constant spin rate and axis of rotation for the observed object.
-These assumptions simplifies the estimation process, but can lead to misleading conclusions when the observed object has a complex rotation state.
+While much work has performed full state attitude inversion on simulated light curve data \cite{burton2024journal, gagnon2024, robinson2025att}, the applicability of many of these methods to real observations is untested.
+Studies that do work with real data often focus on estimating an assumed constant spin rate and axis of rotation for the observed object.
+These assumptions are made to simplify the estimation process but can lead to misleading conclusions when the observed object has a complex rotation state.
 
@@
 
-The simplifying assumptions of constant, single-axis rotation that are commonly made when working with real data enable elegant solutions, but break down when the observed object is in a complex spin profile.
+The simplifying assumptions of constant, single-axis rotation that are commonly made when working with real data enable elegant solutions but break down when the observed object is in a complex spin profile.
 Methods for full attitude state solutions -- time-varying orientation and angular velocities -- have been developed \cite{clark2022, burton2024journal, gagnon2024, linares2014particle}, but are seldom applied to real data.
-Works that have been successful at estimating a full rigid body attitude state from real data include genetic algorithms \cite{piergentili2020, gallucci2020} and brute-force grid search \cite{shafer2017}.
+Works that have been successful at estimating a full rigid-body attitude state from real data include genetic algorithms \cite{piergentili2020, gallucci2020} and brute-force grid search \cite{shafer2017}.
 Any full-state estimation method must tackle the inherent computational complexity, lack of knowledge of material reflectivities, and the presence of measurement noise.
@@
 Further, our method differs from other global estimation methods in that we do not seek to find one optimal solution and instead search for a collection of high-quality local minima to directly study the many ambiguous solutions for a given light curve.
-Our approach is designed specifically for use with real data as we take time-varying measurement noise and uncertainties in the body moments of inertia into account explicitly.
+Our approach is designed specifically for use with real data as we explicitly take time-varying measurement noise and uncertainties in the body moments of inertia into account.
 
@@
 
-This work proceeds by describing the inversion method, discussing efficient methods for computing the time-varying probability density of the light curve.
+This work proceeds by describing the inversion method, discussing efficient methods for computing the time-varying probability density of the light curve, and reporting inversion test case results.
 We present inversion results for a simulated rocket body light curve where the ground truth is available to demonstrate the efficacy of the method, followed by results for real light curve data collected by the Purdue Optical Ground Station for a derelict ECHOSTAR satellite and Delta I rocket body.
@@
 
+We use the Runge-Kutta 4 numerical integration algorithm for all state propagation as it is computationally inexpensive and has low error -- usually below $1^\circ$ over the course of a well-sampled light curve.
+
 \subsection{Objective Function}
@@
 \begin{equation} \label{eq:nll_loss}
- f(\vctr{x}) = \frac{1}{m}\sum_{k=1}^{m}\left[\frac{1}{2} \ln2\pi + \ln\sigma_k + \frac{1}{2}\left(\frac{\vctr{S}_k - \hat{\vctr{S}}_k(\vctr{x})}{\sigma_k}\right)^2 \right].
+  \begin{split}
+ f(\vctr{x}) = \frac{1}{m}\sum_{k=1}^{m}\Biggl[&\frac{1}{2} \ln2\pi + \ln\sigma_k \\ + &\frac{1}{2}\left(\frac{\vctr{S}_k - \hat{\vctr{S}}_k(\vctr{x})}{\sigma_k}\right)^2 \Biggr].
+  \end{split}
  \end{equation}
@@
 In this work, initial state vectors are sampled from a prior distribution, usually defined independently for the MRP, angular velocity, and inertia tensor ratios.
-If no information is available, as is usually the case, a uniform distribution is sampled from for the first six components of the state.
-In particular, $\vctr{p}_{i,0}(0)$ is uniformly randomly sampled from orientation space, $\vctr{\omega}_{i,0}(0)$ is randomly sampled from a uniform distribution of magnitudes up to $\norm{\vctr{\omega}}_\text{max}$ -- the maximum angular rate at which to initialize state samples.
-We determine $\norm{\vctr{\omega}}_\text{max}$ based on the maximum power positive frequency in the observed light curve $f^\star$ in Hertz and an arbitrary safety factor $f_\omega \geq 1$, selected as $f_\omega = 1.25$ for this work:
-
-\begin{equation} \label{eq:ang_vel_max}
-  \norm{\vctr{\omega}}_\text{max} = 2\pi f_\omega f^\star.
-\end{equation}
-
+If no information is available, as is usually the case, a uniform distribution is sampled from the first six components of the state.
+In particular, $\vctr{p}_{i,0}(0)$ is uniformly randomly sampled from orientation space, $\vctr{\omega}_{i,0}(0)$ is randomly sampled from a uniform distribution of magnitudes up to $\norm{\vctr{\omega}}_\text{mid}$ -- the mean angular rate at which to initialize state samples.
+We determine this value based on the maximum power positive frequency in the observed light curve $f^\star$ in Hertz as simply $\norm{\vctr{\omega}}_\text{mid} = 2\pi f_\omega$.
 Uniformly sampling over angular velocity directions, we can express the default sampling strategy for an angular velocity state as:
@@
 \noindent
-where $g_{1:3} \in \mathcal{N}(0,1)$ and $m=\mathcal{U}(c_1 \norm{\vctr{\omega}}_\text{max}, c_2 \norm{\vctr{\omega}}_\text{max})$.
-The constants $c_1<1$ and $c_2>1$ are chosen to correct for the fact that even for a well-chosen bound $\norm{\vctr{\omega}}_\text{max}$, there can still be numerous high-quality solutions at significantly lower or higher angular velocities that possess equal $f^\star$ values.
+where $g_{1:3} \in \mathcal{N}(0,1)$ and $m=\mathcal{U}(c_1 \norm{\vctr{\omega}}_\text{mid}, c_2 \norm{\vctr{\omega}}_\text{mid})$.
+The constants $c_1<1$ and $c_2>1$ are chosen to correct for the fact that even for a well-chosen $\norm{\vctr{\omega}}_\text{mid}$, there can still be numerous high-quality solutions at significantly lower or higher angular velocities that possess equal $f^\star$ values.
 This is a result of object shape symmetries or harmonics in spin and precession rates.
@@
 
-\begin{equation}
+\begin{equation} \label{eq:j_sampler}
   \begin{bmatrix}
@@
 
-\subsection{Generating the Observed Light Curve}
-
-In order to make this simulated case realistic, we use the presented measurement noise models and introduce a mismatch between the ground truth object used to generate the measurements and the model used in inversion.
+\subsection{Generating the Simulated Light Curve}
+
+To make this simulated case realistic, we use the presented measurement noise models and introduce a mismatch between the ground truth object used to generate the measurements and the model used for inversion.
 The selected object is a high-fidelity model of the Saturn V second stage scaled down by a factor $10/3$ to act as a template rocket body object, shown on the right of Figure \ref{fig:rb_models}
-
-% \begin{figure}[H]
-%   \centering
-%   \includegraphics[width=\figsmall]{rocket_bodies2.png}
-%   \caption{Simulated truth rocket body model \cite{nasa_models}.}
-%   \label{fig:meas_model}
-% \end{figure}
 
@@
   \centering
-  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
+  \renewcommand{\arraystretch}{1.2} 
   \caption{Simulated observation characteristics}
@@
   \centering
-  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
+  \renewcommand{\arraystretch}{1.2} 
   \caption{Reflection properties of materials used in the synthetic data test case. The white paint parameters are estimated qualitatively.}
@@
   \centering
-  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
+  \renewcommand{\arraystretch}{1.2} 
   \caption{True object attitude profile for all test cases}
@@
 
-% \begin{figure}[H]
-%   \centering
-%   \includegraphics[width=\figmed]{light_curves_adu_rocket_body_sdc.png}
-%   \caption{Simulated observed flux for the rocket body.}
-%   \label{fig:obs_adu_synth}
-% \end{figure}
-
 \begin{figure}[H]
@@
 
-% \begin{figure}[H]
-%   \centering
-%   \includegraphics[width=\figsmall]{light_curves_std_rocket_body_sdc.png}
-%   \caption{Simulated observed light curve standard deviation computed with Equation \ref{eq:sigma_total}.}
-%   \label{fig:obs_std_synth}
-% \end{figure}
-
-\subsection{Case 1a and 1b: Synthetic Light Curve Optimization Results}
+\subsection{Case 1U and 1P: Synthetic Light Curve Optimization Results}
 
@@
 Two distinct inversion cases are computed.
-Case 1a presents inversion results given no attitude information besides an estimate of the inertia tensor ratios derived from the known geometry, while Case 1b presents results given good prior attitude state distributions.
-Tables \ref{tb:case1a_prior} and \ref{tb:case1b_prior} summarize the initial condition sampling strategies employed for each case.
-
-For both cases, a total of $n_\text{sample}=10^5$ initial state samplees are created via the sampling schemes described their respective tables.
-BFGS minimization is then performed for each initial state sample to locally solve the optimization problem given by Equation \ref{eq:opt_problem}.
+Case 1U presents inversion results given no attitude information besides an estimate of the inertia tensor ratios derived from the known geometry -- hence requiring a uniform search of the solution space.
+In contrast, Case 1P presents results given good prior knowledge of attitude state distributions.
+Table \ref{tb:case1Ub_priors} summarizes the initial condition sampling strategies employed for each case.
+For Case 1P, the standard deviations $\sigma_\omega$ and $\sigma_J$ are for spherically-symmetric Gaussians centered on the ground truth.
+MRPs are sampled by rotating by $\theta\sim \mathcal{N}(0,\sigma_p)$ degrees away from the ground truth orientation around an axis uniformly sampled on the sphere.
 
@@
   \centering
-  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
-  \caption{Initial condition sampling parameters used for inversion in Case 1a}
+  \renewcommand{\arraystretch}{1.2} 
+  \caption{Initial condition sampling parameters used for inversion in Cases 1U (no attitude information) and 1P (prior attitude knowledge)}
   \vspace*{6pt}
@@
     \toprule
-    Variable & Value \\ \midrule
-    $\vctr{p}$ sampling & Uniform, via Equation \ref{eq:mrp_sampler} \\ \grule
-    $\vctr{\omega}$ sampling & Uniform, via Equation \ref{eq:omega_sampler} \\ \grule
-    $J$ sampling & Gaussian, via Equation \ref{eq:omega_sampler} \\ \grule
-    $\sigma_J$ & $0.1$ $[\text{nondim}]$ \\ \grule
-    $\norm{\vctr{\omega}}_\text{max}$ & $5.26$ $[\text{deg}\cdot \text{s}^{-1}]$ \\ \brule
+    \multicolumn{2}{c}{\textbf{Case 1U}} \\
+    \midrule
+ State & Sampling strategy \\ \midrule
+    $\vctr{p}$ & Uniform, via Equation \ref{eq:mrp_sampler} \\ \grule
+    $\vctr{\omega}$ & $\norm{\vctr{\omega}}_\text{mid} = 5.26$ $[\text{deg}\cdot \text{s}^{-1}]$, via Eq.\ \ref{eq:omega_sampler} \\ \grule
+    $J$ &  $\sigma_J = 0.1$, via Equation \ref{eq:j_sampler} \\ \brule
+    \addlinespace
+    \toprule
+    \multicolumn{2}{c}{\textbf{Case 1P}} \\
+    \midrule
+ State & Sampling strategy \\ \midrule
+    $\vctr{p}$ & Gaussian, $\sigma_p=20$ $[\text{deg}]$ \\ \grule
+    $\vctr{\omega}$ & Gaussian, $\sigma_\omega = 0.1$ $[\text{deg}\cdot\text{s}^{-1}]$ \\ \grule
+    $J$ & Gaussian, $\sigma_J = 0.05$ $[\text{nondim}]$ \\ \brule
   \end{tabular}
-  \label{tb:case1a_prior}
+  \label{tb:case1Ub_priors}
 \end{table}
 
-\begin{table}[H]
-  \centering
-  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
-  \caption{Informative prior parameters used for inversion in Case 1b.
-The standard deviations $\sigma_\omega$ and $\sigma_J$ are for spherically-symmetric Gaussians centered on the ground truth.
-MRPs are sampled by rotating by $\theta\sim \mathcal{N}(0,\sigma_p)$ degrees away from the ground truth orientation around an axis uniformly sampled on the sphere.}
-  \vspace*{6pt}
-  \begin{tabular}{@{} l l @{}}
-    \toprule
-    Variable & Value \\ \midrule
-    $\vctr{p}$ sampling & Gaussian, $\sigma_p=20$ $[\text{deg}]$ \\ \grule
-    $\vctr{\omega}$ sampling & Gaussian, $\sigma_\omega = 0.1$ $[\text{deg}\cdot\text{s}^{-1}]$ \\ \grule
-    $J$ sampling & Gaussian, $\sigma_J = 0.05$ $[\text{nondim}]$ \\ \brule
-  \end{tabular}
-  \label{tb:case1b_prior}
-\end{table}
+For both cases, a total of $n_\text{sample}=10^5$ initial state samples are created via the sampling schemes described in their respective tables.
+BFGS minimization is then performed for each initial state sample to locally solve the optimization problem given by Equation \ref{eq:opt_problem}.
+Both cases ran to convergence on an Apple M1 CPU in $3.2$ hours.
 
 Solutions are then classified as candidates for the ground truth with $\ell = 1/2$ via Equation \ref{eq:ell_selection_criteria}.
+$2026$ and $39683$ candidate solutions were identified for Cases 1U and 1P, respectively.
 Figure \ref{fig:conv_lcs_synth} shows the mean light curves produced by the identified candidate solutions.
@@
   \centering
-  \includegraphics[width=\figbig]{converged_solutions_1_Rocket Body.png}
+  \includegraphics[width=\figmed]{converged_solutions_1_Rocket Body.png}
   \caption{Candidate solution light curves found through BFGS optimization.}
@@
 
-Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} display the candidate MRP and angular velocity of the candidate solutions determined via the procedure in Section \ref{sec:candidate_sols}.
-In the left Case 1a figures for orientation and angular velocity, we notice that the ground truth in red lies squarely on a surface containing numerous candidate solutions.
+Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} display the candidate MRP and angular velocity of the candidate solutions determined via the procedure detailed in Section \ref{sec:candidate_sols}.
+In the Case 1U figures on the left for orientation and angular velocity, we notice that the ground truth in red lies squarely on a surface containing numerous candidate solutions.
 This indicates that the local minima have correctly converged to the non-singular subspace of states that approximate the observations well for the assumed shape model.
@@
 
-The Case 1b results on the right within Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} show that the majority of the initial states sampled close to the truth have converged to a local subset of the solution manifolds identified by Case 1a.
+The Case 1P results on the right within Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} show that the majority of the initial states sampled close to the truth have converged to a local subset of the solution manifolds identified by Case 1U.
 We should not expect these solutions to have shrunk down exponentially towards the ground truth -- the mismatch in the assumed shape model and the measurement noise inherently limits the minimum size of the feasible solution space.
@@
     \caption{}
-    \label{fig:sigma_sols1a}
+    \label{fig:sigma_sols1U}
   \end{subfigure}
@@
     \caption{}
-    \label{fig:sigma_sols1b}
-  \end{subfigure}
-
-  \caption{Candidate solution initial orientation MRP vectors for Cases 1a (left) and 1b (right), with the ground truth highlighted in red.}
+    \label{fig:sigma_sols1P}
+  \end{subfigure}
+
+  \caption{Candidate solution initial orientation MRP vectors for Cases 1U (left) and 1P (right), with the ground truth highlighted in red.}
   \label{fig:sigma_sols1}
@@
     \caption{}
-    \label{fig:omega_sols1a}
+    \label{fig:omega_sols1U}
   \end{subfigure}
@@
     \caption{}
-    \label{fig:omega_sols1b}
-  \end{subfigure}
-  \caption{Candidate solution initial body-frame angular velocity vectors for Cases 1a (left) and 1b (right), with the ground truth highlighted in red.
-TODO: change the perspective on the angular velocities}
+    \label{fig:omega_sols1P}
+  \end{subfigure}
+  \caption{Candidate solution initial body-frame angular velocity vectors for Cases 1U (left) and 1P (right), with the ground truth highlighted in red.
+Note that each plot shows a unique perspective to highlight the structured distribution of solutions.}
   \label{fig:omega_sols1}
@@
 
-Figure \ref{fig:i_sols1} displays the inertia tensor solutions for Cases 1a and 1b.
+Figure \ref{fig:i_sols1} displays the inertia tensor solutions for Cases 1U and 1P.
 Notably, the majority of candidate solutions remain within a few standard deviations of the mean of their initial distribution, with the remaining minority exploring directions that maintain the same precession rate.
-When a good prior for the other attitude state components is available in Case 1b, the inertia tensor solutions are more well-clustered, although due to the low observability of the rocket body's spin rate in the observed light curve, there are still ambiguous solutions with higher or lower $J_z / J_x$ values.
+When a good prior for the other attitude state components is available in Case 1P, the inertia tensor solutions are more well-clustered, although due to the low observability of the rocket body's spin rate in the observed light curve, there are still ambiguous solutions with higher or lower $J_z / J_x$ values.
 
@@
     \caption{}
-    \label{fig:i_sols1a}
+    \label{fig:i_sols1U}
   \end{subfigure}
@@
     \caption{}
-    \label{fig:i_sols1b}
-  \end{subfigure}
-  \caption{Candidate solution inertia ratios for Cases 1a (left) and 1b (right), with the ground truth highlighted in red.}
+    \label{fig:i_sols1P}
+  \end{subfigure}
+  \caption{Candidate solution inertia ratios for Cases 1U (left) and 1P (right), with the ground truth highlighted in red.}
   \label{fig:i_sols1}
@@
 Figure \ref{fig:w_vs_ang_error_sols1} summarizes the distribution of candidate solutions with respect to the ground truth by displaying the angular velocity magnitude error and angular orientation error compared to the ground truth.
-Case 1a shows significant variability in angular velocity magnitude, although this is almost entirely due to the unobservability of the spin rate about the object's axis of symmetry in the light curve.
+Case 1U shows significant variability in angular velocity magnitude, although this is almost entirely due to the unobservability of the spin rate about the object's axis of symmetry in the light curve.
 The mean precession rate error is low at $0.3\%$.
+
 Due to the underdetermined orientation solution space, there are far more solutions at high angular offsets from the truth orientation, leading to a large mean angular error of $126^\circ$.
-For Case 1b, the mean orientation error is much lower at $18^\circ$, although convergence here is still inherently limited by the noisy measurements and model mismatch.
+For Case 1P, the mean orientation error is much lower at $18^\circ$, although convergence here is still inherently limited by the noisy measurements and model mismatch.
 
@@
     \caption{}
-    \label{fig:w_vs_ang_error_sols1a}
+    \label{fig:w_vs_ang_error_sols1U}
   \end{subfigure}
@@
     \caption{}
-    \label{fig:w_vs_ang_error_sols1b}
-  \end{subfigure}
-
-  \caption{Candidate solution angular orientation error and angular velocity magnitudes for Cases 1a (left) and 1b (right) compared to the ground truth}
+    \label{fig:w_vs_ang_error_sols1P}
+  \end{subfigure}
+
+  \caption{Candidate solution angular orientation error and angular velocity magnitudes for Cases 1U (left) and 1P (right) compared to the ground truth}
   \label{fig:w_vs_ang_error_sols1}
@@
 
-% While the state vectors of the candidate solutions do not uniquely indicate the ground truth, the object's precession rate is estimated
-
-% \begin{figure}[H]
-%   \centering
-%   \includegraphics[width=\figsmall]{spin_proc_Rocket Body.png}
-%   \caption{Converged solution body frame spin and precession rates.}
-%   \label{fig:spin_proc1}
-% \end{figure}
+While introducing a good prior in Case 1P does constrain the solution space, the ambiguities of the light curve inversion problem are still present and lead to the remaining variability we see in Figure \ref{fig:w_vs_ang_error_sols1P}.
+Our method respects this fact and resolves the structure of these surfaces, while other approaches that consider only one solution fail.
 
@@
 
-% The two objects selected for study in this work are a Delta I upper stage and the ECHOSTAR 2 satellite.
-
-\subsection{Case 2: Delta I Upper Stage}
-
-In Case 2, we perform attitude inversion on real data observed for a Delta I upper stage rocket body.
+In this section, we provide inversion results for a Delta I upper stage in Case 2D, and the ECHOSTAR II satellite in Case 2E, both observed by the Purdue Optical Ground Station.
+
+\subsection{Case 2D: Delta I Upper Stage}
+
+In Case 2D, we perform attitude inversion on real data observed for a Delta I upper stage rocket body.
 The chosen Delta I upper stage (COSPAR ID 1984-115C) is a Star-37E \cite{delta3914_astronautix} solid fuel apogee kick motor, displayed in Figure \ref{fig:star37e}.
@@
   \centering
-  \includegraphics[width=0.25\textwidth]{star37e.JPG}
-  \caption{Star-37E upper stage \cite{star37_af}.}
+  \begin{subfigure}[b]{0.13\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{star37e.JPG}
+    \caption{} 
     \label{fig:star37e}
-\end{figure}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\figbig]{sphx_glr_plot_models_002.png}
-  \caption{Simplified Star-37E model used for inversion.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.32\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{sphx_glr_plot_models_002.png}
+    \caption{}
     \label{fig:star37e_simple}
+  \end{subfigure}
+  \caption{a) Star-37E upper stage \cite{star37_af}, b) Simplified model used for inversion.}
+  \label{fig:star37e_combined}
 \end{figure}
@@
   \centering
-  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
+  \renewcommand{\arraystretch}{1.2} 
   \caption{Delta I rocket body observation characteristics}
@@
   \centering
-  \includegraphics[width=\figbig]{sphx_glr_plot_lcs_001_2_00x.png}
+  \includegraphics[width=\figmed]{sphx_glr_plot_lcs_001_2_00x.png}
   \caption{Selected light curve for the Delta I (Star-37E) rocket body, observed by the Purdue Optical Ground Station}
@@
 
-This light curve shows clear periodicity, but has a high measurement standard deviation and a relatively low sampling rate due to the object's low brightness and the sensor readout time.
-Inversion is performed using $n_\text{sample}=10^4$ samples with the material and angular velocity assumptions listed in Table \ref{tb:case2_ass}.
-As in Case 1a, orientation and angular velocity states are sampled via Equations \ref{eq:mrp_sampler} and \ref{eq:omega_sampler}, respectively.
+This light curve shows clear periodicity but has a high measurement standard deviation and a relatively low sampling rate due to the object's low brightness and the sensor readout time.
+Inversion is performed using $n_\text{sample}=10^4$ samples in $2.4$ hours on an Apple M1 CPU.
+The material and angular velocity assumptions are listed in Table \ref{tb:case2_ass}.
+As in Case 1U, orientation, and angular velocity states are sampled via Equations \ref{eq:mrp_sampler} and \ref{eq:omega_sampler}, respectively.
 All initial inertia tensor samples are set to the estimated $J_\text{R/B}$.
-The candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case2_s}, orientation MRPs and angular velocities in Figure \ref{fig:case2_pw}, and inertia ratios in Figure \ref{fig:case2_i}.
+The $82$ converged candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case2_s}, orientation MRPs and angular velocities in Figure \ref{fig:case2_pw}, and inertia ratios in Figure \ref{fig:case2_i}.
 
@@
   \centering
-  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
+  \renewcommand{\arraystretch}{1.2} 
   \caption{Delta I rocket body inversion assumptions}
@@
  Uniform BRDF & $C_d=0.2$, $C_s=0.8$, $n=30$ \\ \grule
-    $\norm{\vctr{\omega}}_\text{max}$ & $6.93$ $\text{deg} \cdot \text{s}^{-1}$ \\ \grule
+    $\norm{\vctr{\omega}}_\text{mid}$ & $6.93$ $[\text{deg} \cdot \text{s}^{-1}]$ \\ \grule
     $\ell$ & $1/100$ \\ \brule
@@
   \centering
-  \includegraphics[width=\figbig]{sphx_glr_results_002_2_00x.png}
+  \includegraphics[width=\figmed]{sphx_glr_results_002_2_00x.png}
   \caption{Candidate solution light curves compared to the real measurements in ADU for the Delta I (Star-37E) rocket body.}
@@
 
-% \begin{figure}[H]
-%   \centering
-%   \includegraphics[width=\figmed]{sphx_glr_results_006.png}
-%   \caption{}
-%   \label{fig:case2_p}
-% \end{figure}
-
 \begin{figure}[H]
@@
 \end{figure}
-
-% \begin{figure}[H]
-%   \centering
-%   \includegraphics[width=\figmed]{sphx_glr_results_005.png}
-%   \caption{Candidate solution angular velocities for the Delta I (Star-37E) rocket body}
-%   \label{fig:case2_w}
-% \end{figure}
 
@@
 
-The candidate solutions identified by the inversion process produce light curves which adequately approximate the measurements, although they do not reach the lowest observed values.
+The $82$ candidate solutions identified by the inversion process produce light curves that adequately approximate the measurements, although they do not reach the lowest observed values.
 This indicates that the true material is more dominated by specular reflections than our assumed coefficients.
@@
 Taken together, these results indicate that the object could be spinning in a number of attitude profiles, although the light curve inversion process has significantly narrowed the search space to small subsets of the original unconstrained eight-dimensional volume.
-These constraints can be combined with additional light curves, more information about the object's true mass distribution, or reflectivity to further refine the solution set.
-
-\subsection{Case 3: ECHOSTAR II}
-
-In Case 3, we perform attitude inversion on real data observed for a derelict satellite.
+These constraints can be combined with additional light curves,  information about the object's true mass distribution, or reflectivity to further refine the solution set.
+
+\subsection{Case 2E: ECHOSTAR II}
+
+In Case 2E, we perform attitude inversion on real data observed for a derelict satellite.
 ECHOSTAR II (COSPAR ID 1996-055A) was a communications satellite on the AS-7000 bus \cite{as7000_astronautix}, shown in Figure \ref{fig:echostar1}.
@@
   \centering
-  \includegraphics[width=\figbig]{EchoStar-1_image.jpg}
-  \caption{ECHOSTAR II artist's rendition \cite{as7000_astronautix}.}
+  \begin{subfigure}[b]{0.35\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{EchoStar-1_image.jpg}
+    \caption{}
     \label{fig:echostar1}
-\end{figure}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\figbig]{sphx_glr_plot_models_001.png}
-  \caption{Simplified ECHOSTAR II box-wing model used for inversion.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.35\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{sphx_glr_plot_models_001.png}
+    \caption{}
     \label{fig:echostar1_simple}
+  \end{subfigure}
+  \caption{a) Artist's rendition of ECHOSTAR II \cite{as7000_astronautix}, b) Simplified ECHOSTAR II box-wing model used for inversion.}
+  \label{fig:echostar}
 \end{figure}
@@
   \centering
-  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
+  \renewcommand{\arraystretch}{1.2} 
   \caption{ECHOSTAR II observation characteristics}
@@
   \centering
-  \includegraphics[width=\figbig]{sphx_glr_plot_lcs_002_2_00x.png}
+  \includegraphics[width=\figmed]{sphx_glr_plot_lcs_002_2_00x.png}
   \caption{Selected light curve for ECHOSTAR II, observed by the Purdue Optical Ground Station}
@@
 
-This light curve is sampled more densely than the Delta I observations, and clearly has significantly more intense specular peaks.
-Inversion is performed using $n_\text{sample}=10^4$ samples with the material and angular velocity assumptions listed in Table \ref{tb:case3_ass}.
-Initial state sampling is accomplished identically to Case 2, with a different maximum angular veloity and initial inertia ratios.
+This light curve is sampled more densely than the Delta I observations and clearly has significantly more intense specular peaks.
+Inversion is performed using $n_\text{sample}=10^4$ samples in $48$ minutes on an Apple M1 CPU.
+The material and angular velocity assumptions are listed in Table \ref{tb:case3_ass}.
+Initial state sampling is accomplished identically to Case 2D, with a different mean angular velocity and initial inertia ratios.
 The candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case3_s}, orientation MRPs and angular velocities in Figure \ref{fig:case3_pw}, and inertia ratios in Figure \ref{fig:case3_i}.
@@
   \centering
-  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
+  \renewcommand{\arraystretch}{1.2} 
   \caption{ECHOSTAR II inversion assumptions}
@@
  Solar panel BRDF & $C_d=0.05$, $C_s=0.13$, $n=10$ \\ \grule
-    $\norm{\vctr{\omega}}_\text{max}$ & $2.32$ $\text{deg} / \text{s}^{-1}$ \\ \grule
+    $\norm{\vctr{\omega}}_\text{mid}$ & $2.32$ $[\text{deg} \cdot \text{s}^{-1}]$ \\ \grule
     $\ell$ & $1/1000$ \\ \brule
@@
   \centering
-  \includegraphics[width=\figbig]{sphx_glr_results_004_2_00x.png}
+  \includegraphics[width=\figmed]{sphx_glr_results_004_2_00x.png}
   \caption{Candidate solution light curves compared to the real measurements in ADU for ECHOSTAR II.}
@@
   \includegraphics[width=\figsmall]{sphx_glr_results_003_2_00x.png}
-  \caption{Candidate solution inertia ratio distribution for ECHOSTAR II with initial estimate in red.}
+  \caption{Candidate solution inertia ratio distribution for ECHOSTAR II with the initial estimate in red.}
   \label{fig:case3_i}
@@
 
-% \begin{figure}[H]
-%   \centering
-%   \includegraphics[width=\figmed]{sphx_glr_results_008.png}
-%   \caption{Candidate solution orientation MRPs for ECHOSTAR II}
-%   \label{fig:case3_p}
-% \end{figure}
-
-% \begin{figure}[H]
-%   \centering
-%   \includegraphics[width=\figmed]{sphx_glr_results_007.png}
-%   \caption{Candidate solution angular velocities for ECHOSTAR II}
-%   \label{fig:case3_w}
-% \end{figure}
-
-The candidate solutions identified by the inversion process are more sparse than those found in either of the previous rocket body cases, with some vague clustering in both angular velocity and orientation space.
-That said, the candidate light curves fit the data quite well.
+The $19$ candidate solutions identified by the inversion process are more sparse than those found in either of the previous rocket body cases, with some vague clustering in both angular velocity and orientation space.
+The candidate solution light curves fit the data well.
 Notably, the formulation of the negative log-likelihood objective function in Equation \ref{eq:nll_loss} produces the behavior seen in the candidate solutions in Figure \ref{fig:case3_s}, where even the smallest recurrent specular peaks are well-estimated by most solutions.
@@
 The inertia tensor solutions are dispersed throughout the solution space, with less symmetry than the Delta I solutions.
+
 As in the Delta I case, the attitude inversion solver has successfully constrained the search space for later optimization as more data about the mass and reflectivity of the object becomes available to continue refining any of these candidates.
-
-\section{Conclusion}
+To address the unknown solar panel rotation angle at the time of ECHOSTAR II's retirement, a parametric model for the glint timing of the bus and panels could be employed to build an \textit{a priori} estimate of this angle for more accurate full-state solutions.
+
+\section{Conclusions}
 
 Attitude information is critical for many tasks informing space situational awareness.
-Orientation and angular velocity data aids high-fidelity orbit propagation, mission recovery efforts, and object selection for active debris removal.
+Orientation and angular velocity data aid high-fidelity orbit propagation, mission recovery efforts, and object selection for active debris removal.
 Obtaining attitude information is often difficult, especially for passive debris objects.
-Due to atmospheric distortion and diffraction-limited optics, ground-based telescopes can only determine a total brightness received from a space object.
-If a shape model is known, the light curve -- a sequence of these brightness measurements -- is one tool for attitude estimation.
+Due to atmospheric distortion and diffraction-limited optics, ground-based telescopes can only determine the total brightness received from a space object.
+If a shape model is known, the light curve -- a sequence of these brightness measurements -- is one tool for attitude estimation in the far field.
+
 We presented an approach for solving this highly nonlinear and non-convex optimization problem with a gradient-based multi-start method.
 We presented attitude inversion results for one case with synthetic data and two cases with real observations.
-Unlike other approaches, our method successfully identifies the numerous ambiguous solutions families -- originating from symmetries in the observation geometry and the object shape -- which can produce the same observations.
+Results on synthetic data prove that our algorithm converges well to the full set of ambiguous solutions, and that performance improves as more prior information becomes available.
+Results on real data show that our approach can produce viable clusters of potential full-state solutions even when material properties are unavailable and the assumed geometry is heavily simplified, although convergence becomes less frequent.
+
+Unlike other approaches, our method successfully identifies the numerous ambiguous solution families -- originating from symmetries in the observation geometry and the object shape -- which can produce the same observations.
 These disparate candidate solutions can serve as starting points for more detailed exploration as additional light curves or further object shape information is obtained.
 
-\section{Global TODO list}
-
-\begin{itemize}
-  % \item Add the constant term back into the objective function (in text and code) so we can start treating it as a real likelihood, useful for the future
-  \item Mention the unknown solar panel rotation angle for ECHOSTAR
-  \item make a table of execution times
-  % \item Update perspective on some the case 1a angular velocity plot
-  % \item Update $n_\text{sample}$ values everywhere to reflect the scripts
-  % \item Update $\omega$ sampling scheme to go from $0.5-2$ as reflected in scripts
-  % \item Re-run all cases to ensure that values are up to date with recent changes, update where those statistics are reported in the text
-  \item Try to get rid of big gaps where possible
-  % \item Reformat tables so they're attractive to look at
-\end{itemize}
-
 \section*{Acknowledgments}
@@
   \centering
-  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
+  \renewcommand{\arraystretch}{1.2} 
   \caption{Telescope information for synthetic and real data observations}
@@
   \centering
-  \renewcommand{\arraystretch}{1.3} % Increase row height for better readability
-  \caption{BFGS solver configuration for all results cases}
+  \renewcommand{\arraystretch}{1.2} 
+  \caption{BFGS configuration for all results cases}
   \vspace*{6pt}
